### 2012-11-20 Assignment6/paper/paper.tex
Working on paper

% Paper for the ChouchBase assignment.

\documentclass[]{IEEEtran}

\usepackage{float}
\usepackage{url}
\usepackage{graphicx}
\usepackage{color}

\begin{document}

\title{No Dogs On The Couch$_{Base}$}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations

\author{\IEEEauthorblockN{Eriq Augustine, Ryan Hnarkis, Aldrin Montana, Ryan Verdon, Tyler Yero}
\\
\IEEEauthorblockA{Department of Computer Science\\
Cal Poly, San Luis Obispo\\
 \textsf{\{eaugusti, rhnaraki, amontana, rverdon, tyero\}@calpoly.edu}
}
}

\maketitle

\thispagestyle{empty}
\pagestyle{empty}

\section{Introduction}
TODO

\section{Experiment Design}
TODO

\section{Implementation}
TODO

\section{Evaluation}
TODO

\section{Analysis}
TODO

\section{Conclusions}
TODO

% TODO: Uncomment if we use any refs
% \bibliographystyle{acm}
% \bibliography{refs}

\end{document}

### 2012-11-20 Assignment6/MoveSQLDataToCouch.java
Finished JSON generation need to ship the JSON to couchbase now.

@@ -152,7 +152,7 @@ public class MoveSQLDataToCouch {
             memberNames.add(uname);
          }
          //toJSON
-         String JSON = groupToJSON(groupId, name, desc, createDate, memberIds, memberNames); 
+         String JSON = groupToJSON(groupId, name, desc, createDate, memberIds, memberNames);
          //SEND TO COUCHBASE    
       }
 
@@ -163,7 +163,7 @@ public class MoveSQLDataToCouch {
 
 
    private static void moveContigs(Connection conn) throws Exception {
-      String contigsQuery = "SELECT ContigId, Name, Difficulty, Sequence, UploaderId, Source, Species, Status, CreateDate FROM Contigs;";
+      String contigsQuery = "SELECT ContigId, Name, Difficulty, Sequence, UploaderId, UserName, Source, Species, Status, CreateDate FROM Contigs, Users WHERE Contigs.UploaderId = Users.UserId;";
       String expertAnnoQuery = "SELECT AnnotationId FROM Annotations WHERE ExpertSubmission = 1 AND ContigId = ?;";
       String geneAnnoQuery = " SELECT A.AnnotationId, G.Name FROM Annotations A, GeneNames G WHERE A.GeneId = G.GeneId AND A.ContigId = ?;";
       PreparedStatement contigQ = conn.prepareStatement(contigsQuery);
@@ -177,6 +177,7 @@ public class MoveSQLDataToCouch {
          int diff = rs.getInt("Difficulty");
          String seq = rs.getString("Sequence");
          int uploader = rs.getInt("UploaderId");
+         String uploader_name = rs.getString("UserName");
          String source = rs.getString("Source");
          String species = rs.getString("Species");
          String status = rs.getString("Status");
@@ -207,7 +208,7 @@ public class MoveSQLDataToCouch {
 
          //toJSON
          String JSON = contigToJSON(contigId, name, diff, seq, uploader, source, species,
-                                    status, create, expertAnnos, isoforms);     
+                                    status, create, uploader_name, expertAnnos, isoforms);
          //SEND TO COUCHBASE  
       }
 
@@ -254,7 +255,7 @@ public class MoveSQLDataToCouch {
                                               partial, expert, contigId, userId, createDate,
                                               lastModifiedDate, finishedDate, incorrect, 
                                               expertIncorrect, exonStartEndPairs);
-
+         
          //SEND to couchbase
       }
 
@@ -421,9 +422,9 @@ public class MoveSQLDataToCouch {
[…]

### 2012-11-20 Assignment6/MoveSQLDataToCouch.java
Can now move sql data into couchbase. example_json.txt contains one of each type of json object that is stored in couchbase for reference.

@@ -10,6 +10,15 @@ import java.util.Random;
 import java.util.ArrayList;
 import java.util.HashMap;
 
+import java.net.URI;
+import java.util.LinkedList;
+import java.util.List;
+import java.util.concurrent.TimeUnit;
+
+import com.couchbase.client.CouchbaseClient;
+import net.spy.memcached.internal.GetFuture;
+import net.spy.memcached.internal.OperationFuture;
+
 
 
 /**
@@ -20,36 +29,42 @@ public class MoveSQLDataToCouch {
    private static final String DB_USER = "";
    private static final String DB_PASS = "";
 
+   private static List<URI> uris;
+
    public static void main(String[] args) throws Exception {
       Connection conn = null;
+      CouchbaseClient client = null;
+      uris = new LinkedList<URI>();
+      uris.add(URI.create("http://127.0.0.1:8091/pools"));
 
     
       try {
          // Instantiate the DB Driver
          Class.forName("com.mysql.jdbc.Driver");
-
          conn = DriverManager.getConnection(DB_URL, DB_USER, DB_PASS);
+
+         client = new CouchbaseClient(uris, "default", "");
       } catch (Exception ex) {         
          System.err.println("Cannot connect to DB: " + ex);
       }
       
       System.out.println("Starting to move the stuff...\n");
-      moveUsers(conn);
+      moveUsers(conn, client);
       System.out.println("-Finished moving users.");
-      moveGroups(conn);
+      moveGroups(conn, client);
       System.out.println("-Finished moving groups.");
-      moveContigs(conn);
+      moveContigs(conn, client);
       System.out.println("-Finished moving contigs.");
-      moveAnnotations(conn);
+      moveAnnotations(conn, client);
       System.out.println("-Finished moving annotations.");
-      //moveCollabAnnotations(conn);
-      //System.out.println("-Finished moving collab annotations.");
      
       System.out.println("\nDone!");
+      conn.close();
+      client.shutdown();
    }
 
 
-   private static void moveUsers(Connection conn) throws Exception {
+   private static void moveUsers(Connection conn, CouchbaseClient client) throws Exception {
       String usersQuery = "SELECT UserId, FirstName, LastName, UserName, Email, Pass, Salt, LastLoginDate, RegistrationDate, Level, Role, Exp FROM Users;";
[…]

### 2012-11-20 Assignment6/paper/paper.tex
Basic introduction done. Removed one of the nested evaluation sections of analysis about conclusions of comparisons of data about the results of our experiments on the evaluation on couchbase and mysql.

@@
 \section{Introduction}
-TODO
+Our team decided to compare a MySQL cluster versus a Couchbase cluster for running a
+web application. We are using an eight table SQL schema from Ryan Verdon's Thesis. To evaluate
+the performance we autogenerated around 400MB of data and inserted the same data into each cluster.
+Obviously using a different data model for Couchbase to accomodate for the fact it is a NoSQL key-value store
+instead of a relational database. We then created a set of six experiments to run. Five came from common 
+use cases found in the web application. While the other was designed to test the performance on each cluster
+of repeatedly writing and reading from the same object. Our major metrics are write performance, read performance,
+and time-delay between writes and reads of the same data. The rest of the paper is broken down as follows.
+Section 2 describes in depth how we designed the experiment. Section 3 discusses our implementation. Section 4 goes
+into the results of each of our experiments and some analysis on the data we obtained. Lastly, section 5 presents 
+our conclusions.
 
@@
 
-\section{Analysis}
-TODO
-
 \section{Conclusions}

### 2012-11-20 Assignment6/JSONExample.java
Added a JSON library, and an exmaple file for how to use it.

@@ -1,0 +1,73 @@
+import org.json.JSONException;
+import org.json.JSONObject;
+
+import java.util.HashMap;
+import java.util.Map;
+
+/**
+ * This is an example to show everyone how to use the JSON library.
+ */
+public class JSONExample {
+   // Note that most JSON conversions will throw a JSONException if they are given inproper JSON.
+   public static void main(String[] args) throws JSONException {
+      JSONObject json = null;
+
+      // This the most popular constructor.
+      // Just jove is the JSON text.
+      // Note: Single quotes are not valid json, but this library will accept it.
+      String string1 = "{'id': 1}";
+      System.out.println("Parsing: " + string1);
+      json = new JSONObject(string1);
+      System.out.println("JSONObject: " + json);
+      // Note that I pull out the field with the type I know it to be.
+      // There are variations for all valid JSON types.
+      // There is a general get() that returns sn Object.
+      System.out.println("'id' field: " + json.getInt("id"));
+      System.out.println();
+
+      try {
+         // Trying to get a non-existant field will throw a JSONException.
+         System.out.println("'id' field: " + json.getInt("iDontExist"));
+      } catch (JSONException jsonEx) {
+         System.out.println("Threw an exception.");
+      }
+
+      // Use has() if you want to check existance.
+      System.out.println("Does 'iDontExist' exist?: " + json.has("iDontExist"));
+      System.out.println();
+
+
+      // There are many overrides of put() that take the key and the value.
+      // Therefore the value can be an int, bool, Map (will be converted to a JSONObject),
+      // JSONObject, etc.
+      json.put("newIntField", 999);
+      json.put("newBool", true);
+
+      Map myMap = new HashMap<String, String>();
+      myMap.put("hi", "there");
+      myMap.put("foo", "bar");
+      json.put("newMapField", myMap);
+
+      JSONObject newObj = new JSONObject("{'name': 'Aldrin', 'val': 3.4, 'present': true}");
+      json.put("newObjField", newObj);
+
+      System.out.println("New JSONObject: " + json);
+      System.out.println();
+
+      // When you want to set a field, you have to totall override it.
+      // Note that put() returns this (the JSONObject) so you can chain them.
+      json.put("newObjField", json.getJSONObject("newObjField").put("name", "Ryan").put("present", false));
+      System.out.println("Modified JSONObject: " + json);
+      System.out.println();
+
+      // This has some of the same itteration methods as Map: keys() and keySet().
+
[…]

### 2012-11-20 Assignment6/JSONExample.java
Modified a comment

@@ -6,6 +6,7 @@ import java.util.Map;
 
 /**
  * This is an example to show everyone how to use the JSON library.
+ * See here for a complete reference: http://www.json.org/java/
  */
 public class JSONExample {
    // Note that most JSON conversions will throw a JSONException if they are given inproper JSON.

### 2012-11-20 Assignment6/paper/paper.tex

@@
 
+\section{Workloads}
+The following workloads were applied to the both the MySQL and Couchbase clusers.
+\subsection{Assigning a Task}
+TODO
+\subsection{Profiling}
+TODO
+\subsection{Read and Writing Performance}
+TODO
+\subsection{Publishing}
+TODO
+\subsection{Adding and Removing Group Membership}
+Adding or removing users to a group is a task common for classroom situations.
+In MySQL, this operation requires a join between three tables: Groups, Users, and GroupMembership.
+In Couchbase, this requires retrieving the key for the desired group and removing the user from the members field.
+
 \section{Implementation}

### 2012-11-21 Assignment6/ContigUploadWorkload.java
Couch Profile and infrastructure

@@ -83,9 +83,6 @@ public class ContigUploadWorkload extends Workload {
       return new Stats();
    }
 
-   protected void initCouch() {
-   }
-
    protected Stats executeCouchImpl() {
       // Nope
       throw new UnsupportedOperationException();

### 2012-11-21 Assignment6/ProfileWorkload.java

@@ -8,11 +8,16 @@ import java.sql.ResultSet;
 import java.sql.SQLException;
 import java.sql.Statement;
 
+import org.json.JSONArray;
+import org.json.JSONException;
+import org.json.JSONObject;
+
 /**
  * This one measures fetching a complete user's profile.
  */
 public class ProfileWorkload extends Workload {
-   // Should be no more than 100000 (the number of users in the db).
+   private static final int MAX_USERS = 100000;
+
    private static final int TIMES = 1000000;
    //private static final int TIMES = 1;
 
@@ -79,10 +84,80 @@ public class ProfileWorkload extends Workload {
    }
 
    protected void initCouch() {
+      super.initCouch();
+      Random rand = new Random(4);
+
+      for (int i = 0; i < TIMES; i++) {
+         userIds[i] = "" + (rand.nextInt(MAX_USERS) + 1);
+      }
+   }
+
+   protected void cleanupCouch() {
+      super.cleanupCouch();
+
+      userIds = null;
    }
 
    protected Stats executeCouchImpl() {
-      // Nope
-      throw new UnsupportedOperationException();
+      String data = null;
+      JSONObject jsonUser = null;
+      JSONArray groups = null;
+      JSONArray history = null;
+      JSONArray partials = null;
+      JSONArray tasks = null;
+
+      // THis shouldn't even be necessary, but I am being neurotic.
+      Object throwAway = null;
+
+      for (int i = 0; i < TIMES; i++) {
+         try {
+            // I saw the api just casting, so I will too.
+            data = (String)client.get("Users-" + i);
+            jsonUser = new JSONObject(data);
+
+            // Expand groups
+            groups = jsonUser.getJSONArray("groups");
+            for (int j = 0; j < groups.length(); j++) {
+               String groupId = groups.getString(j);
+
+               // The next step would be trivial, get the groups name form the JSON.
+               throwAway = client.get("Groups-" + groupId);
+            }
+
+            // Expand history
+            history = jsonUser.getJSONArray("history");
+            for (int j = 0; j < history.length(); j++) {
+               JSONObject historyJson = history.getJSONObject(j);
+               String annotationId = historyJson.getString("anno_id");
+
+               throwAway = client.get("Annotations-" + annotationId);
+            }
+
+            // Expand partials
+            partials = jsonUser.getJSONArray("incomplete_annotations");
+            for (int j = 0; j < partials.length(); j++) {
+               String annotationId = partials.getString(j);
+
[…]

### 2012-11-21 Assignment6/PublishWorkload.java

@@ -86,9 +86,6 @@ public class PublishWorkload extends Workload {
       return new Stats();
    }
 
-   protected void initCouch() {
-   }
-
    protected Stats executeCouchImpl() {
       // Nope
       throw new UnsupportedOperationException();

### 2012-11-21 Assignment6/StupidWorkload.java

@@ -83,9 +83,6 @@ public class StupidWorkload extends Workload {
       return new ReadWriteStats(readAfterWriteTimes, writeAfterReadTimes);
    }
 
-   protected void initCouch() {
-   }
-
    protected Stats executeCouchImpl() {
       // Nope
       throw new UnsupportedOperationException();

### 2012-11-21 Assignment6/Workload.java

@@ -1,11 +1,18 @@
+import java.util.List;
+import java.util.LinkedList;
+import java.net.URI;
+
 import java.sql.Connection;
 import java.sql.DriverManager;
+
+import com.couchbase.client.CouchbaseClient;
 
 /**
  * The base class for all workloads.
  */
 public abstract class Workload {
    protected Connection conn;
+   protected CouchbaseClient client;
 
    // See if we can get a clean connection.
    public static void main(String[] args) {
@@ -84,7 +91,24 @@ public abstract class Workload {
       }
    }
 
+   /**
+    * Do whatever setup must be done to initialize the Couch version of the workload.
+    */
+   protected void initCouch() {
+      try {
+         List<URI> uris = new LinkedList<URI>();
+         uris.add(URI.create(TestMaster.COUCH_URI));
+         client = new CouchbaseClient(uris, "default", "");
+      } catch (Exception ex) {
+         System.err.println("Failed to get the CouchBase Connection: " + ex);
+         ex.printStackTrace(System.err);
+         throw new RuntimeException();
+      }
+   }
+
    protected void cleanupCouch() {
+      client.shutdown();
+      client = null;
    }
 
    /**
@@ -92,10 +116,6 @@ public abstract class Workload {
     */
    protected abstract Stats executeMySQLImpl();
 
-   /**
-    * Do ehatever setup must be done to initialize the Couch version of the workload.
-    */
-   protected abstract void initCouch();
 
    /**
     * The actual work for the CouchDB.

### 2012-11-21 Assignment6/PublishWorkload.java
Committing PublishWorkload so I can turn my computer off because eriq sucks. Also, it doesn't work.

@@ -1,5 +1,8 @@
-import java.util.ArrayList;
+import java.util.regex.Pattern;
+import java.util.regex.Matcher;
+import java.util.Map;
 import java.util.List;
+import java.util.ArrayList;
 import java.util.Random;
 
 import java.sql.Connection;
@@ -10,14 +13,22 @@ import java.sql.Statement;
 import java.text.SimpleDateFormat;
 import java.sql.Date;
 
+import org.json.JSONException;
+import org.json.JSONObject;
+
+import com.couchbase.client.CouchbaseClient;
+import net.spy.memcached.internal.GetFuture;
+import net.spy.memcached.internal.OperationFuture;
+
 /**
  * This one measures fetching a complete user's profile.
  */
 public class PublishWorkload extends Workload {
-   private static final int TIMES = 1000000;
-   //private static final int TIMES = 10;
+   //private static final int TIMES = 1000000;
+   private static final int TIMES = 10;
 
    private String[] userIds, annotationIds, exp;
+   List<String> annotationIdList, userIdList;
 
    /**
     * Fetch the user ids here, and preserve the connection.
@@ -50,6 +61,7 @@ public class PublishWorkload extends Workload {
    protected void cleanupMySQL() {
       super.cleanupMySQL();
       userIds = null;
+      annotationIds = null;
    }
 
    protected Stats executeMySQLImpl() {
@@ -87,10 +99,43 @@ public class PublishWorkload extends Workload {
    }
 
    protected void initCouch() {
+      super.initCouch();
+
+      Random rand = new Random(4);
+
+      annotationIdList = new ArrayList<String>(TIMES);
+      userIdList = new ArrayList<String>(TIMES);
+
+      for (int repeat = 0; repeat < TIMES; repeat++) {
+         annotationIdList.add(String.format("Annotations-%d", rand.nextInt(1000)));
+         userIdList.add(String.format("Users-%d", rand.nextInt(100)));
+      }
    }
 
    protected Stats executeCouchImpl() {
-      // Nope
-      throw new UnsupportedOperationException();
+      JSONObject userJSON = null, annotationJSON = null, contigJSON = null;
+
+      for (int iter_ndx = 0; iter_ndx < TIMES; iter_ndx++) {
+         Object user = client.get(userIdList.get(iter_ndx));
+         Object annotation = client.get(annotationIdList.get(iter_ndx));
+
+         try {
+            userJSON = new JSONObject(user.toString());
+            annotationJSON = new JSONObject(annotation.toString());
+
[…]

### 2012-11-21 Assignment6/PublishWorkload.java
Re-committing PublishWorkload to have all non-working code (probably not working code) commented out. That way this commit shouldn't break anything but I can continue where I left off

@@ -100,6 +100,7 @@ public class PublishWorkload extends Workload {
 
    protected void initCouch() {
       super.initCouch();
+      /*
 
       Random rand = new Random(4);
 
@@ -110,9 +111,11 @@ public class PublishWorkload extends Workload {
          annotationIdList.add(String.format("Annotations-%d", rand.nextInt(1000)));
          userIdList.add(String.format("Users-%d", rand.nextInt(100)));
       }
+      */
    }
 
    protected Stats executeCouchImpl() {
+      /*
       JSONObject userJSON = null, annotationJSON = null, contigJSON = null;
 
       for (int iter_ndx = 0; iter_ndx < TIMES; iter_ndx++) {
@@ -137,5 +140,7 @@ public class PublishWorkload extends Workload {
       }
 
       return new Stats();
+      */
+      throw new UnsupportedOperationException();
    }
 }

### 2012-11-21 Assignment6/ContigUploadWorkload.java
Roughly finished couchbase version of contig uploading. Still need to test it, getting an annoying error with couchbase.

@@ -2,8 +2,12 @@ import java.util.ArrayList;
 import java.util.List;
 import java.util.Random;
 import java.util.Date;
+import java.util.HashMap;
 
 import java.text.SimpleDateFormat;
+
+import org.json.JSONException;
+import org.json.JSONObject;
 
 import java.sql.Connection;
 import java.sql.DriverManager;
@@ -16,7 +20,7 @@ import java.sql.PreparedStatement;
  */
 public class ContigUploadWorkload extends Workload {
    // Should be no more than 9999999999999999999999999999
-   private static final int TIMES = 10000;
+   private static final int TIMES = 10;
    //private static final int TIMES = 5;
    private static final int MIN_SEQ_LENGTH = 45000;
    private static final int MAX_ADDITIONAL_LENGTH = 15001;
@@ -84,8 +88,33 @@ public class ContigUploadWorkload extends Workload {
    }
 
    protected Stats executeCouchImpl() {
-      // Nope
-      throw new UnsupportedOperationException();
+      String json = "";
+      
+      for (int i = 0; i < TIMES; i++) {
+         int uploaderId = rand.nextInt(TIMES)+1;
+         int newContigId = 9000 + i;
+         String user_json = (String)client.get("Users-1");
+
+         try {
+            JSONObject user = new JSONObject(user_json);
+         
+            json = contigToJSON(newContigId, "contig-" + randomString(15), rand.nextInt(10)+1, 
+                                      seqs.get(rand.nextInt(seqs.size())), 
+                                      uploaderId, randomString(7), randomString(20), 
+                                      randomString(10), new Date(), user.getString("meta.user_name"),
+                                      new ArrayList<Integer>(),
+                                      new HashMap<String, ArrayList<Integer>>());
+         }
+         catch (Exception e) {
+            System.err.println("Caught an error uploading a contig. ERROR: " + e);
+         }
+
+         client.set("Contigs-" + newContigId, 0, json); 
+      }      
+
+
+      json = null;
+      return new Stats();
    }
 
    private String randomSequence(int length) {
@@ -113,6 +142,68 @@ public class ContigUploadWorkload extends Workload {
    private String randomDate() {
       SimpleDateFormat formatter = new SimpleDateFormat("yyyy-MM-dd HH:mm:ss");
       return formatter.format(new Date());
+   }
+
+   private static String contigToJSON(int contigId, String name, int diff, String seq, 
[…]

### 2012-11-21 Assignment6/ContigUploadWorkload.java
ContigUploadWorkload fully tested and operational.

@@ -53,6 +53,15 @@ public class ContigUploadWorkload extends Workload {
       }
    }
 
+   protected void initCouch() {
+      super.initCouch();
+
+      for (int i = 0; i < TIMES; i++) {
+         int extra_length = rand.nextInt(MAX_ADDITIONAL_LENGTH);
+         seqs.add(randomSequence(MIN_SEQ_LENGTH + extra_length));
+      }
+   }
+
    protected void cleanupMySQL() {
       super.cleanupMySQL();
 
@@ -93,15 +102,16 @@ public class ContigUploadWorkload extends Workload {
       for (int i = 0; i < TIMES; i++) {
          int uploaderId = rand.nextInt(TIMES)+1;
          int newContigId = 9000 + i;
-         String user_json = (String)client.get("Users-1");
+         String user_json = (String)client.get("Users-"+uploaderId);
 
          try {
             JSONObject user = new JSONObject(user_json);
+            String name = user.getJSONObject("meta").getString("user_name");
          
             json = contigToJSON(newContigId, "contig-" + randomString(15), rand.nextInt(10)+1, 
                                       seqs.get(rand.nextInt(seqs.size())), 
                                       uploaderId, randomString(7), randomString(20), 
-                                      randomString(10), new Date(), user.getString("meta.user_name"),
+                                      randomString(10), new Date(), name,
                                       new ArrayList<Integer>(),
                                       new HashMap<String, ArrayList<Integer>>());
          }

### 2012-11-21 Assignment6/StupidWorkload.java
Fixed bug in SQL of StupidWorkload, it was using an old table.

@@ -34,7 +34,7 @@ public class StupidWorkload extends Workload {
 
    protected Stats executeMySQLImpl() {
       String readQuery = "SELECT name FROM ReadWriteTest WHERE id = 1";
-      String writeUpdate = "UPDATE users SET name = '%s' WHERE id = 1";
+      String writeUpdate = "UPDATE ReadWriteTest SET name = '%s' WHERE id = 1";
       List<Long> readAfterWriteTimes = new ArrayList<Long>(TIMES);
       List<Long> writeAfterReadTimes = new ArrayList<Long>(TIMES);
 

### 2012-11-21 Assignment6/GroupMemWorkload.java
Some tuning for runs

@@ -13,7 +13,7 @@ import java.sql.Statement;
  */
 public class GroupMemWorkload extends Workload {
    // Should be no more than 100000 (the number of users in the db).
-   private static final int TIMES = 1000000;
+   private static final int TIMES = 100000;
    //private static final int TIMES = 1;
 
    private String[] userIds;

### 2012-11-21 Assignment6/PublishWorkload.java

@@ -14,7 +14,7 @@ import java.sql.Date;
  * This one measures fetching a complete user's profile.
  */
 public class PublishWorkload extends Workload {
-   private static final int TIMES = 1000000;
+   private static final int TIMES = 100000;
    //private static final int TIMES = 10;
 
    private String[] userIds, annotationIds, exp;

### 2012-11-21 Assignment6/StupidWorkload.java

@@ -13,7 +13,7 @@ import java.sql.Statement;
  * Ryan thinks that it is stupid.
  */
 public class StupidWorkload extends Workload {
-   private static final int TIMES = 1000000;
+   private static final int TIMES = 100000;
    //private static final int TIMES = 10;
 
    private String[] randomStrings;

### 2012-11-21 Assignment6/TestMaster.java

@@ -18,11 +18,11 @@ public class TestMaster {
 
    public static void main(String[] args) {
       Workload[] workloads = {
-                              new ContigUploadWorkload(),
-                              new ProfileWorkload(),
-                              new AssignTaskWorkload(),
-                              new GroupMemWorkload(),
-                              new PublishWorkload(),
+                              //new ContigUploadWorkload(),
+                              //new ProfileWorkload(),
+                              //new AssignTaskWorkload(),
+                              //new GroupMemWorkload(),
+                              //new PublishWorkload(),
                               new StupidWorkload(),
                               };
 

### 2012-11-21 Assignment6/StupidWorkload.java
StupidWorkload Couch complete, but not tested.

@@ -84,7 +84,25 @@ public class StupidWorkload extends Workload {
    }
 
    protected Stats executeCouchImpl() {
-      // Nope
-      throw new UnsupportedOperationException();
+      List<Long> readAfterWriteTimes = new ArrayList<Long>(TIMES);
+      List<Long> writeAfterReadTimes = new ArrayList<Long>(TIMES);
+      long time = System.currentTimeMillis();
+      Object throwAway;
+
+      for (int i = 0; i < TIMES; i++) {
+         throwAway = client.get("ReadAfterWrite-1");
+
+         if (i != 0) {
+            readAfterWriteTimes.add(System.currentTimeMillis() - time);
+         }
+
+         time = System.currentTimeMillis();
+         client.set("ReadAfterWrite-1", 0, randomStrings[i]);
+         writeAfterReadTimes.add(System.currentTimeMillis() - time);
+
+         time = System.currentTimeMillis();
+      }
+
+      return new ReadWriteStats(readAfterWriteTimes, writeAfterReadTimes);
    }
 }

### 2012-11-21 Assignment6/AssignTaskWorkload.java
Did couch impl for AssignTask. Might want to move taskToJSON helper method to another class

@@ -35,22 +35,54 @@ public class AssignTaskWorkload extends Workload {
    }
 
    protected Stats executeMySQLImpl() {
-         String assignTaskInsert = String.format("REPLACE INTO Tasks (UserId, ContigId, Description, EndDate)" +
-                                                 " SELECT UserId, %d, '%s', '%s'" +
-                                                 " FROM GroupMembership" +
-                                                 " WHERE GroupId = %%d",
-                                                 CONTIG_ID, DESCRIPTION, DATE);
+      String assignTaskInsert = String.format("REPLACE INTO Tasks (UserId, ContigId, Description, EndDate)" +
+                                              " SELECT UserId, %d, '%s', '%s'" +
+                                              " FROM GroupMembership" +
+                                              " WHERE GroupId = %%d",
+                                              CONTIG_ID, DESCRIPTION, DATE);
 
       for (int i = 0; i < TIMES; i++) {
-         Util.doUpdate(conn, String.format(
-               assignTaskInsert, (rand.nextInt() % MAX_GROUP_ID) + 1));
+         Util.doUpdate(conn, String.format(assignTaskInsert, (rand.nextInt() % MAX_GROUP_ID) + 1));
       }
 
       return new Stats();
    }
 
-   protected Stats executeCouchImpl() {
-      // Nope
-      throw new UnsupportedOperationException();
+   protected Stats executeCouchImpl() {	  
+	  String data = null;
+      JSONObject jsonUser = null;
+	  JSONObject jsonGroup = null;
+      JSONArray users = null;
+	  
+      for (int i = 0; i < TIMES; i++) {
+         try {
+            data = (String)client.get("Groups-" + (rand.nextInt() % MAX_GROUP_ID) + 1);
+			jsonGroup = new JSONObject(data);
+            users = jsonGroup.getJSONArray("users");
+			
+			// assign the task to every user in the group we randomly selected
+			for (int j = 0; j < users.length(); j++) {
+               jsonUser = users.getJSONObject(j);
+			   String userId = jsonUser.getString("user_id");
+			   
+			   taskJSON = taskToJSON(userId, CONTIG_ID, DESCRIPTION, DATE);
+		       client.set("Tasks-" + userId + '-' + CONTIG_ID, 0, taskJSON);
+			}
+         } catch (JSONException jsonEx) {
+            System.err.println("Error parsing json (" + data + "): " + jsonEx);
[…]

### 2012-11-21 Assignment6/AssignTaskWorkload.java
Fixed broken code

@@ -8,6 +8,10 @@ import java.sql.ResultSet;
 import java.sql.SQLException;
 import java.sql.Statement;
 
+import org.json.JSONArray;
+import org.json.JSONException;
+import org.json.JSONObject;
+
 /**
  * This one measures assigning a task to an arbitraty group.
  * This is pretty much exacly a 50% R/W Workload.
@@ -48,26 +52,41 @@ public class AssignTaskWorkload extends Workload {
       return new Stats();
    }
 
-   protected Stats executeCouchImpl() {	  
-	  String data = null;
+   protected Stats executeCouchImpl() {
+      String data = null;
       JSONObject jsonUser = null;
-	  JSONObject jsonGroup = null;
+      JSONObject jsonGroup = null;
       JSONArray users = null;
-	  
+
+      JSONObject task = new JSONObject();
+      try {
+         // Since the rest of the db stringifys their numbers, stringify this.
+         task.put("contig_id", "" + CONTIG_ID);
+         task.put("end_date", DATE);
+         task.put("desc", DESCRIPTION);
+      } catch (JSONException ex) {
+         System.err.println("Error creating core task object.");
+         ex.printStackTrace(System.err);
+         return null;
+      }
+
       for (int i = 0; i < TIMES; i++) {
          try {
-            data = (String)client.get("Groups-" + (rand.nextInt() % MAX_GROUP_ID) + 1);
-			jsonGroup = new JSONObject(data);
+            data = (String)client.get("Groups-" + (rand.nextInt(MAX_GROUP_ID) + 1));
+            jsonGroup = new JSONObject(data);
             users = jsonGroup.getJSONArray("users");
-			
-			// assign the task to every user in the group we randomly selected
-			for (int j = 0; j < users.length(); j++) {
-               jsonUser = users.getJSONObject(j);
-			   String userId = jsonUser.getString("user_id");
-			   
-			   taskJSON = taskToJSON(userId, CONTIG_ID, DESCRIPTION, DATE);
-		       client.set("Tasks-" + userId + '-' + CONTIG_ID, 0, taskJSON);
-			}
+
+            // assign the task to every user in the group we randomly selected
+            for (int j = 0; j < users.length(); j++) {
+               String userId = users.getString(j);
+
+               jsonUser = new JSONObject(client.get("Users-" + userId));
+               // Update the user's tasks.
+               jsonUser.getJSONArray("tasks").put(task);
+
+               // Rewrite the user.
[…]

### 2012-11-22 Assignment6/GroupMemWorkload.java
Couchbase stuff inserted

@@ -8,6 +8,9 @@ import java.sql.ResultSet;
 import java.sql.SQLException;
 import java.sql.Statement;
 
+import org.json.JSONArray;
+import org.json.JSONException;
+import org.json.JSONObject;
 /**
  * This one measures fetching a complete user's profile.
  */
@@ -16,9 +19,15 @@ public class GroupMemWorkload extends Workload {
    private static final int TIMES = 100000;
    //private static final int TIMES = 1;
 
+   private static final int MAX_GROUP_ID = 100;
+
+   private static final String DESCRIPTION = "Some task description.";
+   private static final String DATE = "2012-12-12";
+   private static final int CONTIG_ID = 5;
+
    private String[] userIds;
    private String[] groupIds;
-
+   private Random rand;
    /**
     * Fetch the user ids here, and preserve the connection.
     */
@@ -26,6 +35,7 @@ public class GroupMemWorkload extends Workload {
       // Keeping as a string to reduce conversion time.
       userIds = new String[TIMES];
       groupIds = new String[TIMES];
+      rand = new Random(4);
    }
 
    protected void initMySQL() {
@@ -63,7 +73,47 @@ public class GroupMemWorkload extends Workload {
    }
 
    protected Stats executeCouchImpl() {
-      // Nope
-      throw new UnsupportedOperationException();
+      String data = null;
+      JSONObject jsonUser = null;
+      JSONObject jsonGroup = null;
+      JSONArray users = null;
+      int groupId = 0;
+      JSONObject task = new JSONObject();
+      try {
+         // Since the rest of the db stringifys their numbers, stringify this.
+         task.put("contig_id", "" + CONTIG_ID);
+         task.put("end_date", DATE);
+         task.put("desc", DESCRIPTION);
+      } catch (JSONException ex) {
+         System.err.println("Error creating core task object.");
+         ex.printStackTrace(System.err);
+         return null;
+      }
+
+      for (int i = 0; i < TIMES; i++) {
+         try {
+            groupId = rand.nextInt(MAX_GROUP_ID) + 1;
+            data = (String)client.get("Groups-" + (groupId));
+            jsonGroup = new JSONObject(data);
+            users = jsonGroup.getJSONArray("users");
+            // Remove the first user from the group users list
+            if(0 < users.length()) {
+               String userId = users.getString(0);
+               users.remove(0);
+               //Not needed jsonUser = new JSONObject(client.get("Users-" + userId));
+               //Put the modified user list back into the group JSON
[…]

### 2012-11-22 Assignment6/TestMaster.java

@@ -16,21 +16,21 @@ public class TestMaster {
    public static final String DB_USER = "cgat";
    public static final String DB_PASS = "csc560";
 
-   public static final String COUCH_URI = "http://127.0.0.1:8091/pools";
+   public static final String COUCH_URI = "http://ec2-204-236-188-48.us-west-1.compute.amazonaws.com:8091/pools";
 
    public static void main(String[] args) {
       Workload[] workloads = {
-                              //new ContigUploadWorkload(),
+                              new ContigUploadWorkload(),
                               //new ProfileWorkload(),
                               //new AssignTaskWorkload(),
                               //new GroupMemWorkload(),
                               //new PublishWorkload(),
-                              new StupidWorkload(),
+                              //new StupidWorkload(),
                               };
 
       for (Workload workload : workloads) {
          System.out.println("Running Workload [" + workload.getClass().getCanonicalName() + "]...");
-         Stats results = workload.executeMySQL();
+         Stats results = workload.executeCouchImpl();
          System.out.println("Completed Workload [" + workload.getClass().getCanonicalName() + "]...");
          System.out.println(results + "\n");
       }

### 2012-11-22 Assignment6/GroupMemWorkload.java
Couchbase stuff inserted

@@ -100,7 +100,7 @@ public class GroupMemWorkload extends Workload {
             if(0 < users.length()) {
                String userId = users.getString(0);
                users.remove(0);
-               //Not needed jsonUser = new JSONObject(client.get("Users-" + userId));
+               //Not needed at the momemnt jsonUser = new JSONObject(client.get("Users-" + userId));
                //Put the modified user list back into the group JSON
                jsonGroup.getJSONArray("users").put(users);
                //Overwrite the group object

### 2012-11-22 Assignment6/AssignTaskWorkload.java
Removed bad lib, updated connection, checking over couchbase

@@ -20,8 +20,8 @@ import org.json.JSONObject;
  */
 public class AssignTaskWorkload extends Workload {
    // Should be no more than 100000 (the number of users in the db).
-   private static final int TIMES = 1000000;
-   //private static final int TIMES = 10;
+   //private static final int TIMES = 1000000;
+   private static final int TIMES = 10;
 
    private static final int MAX_GROUP_ID = 100;
 
@@ -70,7 +70,7 @@ public class AssignTaskWorkload extends Workload {
          return null;
       }
 
-      for (int i = 0; i < TIMES; i++) {
+      for (int i = 1; i <= TIMES; i++) {
          try {
             data = (String)client.get("Groups-" + (rand.nextInt(MAX_GROUP_ID) + 1));
             jsonGroup = new JSONObject(data);

### 2012-11-22 Assignment6/ContigUploadWorkload.java

@@ -20,8 +20,8 @@ import java.sql.PreparedStatement;
  */
 public class ContigUploadWorkload extends Workload {
    // Should be no more than 9999999999999999999999999999
-   private static final int TIMES = 10;
-   //private static final int TIMES = 5;
+   //private static final int TIMES = 10000;
+   private static final int TIMES = 1;
    private static final int MIN_SEQ_LENGTH = 45000;
    private static final int MAX_ADDITIONAL_LENGTH = 15001;
 

### 2012-11-22 Assignment6/ProfileWorkload.java

@@ -18,8 +18,8 @@ import org.json.JSONObject;
 public class ProfileWorkload extends Workload {
    private static final int MAX_USERS = 100000;
 
-   private static final int TIMES = 1000000;
-   //private static final int TIMES = 1;
+   //private static final int TIMES = 1000000;
+   private static final int TIMES = 1;
 
    private String[] userIds;
 
@@ -109,10 +109,13 @@ public class ProfileWorkload extends Workload {
       // THis shouldn't even be necessary, but I am being neurotic.
       Object throwAway = null;
 
-      for (int i = 0; i < TIMES; i++) {
+      for (int i = 1; i <= TIMES; i++) {
          try {
             // I saw the api just casting, so I will too.
             data = (String)client.get("Users-" + i);
+            //TEST
+            System.out.println(data);
+
             jsonUser = new JSONObject(data);
 
             // Expand groups
@@ -122,6 +125,9 @@ public class ProfileWorkload extends Workload {
 
                // The next step would be trivial, get the groups name form the JSON.
                throwAway = client.get("Groups-" + groupId);
+
+               //TEST
+               System.out.println(throwAway);
             }
 
             // Expand history
@@ -131,6 +137,9 @@ public class ProfileWorkload extends Workload {
                String annotationId = historyJson.getString("anno_id");
 
                throwAway = client.get("Annotations-" + annotationId);
+
+               //TEST
+               System.out.println(throwAway);
             }
 
             // Expand partials
@@ -139,6 +148,9 @@ public class ProfileWorkload extends Workload {
                String annotationId = partials.getString(j);
 
                throwAway = client.get("Annotations-" + annotationId);
+
+               //TEST
+               System.out.println(throwAway);
             }
 
             // Expand tasks
@@ -148,6 +160,9 @@ public class ProfileWorkload extends Workload {
                String contigId = taskJson.getString("contig_id");
 
                throwAway = client.get("Contigs-" + contigId);
+
+               //TEST
+               System.out.println(throwAway);
             }
          } catch (JSONException jsonEx) {
             System.err.println("Error parsing json (" + data + "): " + jsonEx);

### 2012-11-22 Assignment6/StupidWorkload.java

@@ -13,8 +13,8 @@ import java.sql.Statement;
  * Ryan thinks that it is stupid.
  */
 public class StupidWorkload extends Workload {
-   private static final int TIMES = 100000;
-   //private static final int TIMES = 10;
+   //private static final int TIMES = 100000;
+   private static final int TIMES = 10;
 
    private String[] randomStrings;
 
@@ -81,6 +81,13 @@ public class StupidWorkload extends Workload {
       }
 
       return new ReadWriteStats(readAfterWriteTimes, writeAfterReadTimes);
+   }
+
+   protected void initCouch() {
+      super.initCouch();
+
+      // Make an initial set of the key.
+      client.set("ReadAfterWrite-1", 0, "0");
    }
 
    protected Stats executeCouchImpl() {

### 2012-11-22 Assignment6/TestMaster.java

@@ -16,12 +16,13 @@ public class TestMaster {
    public static final String DB_USER = "cgat";
    public static final String DB_PASS = "csc560";
 
-   public static final String COUCH_URI = "http://ec2-204-236-188-48.us-west-1.compute.amazonaws.com:8091/pools";
+   //public static final String COUCH_URI = "http://ec2-204-236-188-48.us-west-1.compute.amazonaws.com:8091/pools";
+   public static final String COUCH_URI = "http://ip-10-168-85-10.us-west-1.compute.internal:8091/pools";
 
    public static void main(String[] args) {
       Workload[] workloads = {
-                              new ContigUploadWorkload(),
-                              //new ProfileWorkload(),
+                              //new ContigUploadWorkload(),
+                              new ProfileWorkload(),
                               //new AssignTaskWorkload(),
                               //new GroupMemWorkload(),
                               //new PublishWorkload(),
@@ -30,7 +31,7 @@ public class TestMaster {
 
       for (Workload workload : workloads) {
          System.out.println("Running Workload [" + workload.getClass().getCanonicalName() + "]...");
-         Stats results = workload.executeCouchImpl();
+         Stats results = workload.executeCouch();
          System.out.println("Completed Workload [" + workload.getClass().getCanonicalName() + "]...");
          System.out.println(results + "\n");
       }

### 2012-11-23 Assignment6/MoveSQLDataToCouch.java
More pageination to moving.

@@ -248,15 +248,33 @@ public class MoveSQLDataToCouch {
                                " WHERE A.GeneId = G.GeneId" +
                                " ORDER BY A.AnnotationId" +
                                " LIMIT ?, " + pageSize;
-      String exonQuery = "SELECT StartPos, EndPos FROM Exons WHERE AnnotationId = ?;";
+      String exonQuery = "SELECT AnnotationId, StartPos, EndPos FROM Exons WHERE AnnotationId BETWEEN ? AND ?";
       PreparedStatement annotationQ = conn.prepareStatement(annotationQuery);
       PreparedStatement exonQ = conn.prepareStatement(exonQuery);
 
       ResultSet rs = null;
 
       // We know we have 1 million annotations.
-      for (int count = 0; count < (1000000 / pageSize); count++) {
-         annotationQ.setInt(1, pageSize * count);
+      for (int page = 0; page < (1000000 / pageSize); page++) {
+         annotationQ.setInt(1, pageSize * page);
+
+         List<ArrayList<Integer>> exonsBatch = new ArrayList<ArrayList<Integer>>();
+         for (int i = 0; i < pageSize; i++) {
+            exonsBatch.add(new ArrayList());
+         }
+
+         // It costs too much to do the exon queries one at a time.
+         // Do a batch at a time.
+         exonQ.setInt(1, page * pageSize + 1);
+         exonQ.setInt(2, page * (pageSize + 1) + 1);
+         ResultSet exonrs = exonQ.executeQuery();
+         while(exonrs.next()) {
+            int annotationId = exonrs.getInt("AnnotationId");
+            int exonStartPos = exonrs.getInt("StartPos");
+            int exonEndPos = exonrs.getInt("EndPos");
+            exonsBatch.get(annotationId).add(exonStartPos);
+            exonsBatch.get(annotationId).add(exonEndPos);
+         }
 
          rs = annotationQ.executeQuery();
          while(rs.next()) {
@@ -274,16 +292,7 @@ public class MoveSQLDataToCouch {
             Date finishedDate = rs.getDate("A.FinishedDate");
             boolean incorrect = rs.getInt("A.Incorrect") == 1;
             boolean expertIncorrect = rs.getInt("A.ExpertIncorrect") == 1;
-            ArrayList<Integer> exonStartEndPairs = new ArrayList<Integer>();
-
-            exonQ.setInt(1,annoId);
-            ResultSet exonrs = exonQ.executeQuery();
-            while(exonrs.next()) {
[…]

### 2012-11-23 Assignment6/MoveSQLDataToCouch.java
fixed connection string

@@ -25,9 +25,9 @@ import net.spy.memcached.internal.OperationFuture;
  * Copy the SQL data into JSON format, then send it to Couchbase.
  */
 public class MoveSQLDataToCouch {
-   private static final String DB_URL = "jdbc:mysql://localhost:3306/test?autoReconnect=true";
-   private static final String DB_USER = "";
-   private static final String DB_PASS = "";
+   private static final String DB_URL = "jdbc:mysql://localhost:3306/cgat?autoReconnect=true";
+   private static final String DB_USER = "cgat";
+   private static final String DB_PASS = "csc560";
 
    private static List<URI> uris;
 

### 2012-11-23 Assignment6/MoveSQLDataToCouch.java
Changed from List to Map

@@ -9,6 +9,7 @@ import java.util.Date;
 import java.util.Random;
 import java.util.ArrayList;
 import java.util.HashMap;
+import java.util.Map;
 
 import java.net.URI;
 import java.util.LinkedList;
@@ -258,9 +259,9 @@ public class MoveSQLDataToCouch {
       for (int page = 0; page < (1000000 / pageSize); page++) {
          annotationQ.setInt(1, pageSize * page);
 
-         List<ArrayList<Integer>> exonsBatch = new ArrayList<ArrayList<Integer>>();
+         Map<Integer, ArrayList<Integer>> exonsBatch = new HashMap<Integer, ArrayList<Integer>>();
          for (int i = 0; i < pageSize; i++) {
-            exonsBatch.add(new ArrayList());
+            exonsBatch.put(new Integer(page * pageSize + i + 1), new ArrayList());
          }
 
          // It costs too much to do the exon queries one at a time.

### 2012-11-23 Assignment6/AssignTaskWorkload.java
Eriq's are verified working

@@ -20,14 +20,15 @@ import org.json.JSONObject;
  */
 public class AssignTaskWorkload extends Workload {
    // Should be no more than 100000 (the number of users in the db).
-   //private static final int TIMES = 1000000;
-   private static final int TIMES = 10;
+   // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
+   //private static final int TIMES = 100000;
+   private static final int TIMES = 1;
 
    private static final int MAX_GROUP_ID = 100;
 
    private static final String DESCRIPTION = "Some task description.";
    private static final String DATE = "2012-12-12";
-   private static final int CONTIG_ID = 5;
+   private static final int CONTIG_ID = 2;
 
    private Random rand;
 
@@ -78,11 +79,11 @@ public class AssignTaskWorkload extends Workload {
 
             // assign the task to every user in the group we randomly selected
             for (int j = 0; j < users.length(); j++) {
-               String userId = users.getString(j);
+               String userId = users.getJSONObject(j).getString("user_id");
 
-               jsonUser = new JSONObject(client.get("Users-" + userId));
+               jsonUser = new JSONObject((String)client.get("Users-" + userId));
                // Update the user's tasks.
-               jsonUser.getJSONArray("tasks").put(task);
+               jsonUser.put("tasks", jsonUser.getJSONArray("tasks").put(task));
 
                // Rewrite the user.
                client.set("Users-" + userId, 0, jsonUser.toString());

### 2012-11-23 Assignment6/ContigUploadWorkload.java

@@ -20,6 +20,7 @@ import java.sql.PreparedStatement;
  */
 public class ContigUploadWorkload extends Workload {
    // Should be no more than 9999999999999999999999999999
+   // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
    //private static final int TIMES = 10000;
    private static final int TIMES = 1;
    private static final int MIN_SEQ_LENGTH = 45000;

### 2012-11-23 Assignment6/GroupMemWorkload.java

@@ -16,6 +16,7 @@ import org.json.JSONObject;
  */
 public class GroupMemWorkload extends Workload {
    // Should be no more than 100000 (the number of users in the db).
+   // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
    private static final int TIMES = 100000;
    //private static final int TIMES = 1;
 

### 2012-11-23 Assignment6/ProfileWorkload.java

@@ -18,6 +18,7 @@ import org.json.JSONObject;
 public class ProfileWorkload extends Workload {
    private static final int MAX_USERS = 100000;
 
+   // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
    //private static final int TIMES = 1000000;
    private static final int TIMES = 1;
 
@@ -120,6 +121,9 @@ public class ProfileWorkload extends Workload {
 
             // Expand groups
             groups = jsonUser.getJSONArray("groups");
+            //TEST
+            System.out.println("Num Groups: " + groups.length());
+
             for (int j = 0; j < groups.length(); j++) {
                String groupId = groups.getString(j);
 
@@ -132,6 +136,9 @@ public class ProfileWorkload extends Workload {
 
             // Expand history
             history = jsonUser.getJSONArray("history");
+            //TEST
+            System.out.println("Num History: " + history.length());
+
             for (int j = 0; j < history.length(); j++) {
                JSONObject historyJson = history.getJSONObject(j);
                String annotationId = historyJson.getString("anno_id");
@@ -144,6 +151,9 @@ public class ProfileWorkload extends Workload {
 
             // Expand partials
             partials = jsonUser.getJSONArray("incomplete_annotations");
+            //TEST
+            System.out.println("Num Partials: " + partials.length());
+
             for (int j = 0; j < partials.length(); j++) {
                String annotationId = partials.getString(j);
 
@@ -155,6 +165,9 @@ public class ProfileWorkload extends Workload {
 
             // Expand tasks
             tasks = jsonUser.getJSONArray("tasks");
+            //TEST
+            System.out.println("Num Tasks: " + tasks.length());
+
             for (int j = 0; j < tasks.length(); j++) {
                JSONObject taskJson = tasks.getJSONObject(j);
                String contigId = taskJson.getString("contig_id");

### 2012-11-23 Assignment6/PublishWorkload.java

@@ -24,8 +24,9 @@ import net.spy.memcached.internal.OperationFuture;
  * This one measures fetching a complete user's profile.
  */
 public class PublishWorkload extends Workload {
-   private static final int TIMES = 100000;
-   //private static final int TIMES = 10;
+   // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
+   //private static final int TIMES = 100000;
+   private static final int TIMES = 10;
 
    private String[] userIds, annotationIds, exp;
    List<String> annotationIdList, userIdList;

### 2012-11-23 Assignment6/StupidWorkload.java

@@ -13,6 +13,7 @@ import java.sql.Statement;
  * Ryan thinks that it is stupid.
  */
 public class StupidWorkload extends Workload {
+   // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
    //private static final int TIMES = 100000;
    private static final int TIMES = 10;
 

### 2012-11-23 Assignment6/TestMaster.java

@@ -22,8 +22,8 @@ public class TestMaster {
    public static void main(String[] args) {
       Workload[] workloads = {
                               //new ContigUploadWorkload(),
-                              new ProfileWorkload(),
-                              //new AssignTaskWorkload(),
+                              //new ProfileWorkload(),
+                              new AssignTaskWorkload(),
                               //new GroupMemWorkload(),
                               //new PublishWorkload(),
                               //new StupidWorkload(),

### 2012-11-24 Assignment6/GroupMemWorkload.java
GroupMemWorkload is ready to run

@@ -16,7 +16,6 @@ import org.json.JSONObject;
  */
 public class GroupMemWorkload extends Workload {
    // Should be no more than 100000 (the number of users in the db).
-   // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
    private static final int TIMES = 100000;
    //private static final int TIMES = 1;
 
@@ -80,8 +79,7 @@ public class GroupMemWorkload extends Workload {
       JSONArray users = null;
       int groupId = 0;
       JSONObject task = new JSONObject();
-      try {
-         // Since the rest of the db stringifys their numbers, stringify this.
+      /**try {
          task.put("contig_id", "" + CONTIG_ID);
          task.put("end_date", DATE);
          task.put("desc", DESCRIPTION);
@@ -89,22 +87,38 @@ public class GroupMemWorkload extends Workload {
          System.err.println("Error creating core task object.");
          ex.printStackTrace(System.err);
          return null;
-      }
+      }**/
 
       for (int i = 0; i < TIMES; i++) {
          try {
+            //Get a random group from Couchbase
             groupId = rand.nextInt(MAX_GROUP_ID) + 1;
-            data = (String)client.get("Groups-" + (groupId));
+            data = (String)client.get("Groups-" + groupId);
+            //Make group into JSON and extract the users from the group
             jsonGroup = new JSONObject(data);
             users = jsonGroup.getJSONArray("users");
-            // Remove the first user from the group users list
             if(0 < users.length()) {
-               String userId = users.getString(0);
+               //Debug System.out.println(users.get(0)); 
+               //Remove the first user from that group
+               JSONObject removedUser = (JSONObject) users.get(0);
                users.remove(0);
-               //Not needed at the momemnt jsonUser = new JSONObject(client.get("Users-" + userId));
                //Put the modified user list back into the group JSON
-               jsonGroup.getJSONArray("users").put(users);
-               //Overwrite the group object
+               jsonGroup.remove("users");
+               jsonGroup.put("users", users);
+               //Overwrite the group in Couchbase
+               client.set("Groups-" + groupId, 0, jsonGroup.toString());
+               //Get the same group back from Couchbase
+               data = (String)client.get("Groups-" + groupId);
+               //Make group into JSON and extract the users from the group
+               jsonGroup = new JSONObject(data);
+               users = jsonGroup.getJSONArray("users");
+               //Put the user back into the group
+               users.put(0, removedUser);
[…]

### 2012-11-24 Assignment6/PublishWorkload.java
Committing changes to PublishWorkload. This should be close to functional though it still needs testing.

@@ -13,6 +13,7 @@ import java.sql.Statement;
 import java.text.SimpleDateFormat;
 import java.sql.Date;
 
+import org.json.JSONArray;
 import org.json.JSONException;
 import org.json.JSONObject;
 
@@ -100,7 +101,6 @@ public class PublishWorkload extends Workload {
 
    protected void initCouch() {
       super.initCouch();
-      /*
 
       Random rand = new Random(4);
 
@@ -111,36 +111,134 @@ public class PublishWorkload extends Workload {
          annotationIdList.add(String.format("Annotations-%d", rand.nextInt(1000)));
          userIdList.add(String.format("Users-%d", rand.nextInt(100)));
       }
-      */
    }
 
    protected Stats executeCouchImpl() {
-      /*
       JSONObject userJSON = null, annotationJSON = null, contigJSON = null;
 
       for (int iter_ndx = 0; iter_ndx < TIMES; iter_ndx++) {
-         Object user = client.get(userIdList.get(iter_ndx));
-         Object annotation = client.get(annotationIdList.get(iter_ndx));
+         String user_id = userIdList.get(iter_ndx);
+         String anno_id = annotationIdList.get(iter_ndx);
+
+         /*
+          * Retrieve the user and annotation documents. Everything revolves
+          * around these (to some extent)
+          */
+         Object user = client.get(user_id);
+         Object annotation = client.get(anno_id);
 
          try {
             userJSON = new JSONObject(user.toString());
             annotationJSON = new JSONObject(annotation.toString());
 
-            int contigId = annotationJSON.getInt("contig_id");
-            Object contig = client.get("Contigs-" + String.valueOf(contigId));
+            /*###############################################################*/
+            //CONTIG DIFFICULTY
+
+            /*
+             * Get the contig document based on the contig_id in the annotation
+             * document
+             */
+            String contigId = annotationJSON.getString("contig_id");
+            Object contig = client.get("Contigs-" + contigId);
             contigJSON = new JSONObject(contig.toString());
+
+            //Get the difficulty from the contig document
+            Integer contigDifficulty = Integer.parseInt(contigJSON.getString("difficulty"));
+
+            /*###############################################################*/
+            //USER META
+
+            /*
+             * Get the meta object from the user document. This contains the
+             * user's total experience
+             */
+            JSONObject userMeta = new JSONObject(userJSON.getJSONObject("meta"));
+            Integer userXP = Integer.parseInt(userMeta.getString("exp"));
+
+            /*
[…]

### 2012-11-24 Assignment6/PublishWorkload.java
Fixed errors in PublishWorkload so that it is now verified to run. The number of data points to run its test on is not updated however. I'll leave that to Eriq

@@ -46,6 +46,11 @@ public class PublishWorkload extends Workload {
    }
 
    private static String currDate() {
+      SimpleDateFormat formatter = new SimpleDateFormat("yyyy-MM-dd");
+      return formatter.format(new Date(System.currentTimeMillis()));
+   }
+
+   private static String currDateTime() {
       SimpleDateFormat formatter = new SimpleDateFormat("yyyy-MM-dd HH:mm:ss");
       return formatter.format(new Date(System.currentTimeMillis()));
    }
@@ -90,7 +95,7 @@ public class PublishWorkload extends Workload {
                                               userIds[i]));
 
             Util.doUpdate(conn, String.format(annotationUpdateQuery,
-                                              currDate(),
+                                              this.currDateTime(),
                                               annotationExp,
                                               annotationIds[i]));
          }
@@ -134,6 +139,7 @@ public class PublishWorkload extends Workload {
 
             /*###############################################################*/
             //CONTIG DIFFICULTY
+            //System.out.println("getting contig difficulty");
 
             /*
              * Get the contig document based on the contig_id in the annotation
@@ -143,17 +149,20 @@ public class PublishWorkload extends Workload {
             Object contig = client.get("Contigs-" + contigId);
             contigJSON = new JSONObject(contig.toString());
 
+            JSONObject contigMeta = contigJSON.getJSONObject("meta");
+
             //Get the difficulty from the contig document
-            Integer contigDifficulty = Integer.parseInt(contigJSON.getString("difficulty"));
+            Integer contigDifficulty = Integer.parseInt(contigMeta.getString("difficulty"));
 
             /*###############################################################*/
             //USER META
+            //System.out.println("increasing user experience");
 
             /*
              * Get the meta object from the user document. This contains the
              * user's total experience
              */
-            JSONObject userMeta = new JSONObject(userJSON.getJSONObject("meta"));
+            JSONObject userMeta = userJSON.getJSONObject("meta");
             Integer userXP = Integer.parseInt(userMeta.getString("exp"));
 
             /*
@@ -171,6 +180,7 @@ public class PublishWorkload extends Workload {
 
             /*###############################################################*/
             //USER HISTORY
+            //System.out.println("appending annotation to user history");
 
             /*
              * Add annotation to user history
@@ -196,6 +206,7 @@ public class PublishWorkload extends Workload {
 
             /*###############################################################*/
[…]

### 2012-11-26 Assignment6/AssignTaskWorkload.java
Publish workload had a bug where it generated an invalid id

@@ -21,8 +21,8 @@ import org.json.JSONObject;
 public class AssignTaskWorkload extends Workload {
    // Should be no more than 100000 (the number of users in the db).
    // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
-   //private static final int TIMES = 100000;
-   private static final int TIMES = 1;
+   private static final int TIMES = 100000;
+   //private static final int TIMES = 1;
 
    private static final int MAX_GROUP_ID = 100;
 

### 2012-11-26 Assignment6/ContigUploadWorkload.java

@@ -21,8 +21,8 @@ import java.sql.PreparedStatement;
 public class ContigUploadWorkload extends Workload {
    // Should be no more than 9999999999999999999999999999
    // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
-   //private static final int TIMES = 10000;
-   private static final int TIMES = 1;
+   private static final int TIMES = 10000;
+   //private static final int TIMES = 1;
    private static final int MIN_SEQ_LENGTH = 45000;
    private static final int MAX_ADDITIONAL_LENGTH = 15001;
 

### 2012-11-26 Assignment6/ProfileWorkload.java

@@ -19,8 +19,8 @@ public class ProfileWorkload extends Workload {
    private static final int MAX_USERS = 100000;
 
    // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
-   //private static final int TIMES = 1000000;
-   private static final int TIMES = 1;
+   private static final int TIMES = 100000;
+   //private static final int TIMES = 1;
 
    private String[] userIds;
 
@@ -114,68 +114,41 @@ public class ProfileWorkload extends Workload {
          try {
             // I saw the api just casting, so I will too.
             data = (String)client.get("Users-" + i);
-            //TEST
-            System.out.println(data);
-
             jsonUser = new JSONObject(data);
 
             // Expand groups
             groups = jsonUser.getJSONArray("groups");
-            //TEST
-            System.out.println("Num Groups: " + groups.length());
-
             for (int j = 0; j < groups.length(); j++) {
                String groupId = groups.getString(j);
 
                // The next step would be trivial, get the groups name form the JSON.
                throwAway = client.get("Groups-" + groupId);
-
-               //TEST
-               System.out.println(throwAway);
             }
 
             // Expand history
             history = jsonUser.getJSONArray("history");
-            //TEST
-            System.out.println("Num History: " + history.length());
-
             for (int j = 0; j < history.length(); j++) {
                JSONObject historyJson = history.getJSONObject(j);
                String annotationId = historyJson.getString("anno_id");
 
                throwAway = client.get("Annotations-" + annotationId);
-
-               //TEST
-               System.out.println(throwAway);
             }
 
             // Expand partials
             partials = jsonUser.getJSONArray("incomplete_annotations");
-            //TEST
-            System.out.println("Num Partials: " + partials.length());
-
             for (int j = 0; j < partials.length(); j++) {
                String annotationId = partials.getString(j);
 
                throwAway = client.get("Annotations-" + annotationId);
-
-               //TEST
-               System.out.println(throwAway);
             }
 
             // Expand tasks
             tasks = jsonUser.getJSONArray("tasks");
-            //TEST
-            System.out.println("Num Tasks: " + tasks.length());
-
             for (int j = 0; j < tasks.length(); j++) {
                JSONObject taskJson = tasks.getJSONObject(j);
                String contigId = taskJson.getString("contig_id");
 
                throwAway = client.get("Contigs-" + contigId);
-
-               //TEST
-               System.out.println(throwAway);
             }
[…]

### 2012-11-26 Assignment6/PublishWorkload.java

@@ -26,8 +26,8 @@ import net.spy.memcached.internal.OperationFuture;
  */
 public class PublishWorkload extends Workload {
    // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
-   //private static final int TIMES = 100000;
-   private static final int TIMES = 10;
+   private static final int TIMES = 100000;
+   //private static final int TIMES = 10;
 
    private String[] userIds, annotationIds, exp;
    List<String> annotationIdList, userIdList;
@@ -114,8 +114,8 @@ public class PublishWorkload extends Workload {
       userIdList = new ArrayList<String>(TIMES);
 
       for (int repeat = 0; repeat < TIMES; repeat++) {
-         annotationIdList.add(String.format("Annotations-%d", rand.nextInt(1000)));
-         userIdList.add(String.format("Users-%d", rand.nextInt(100)));
+         annotationIdList.add(String.format("Annotations-%d", rand.nextInt(1000) + 1));
+         userIdList.add(String.format("Users-%d", rand.nextInt(100) + 1));
       }
    }
 
@@ -249,7 +249,7 @@ public class PublishWorkload extends Workload {
          }
          catch (Exception ex) {
             System.err.println("Error in PublishWorkload Couchbase code: " + ex);
-            System.exit(1);
+            ex.printStackTrace(System.err);
          }
 
       }

### 2012-11-26 Assignment6/StupidWorkload.java

@@ -14,8 +14,8 @@ import java.sql.Statement;
  */
 public class StupidWorkload extends Workload {
    // DON'T TOUCH THIS NUMBER. I need to keep it consistent for testing. -Eriq
-   //private static final int TIMES = 100000;
-   private static final int TIMES = 10;
+   private static final int TIMES = 100000;
+   //private static final int TIMES = 10;
 
    private String[] randomStrings;
 

### 2012-11-26 Assignment6/TestMaster.java

@@ -23,10 +23,10 @@ public class TestMaster {
       Workload[] workloads = {
                               //new ContigUploadWorkload(),
                               //new ProfileWorkload(),
-                              new AssignTaskWorkload(),
-                              //new GroupMemWorkload(),
-                              //new PublishWorkload(),
                               //new StupidWorkload(),
+                              //new GroupMemWorkload(),
+                              new PublishWorkload(),
+                              new AssignTaskWorkload(),
                               };
 
       for (Workload workload : workloads) {

### 2012-11-26 Assignment6/paper/paper.tex

@@
 and time-delay between writes and reads of the same data. The rest of the paper is broken down as follows.
-Section 2 describes in depth how we designed the experiment. Section 3 discusses our implementation. Section 4 goes
-into the results of each of our experiments and some analysis on the data we obtained. Lastly, section 5 presents 
+Section 2 describes in depth how we designed the experiment. Section 3 discusses the workloads we ran on each system. Section 4 describes our implementation. Section 5 goes
+into the results of each of our experiments and some analysis on the data we obtained. Lastly, section 6 presents 
 our conclusions.
@@
 \section{Experiment Design}
+For our experiments we are using a total of five EC2 instances from the Amazon Web Services cloud.
+For both of our systems four servers were used for the database while the last was used for running our workload harness.
+The MySQL system is setup with one master server for writes and three slaves servers that replicate all of the masters data.
+Unlike the Couchbase system where we had four servers setup for reads and writes of the keys they were incharge of.
+
+Our major driving force of this experiment was to see how Couchbase compared to MySQL for running a web application.
+To test which is better we created a workload harness.
+The main task of the workload harness was to define how many times a workload would run.
+In our case workloads define a certain task.
+For example, getting of the data necessary to display a users profile.
+We leave describing each workload we ran to the next section.
+Here we will focus on the design of the systems.
+
+As each workload is run the workload harness records important stats.
+Some of the stats include how long the workload took to run, the min and max times of an individual query or insertion, and the standard deviations of individual runs.
+
+\subsection{Data}
+The data we are storing comes from Ryan Verdon's masters thesis.
+It is the backend for a community genome annotation database.
+So it has data like genome sequences, genes, and species information.
+The representation of the data is vastly different in the two systems.
+There exists four main pieces of data that are cruicial to either representation.
+They are Contigs, Annotations, Users, and Groups.
+Contigs are what a biologist annotates which basically means looks at to find a gene.
+The main portion of a Contig is a DNA sequence.
+Annotations are the record or formal representation of a gene that has been found in a Contig.
+An annotation has several parts.
+The main ones are coordinates in the Contig of where individual exons are located and a ton of meta information about the annotation like who uploaded it and when it was done, etc.
+Users is an important part of the data it is one of the central pillars that connects the data.
+Users upload Contigs, create Annotations, and join groups.
+Groups are an important part of the usefulness of the system.
+Users of likemindness can join the same group to receive important information that is important to them.
+Below we discuss the different ways we store the data in SQL versus JSON.
+
+\subsubsection{SQL}
 TODO
+ER DIAGRAM
+
+\subsubsection{JSON}
+TODO
+PIC OF EACH JSON OBJECT
 
 \section{Workloads}
-The following workloads were applied to the both the MySQL and Couchbase clusers.
+The following workloads were ran on both the MySQL and Couchbase clusters.
 \subsection{Assigning a Task}
-TODO
+This workload comes from a use-case in the web application. The main idea is that a new Contig has been uploaded and some uses that care about this Contig need to be notified.
+
+Subsubsections for each system?
 \subsection{Profiling}
-TODO
-\subsection{Read and Writing Performance}
 TODO
@@
 In Couchbase, this requires retrieving the key for the desired group and removing the user from the members field.
+\subsection{Uploading Contigs}
+
+\subsection{Read and Writing Performance}
+TODO
 
 \section{Implementation}
-TODO
+Our experiment was implemented completly in Java. To do the MySQL portion of the workload we used
+the MySQL JDBC driver. For the Couchbase portion we used their Java library from their website.
 

### 2012-11-27 Assignment6/paper/paper.tex
Added workload descriptions and SQL diagram

@@
 Contigs are what a biologist annotates which basically means looks at to find a gene.
-The main portion of a Contig is a DNA sequence.
-Annotations are the record or formal representation of a gene that has been found in a Contig.
+The main portion of a contig is a DNA sequence.
+Annotations are the record or formal representation of a gene that has been found in a contig.
 An annotation has several parts.
-The main ones are coordinates in the Contig of where individual exons are located and a ton of meta information about the annotation like who uploaded it and when it was done, etc.
+The main ones are coordinates in the contig of where individual exons are located and a ton of meta information about the annotation like who uploaded it and when it was done, etc.
 Users is an important part of the data it is one of the central pillars that connects the data.
-Users upload Contigs, create Annotations, and join groups.
+Users upload contigs, create annotations, and join groups.
 Groups are an important part of the usefulness of the system.
@@
 \subsubsection{SQL}
-TODO
-ER DIAGRAM
+The relational schema for the MySQL cluster is shown in Figure~\ref{fig:SQLDiagram}.
 
 \subsubsection{JSON}
-TODO
-PIC OF EACH JSON OBJECT
+The JSON objects for the Couchbase cluster are shown in Figure~\ref{fig:JSON}.
 
+\clearpage
+\begin{figure}%
+{%
+\setlength{\fboxsep}{0pt}%
+\setlength{\fboxrule}{1pt}%
+\fbox{\includegraphics[width=180mm]{SQLDiagram.png}}%
+}%
+\caption{The SQL Schema for CGAT.}
+\label{fig:SQLDiagram}
+\end{figure}%
+
+\clearpage
+\begin{figure}%
+{%
+\setlength{\fboxsep}{0pt}%
+\setlength{\fboxrule}{1pt}%
+\fbox{\includegraphics[width=180mm]{SQLDiagram.png}}%
+}%
+\caption{The JSON object definitions for CGAT.}
+\label{fig:JSON}
+\end{figure}%
+
+\clearpage
 \section{Workloads}
@@
 \subsection{Assigning a Task}
-This workload comes from a use-case in the web application. The main idea is that a new Contig has been uploaded and some uses that care about this Contig need to be notified.
+This workload comes from a use-case in the web application. The main idea is that a new contig has been uploaded and some users that care about this contig need to be notified.
 
-Subsubsections for each system?
 \subsection{Profiling}
-TODO
+The profiling workload simulates what happens when all of the information about a user is gathered.
+This includes: all groups they are a member of, all full and partial annotations they have submitted, and all tasks they have assigned.
+This is a read-heavy workload.
+
 \subsection{Publishing}
-TODO
+This workload involves updating a user's experience because of a completed annotation.
+The corresponding annontation record must also be set to complete, have its finished date updated, and have its gained experience set.
+This is a write-heavy workload.
+
 \subsection{Adding and Removing Group Membership}
@@
 In Couchbase, this requires retrieving the key for the desired group and removing the user from the members field.
+This is a mixed workload involving reading and writing.
+
 \subsection{Uploading Contigs}
+This workload represents a contig being inserted into the database. All of the attributes of a contig are set, including the user who uploaded it and its sequence. This is a write-heavy workload.
 

### 2012-11-27 ui/js/annotation.js
Working on UI mainly for annotation page

@@ -388,18 +388,25 @@ function placeExons() {
 }
 
 // markPosition is meant to mark the first nucleotide in each line.
-function createNucleotideDiv(nucleotide, position, markPosition) {
+function createNucleotideDiv(nucleotide, position, markPosition, codon, codonWindow) {
    var positionMarker = '';
    if (markPosition) {
       positionMarker = '<div class="line-position-marker">' + position + '</div>';
    }
 
-   return "<div class='nucleotide nucleotide-" + nucleotide + "'" +
+   var codonElement = '';
+   if (codon) {
+      codonElement = '<div class="codon codon-' + codon.replace(/\W/, '') + ' codon-window-' +
+                     codonWindow + '">' + codon + '</div>';
+   }
+
+   return "<div class='nucleotide'" +
           " data-position='" + position + "'" +
           " onClick='nucleotideClicked(" + position + ");'" +
           " id='" + "nucleotide-" + position + "'>" +
-          nucleotide +
+          "<div class='nucleotide-base nucleotide-" + nucleotide + "'>" + nucleotide + "</div>" +
           positionMarker +
+          codonElement +
           "</div>";
 }
 
@@ -430,12 +437,31 @@ function updateDnaSelection(leftEdge) {
       sequence = window.cgat.dna.substring(start, start + window.cgat.nucleoditesPerWindow);
    }
 
+   // Get the translated codons.
+   // Remember, because of windows, not every nucleotide will be included in a codon.
+   // So, start from the first multiple of three.
+   var codonStartOffset = (Math.ceil(start / 3) * 3) - start;
+   var windowedCodons = [];
+   windowedCodons.push(codonTranslation(sequence, codonStartOffset));
+   windowedCodons.push(codonTranslation(sequence, codonStartOffset + 1));
+   windowedCodons.push(codonTranslation(sequence, codonStartOffset + 2));
+
+   var codonCounter = 0;
    var sequenceDivs = [];
    var markPosition = false;
    for (var i = 0; i < sequence.length; i++) {
       // TODO(eriq): 50 is magic for number of nucleotides per line. Put it in constants.
       markPosition = (i % 50) === 0;
-      sequenceDivs.push(createNucleotideDiv(sequence[i], start + i, markPosition));
+
+      // If there is a codon associated with the nucleodite, pair it with the nucleotide.
+      if (i >= codonStartOffset) {
+         var currentCodon = windowedCodons[codonCounter % 3][Math.floor(codonCounter / 3)];
[…]

### 2012-11-27 ui/js/bio.js

@@ -1,0 +1,66 @@
+"use strict";
+
+/**
+ * This should be the place to put core bio/chemistry utilities.
+ * Ideally, this file should be able to be riped out and used as
+ * a library for another project.
+ */
+
+window.bioConstants = {};
+window.bioConstants.codons = {'TTT': 'PHE', 'TCT': 'SER', 'TAT': 'TYR', 'TGT': 'CYS', 'TTC': 'PHE',
+                              'TCC': 'SER', 'TAC': 'TYR', 'TGC': 'CYS', 'TTA': 'LEU', 'TCA': 'SER',
+                              'TAA': 'STOP', 'TGA': 'STOP', 'TTG': 'LEU', 'TCG': 'SER', 'TAG': 'STOP',
+                              'TGG': 'TRP', 'CTT': 'LEU', 'CCT': 'PRO', 'CAT': 'HIS', 'CGT': 'ARG',
+                              'CTC': 'LEU', 'CCC': 'PRO', 'CAC': 'HIS', 'CGC': 'ARG', 'CTA': 'LEU',
+                              'CCA': 'PRO', 'CAA': 'GLN', 'CGA': 'ARG', 'CTG': 'LEU', 'CCG': 'PRO',
+                              'CAG': 'GLN', 'CGG': 'ARG', 'ATT': 'ILE', 'ACT': 'THR', 'AAT': 'ASN',
+                              'AGT': 'SER', 'ATC': 'ILE', 'ACC': 'THR', 'AAC': 'ASN', 'AGC': 'SER',
+                              'ATA': 'ILE', 'ACA': 'THR', 'AAA': 'LYS', 'AGA': 'ARG', 'ATG': 'MET*',
+                              'ACG': 'THR', 'AAG': 'LYS', 'AGG': 'ARG', 'GTT': 'VAL', 'GCT': 'ALA',
+                              'GAT': 'ASP', 'GGT': 'GLY', 'GTC': 'VAL', 'GCC': 'ALA', 'GAC': 'ASP',
+                              'GGC': 'GLY', 'GTA': 'VAL', 'GCA': 'ALA', 'GAA': 'GLU', 'GGA': 'GLY',
+                              'GTG': 'VAL', 'GCG': 'ALA', 'GAG': 'GLU', 'GGG': 'GLY'};
+
+// Get the reverse complement of a sequence
+// Will return undefined if the sequence is not valid (not upcase ATCG).
+function reverseComplement(sequence) {
+   var rc = [];
+
[…]

### 2012-11-27 ui/js/group.js

@@ -54,10 +54,10 @@ document.addEventListener('DOMContentLoaded', function() {
             }
 
             if (inGroup) {
-               $('#join-leave-button').text('Leave ' + window.cgat.groupName);
+               $('#join-leave-button').text('Click To Leave ' + window.cgat.groupName);
                $('#join-leave-button').click(leaveGroup);
             } else {
-               $('#join-leave-button').text('Join ' + window.cgat.groupName);
+               $('#join-leave-button').text('Click To Join ' + window.cgat.groupName);
                $('#join-leave-button').click(joinGroup);
             }
          }
@@ -85,7 +85,7 @@ function joinGroup() {
       },
       success: function(data, textStatus, jqXHR) {
          $('#join-leave-button').unbind('click');
-         $('#join-leave-button').text('Leave ' + window.cgat.groupName);
+         $('#join-leave-button').text('Click To Leave ' + window.cgat.groupName);
          $('#join-leave-button').click(leaveGroup);
          window.cgat.disableGroupButton = false;
       }
@@ -111,7 +111,7 @@ function leaveGroup() {
       },
       success: function(data, textStatus, jqXHR) {
          $('#join-leave-button').unbind('click');
-         $('#join-leave-button').text('Join ' + window.cgat.groupName);
+         $('#join-leave-button').text('Click To Join ' + window.cgat.groupName);
          $('#join-leave-button').click(joinGroup);
          window.cgat.disableGroupButton = false;
       }

### 2012-11-27 ui/js/profile.js

@@ -38,7 +38,7 @@ document.addEventListener('DOMContentLoaded', function () {
 
          // Rank info
          $('#stats-level').text('Level ' + data.meta.level);
-         $('#stats-exp').text(data.meta.exp + 'XP');
+         $('#stats-exp').text(data.meta.exp + ' Experience');
 
          // Notifications
          if (window.cgatSession && window.cgatSession.userId === data['_id']['$id']) {

### 2012-11-27 ui/js/script.js

@@ -1,32 +1,5 @@
 "use strict";
 
-// Get the reverse complement of a sequence
-// Will return undefined if the sequence is not valid (not upcase ATCG).
-function reverseComplement(sequence) {
-   var rc = [];
-
-   for (var i = sequence.length - 1; i >= 0; i--) {
-      switch(sequence[i]) {
-         case 'A':
-            rc.push('T');
-            break;
-         case 'T':
-            rc.push('A');
-            break;
-         case 'C':
-            rc.push('G');
-            break;
-         case 'G':
-            rc.push('C');
-            break;
-         default:
-            return undefined;
-      }
-   }
-
-   return rc.join('');
-}
-
 function collapse(buttonId, collapseAreaId) {
    $('#' + buttonId).addClass('collapse-on');
    $('#' + collapseAreaId).addClass('collapse');

### 2012-11-27 Assignment6/paper/paper.tex
Working on UI mainly for annotation page
Adding paper stuff, just need evaluation/conclusion

@@
 Our team decided to compare a MySQL cluster versus a Couchbase cluster for running a
-web application. We are using an eight table SQL schema from Ryan Verdon's Thesis. To evaluate
-the performance we autogenerated around 400MB of data and inserted the same data into each cluster.
-Obviously using a different data model for Couchbase to accomodate for the fact it is a NoSQL key-value store
+web application. We are using an eight table SQL schema from Ryan Verdon's Thesis plus an addition testing table. To evaluate
+the performance we auto-generated around 400MB of data and inserted the same data into each cluster.
+Obviously using a different data model for Couchbase to accommodate for the fact it is a NoSQL key-value store
 instead of a relational database. We then created a set of six experiments to run. Five came from common 
@@
 The MySQL system is setup with one master server for writes and three slaves servers that replicate all of the masters data.
-Unlike the Couchbase system where we had four servers setup for reads and writes of the keys they were incharge of.
+Unlike the Couchbase system where each server handled a subset of the data.
 
@@
 To test which is better we created a workload harness.
-The main task of the workload harness was to define how many times a workload would run.
+The main tasks of the workload harness was to define how many times a workload would run and which workloads to run.
 In our case workloads define a certain task.
 For example, getting of the data necessary to display a users profile.
-We leave describing each workload we ran to the next section.
-Here we will focus on the design of the systems.
+In our system a workload knows how to do its job in both MySQL and the Couchbase API.
+We leave describing each type workload we ran to the next section.
 
@@
 It is the backend for a community genome annotation database.
-So it has data like genome sequences, genes, and species information.
+It contains data like genome sequences, genes, and species information.
 The representation of the data is vastly different in the two systems.
-There exists four main pieces of data that are cruicial to either representation.
-They are Contigs, Annotations, Users, and Groups.
-Contigs are what a biologist annotates which basically means looks at to find a gene.
+
+But there exists four main pieces of data that are crucial to either representation.
+They are contigs, annotations, users, and groups.
+Contigs are what a biologist annotates to find genes.
 The main portion of a contig is a DNA sequence.
-Annotations are the record or formal representation of a gene that has been found in a contig.
-An annotation has several parts.
-The main ones are coordinates in the contig of where individual exons are located and a ton of meta information about the annotation like who uploaded it and when it was done, etc.
-Users is an important part of the data it is one of the central pillars that connects the data.
+Annotations are the record or formal representation of a gene that has been found in a specific contig.
+An annotation has several parts including coordinates in the contig which specify individual exons and meta information about the annotation like who uploaded it and when it was done, etc.
+
+Groups are another important part of the usefulness of the system.
+Users of like minds can join the same group to receive important information.
+Users is a major part of the data and is one of the central pillars that connects the data.
 Users upload contigs, create annotations, and join groups.
-Groups are an important part of the usefulness of the system.
-Users of likemindness can join the same group to receive important information that is important to them.
-Below we discuss the different ways we store the data in SQL versus JSON.
+Below we show the different ways we store the data in SQL and JSON.
 
 \subsubsection{SQL}
-The relational schema for the MySQL cluster is shown in Figure~\ref{fig:SQLDiagram}.
+In SQL the data is represented as nine tables. A diagram of the tables with links to represent foreign key constraints is shown in Figure 1.
 
 \subsubsection{JSON}
-The JSON objects for the Couchbase cluster are shown in Figure~\ref{fig:JSON}.
+For Couchbase we used four JSON objects which are shown in Figures 2-5.
+We left out the JSON object for the test table.
+It just consisted of two fields (Id and Name).
+All of the information from the SQL tables is represented in these JSON objects.
+We take advantage of nested objects to reduce as much subsequent requerying of the database as possible.
 
-\clearpage
 \begin{figure}%
-{%
-\setlength{\fboxsep}{0pt}%
-\setlength{\fboxrule}{1pt}%
-\fbox{\includegraphics[width=180mm]{SQLDiagram.png}}%
-}%
-\caption{The SQL Schema for CGAT.}
+	\includegraphics[width=5.5in]{SQLDiagram.png}%
+	\caption{The SQL diagram for CGAT.}
 	\label{fig:SQLDiagram}
@@
 \clearpage
+
 \begin{figure}%
-{%
-\setlength{\fboxsep}{0pt}%
-\setlength{\fboxrule}{1pt}%
-\fbox{\includegraphics[width=180mm]{SQLDiagram.png}}%
-}%
-\caption{The JSON object definitions for CGAT.}
-\label{fig:JSON}
+	\includegraphics[height=90mm,width=100mm]{user.png}
+	\caption{The JSON for a user.}
+	\label{fig:JSON-user}
+\end{figure}%
+
+\begin{figure}%
+	\includegraphics[height=70mm]{contig.png}
+	\caption{The JSON for a contig.}
+	\label{fig:JSON-contig}
+\end{figure}%
+
+\begin{figure}%
+	\includegraphics[height=100mm]{annotation.png}
+	\caption{The JSON for an annotation.}
+	\label{fig:JSON-annotation}
+\end{figure}%
+
+\begin{figure}%
+	\includegraphics[height=20mm]{group.png}
+	\caption{The JSON for a group.}
+	\label{fig:JSON-group}
 \end{figure}%
@@
 \clearpage
+
 \section{Workloads}
 The following workloads were ran on both the MySQL and Couchbase clusters.
+
 \subsection{Assigning a Task}
@@
 This workload involves updating a user's experience because of a completed annotation.
-The corresponding annontation record must also be set to complete, have its finished date updated, and have its gained experience set.
+The corresponding annotation record must also be set to complete, have its finished date updated, and have its gained experience set.
 This is a write-heavy workload.
@@
 \subsection{Read and Writing Performance}
-TODO
+This is a test workload to examine the performance of writing and reading from the same object in succession. We are interested in the differences in performance of the two systems.
 
 \section{Implementation}
-Our experiment was implemented completly in Java. To do the MySQL portion of the workload we used
-the MySQL JDBC driver. For the Couchbase portion we used their Java library from their website.
+Our experiment was implemented completely in Java. To do the MySQL portion of the workload we used
+the MySQL JDBC driver. For the Couchbase portion we used the Couchbase Java library from their website.
 
 \section{Evaluation}
-TODO
+In this section we discuss the results from each workload in turn.
+
+%section on each workload and the results on each system (could be a table tbh and not a graph)
+%probably some weak analysis on why we expected the results
+%need to ask eriq the # of times each workload was  run, might be in an email?
+
+\subsection{Assigning a Task}
+
+\subsection{Profiling}
+
+\subsection{Publishing}
+
+\subsection{Adding and Removing Group Membership}
+
+\subsection{Uploading Contigs}
+
+\subsection{Read and Writing Performance}
 

### 2012-11-27 Assignment6/paper/paper.tex
Converted images from png to eps so that they have bounding boxes and latex isn't upset.
modified paper.tex to be way better than before but it still sucks because I
hit a wall of wtf sleepy

@@
 
-\section{Introduction}
-Our team decided to compare a MySQL cluster versus a Couchbase cluster for running a
-web application. We are using an eight table SQL schema from Ryan Verdon's Thesis plus an addition testing table. To evaluate
-the performance we auto-generated around 400MB of data and inserted the same data into each cluster.
-Obviously using a different data model for Couchbase to accommodate for the fact it is a NoSQL key-value store
-instead of a relational database. We then created a set of six experiments to run. Five came from common 
-use cases found in the web application. While the other was designed to test the performance on each cluster
-of repeatedly writing and reading from the same object. Our major metrics are write performance, read performance,
-and time-delay between writes and reads of the same data. The rest of the paper is broken down as follows.
-Section 2 describes in depth how we designed the experiment. Section 3 discusses the workloads we ran on each system. Section 4 describes our implementation. Section 5 goes
-into the results of each of our experiments and some analysis on the data we obtained. Lastly, section 6 presents 
-our conclusions.
+\section{Introduction}\label{sec:introduction}
+Gene annotation is the process of associating metadata about a gene with the
+contig on which the gene resides. This metadata is necessary for conducting
+genomic research projects and analysis involving the contig's genomic sequence.
+Currently, there is no specifically gene annotation software. Genome browsers
+all maintain lots of genomic data and are used to be able to annotate genes and
+contigs, but do not provide a user-friendly interface. While genome browsers
+will likely never be replaced (UCSC genome browser, etc. are well
+established), it would be desirable to have a software system that better
+accomodates the visual and informational needs of gene annotation.
 
-\section{Experiment Design}
-For our experiments we are using a total of five EC2 instances from the Amazon Web Services cloud.
-For both of our systems four servers were used for the database while the last was used for running our workload harness.
-The MySQL system is setup with one master server for writes and three slaves servers that replicate all of the masters data.
-Unlike the Couchbase system where each server handled a subset of the data.
+\textit{CGAT} is a web-based gene annotation application designed with
+usability, simplicity, and efficiency in mind. It is not a feature and data
+rich genome browser like the \textit{UCSC genome brower}. Nor is \textit{CGAT}
+designed to be a replacement for other genome browsers in any aspect other than
+gene annotation. Other gene browsers attempt to accommodate many needs of the
+biology community and so offer many features, and a cluttered, chaotic
+interface. Gene annotation, while not a trivial task, can be well-serviced by a
+simple data model. Ideally, by focusing on a simple data model, \textit{CGAT}
+is able to provide a clean, streamlined experience. \textit{CGAT} aims to
+provide users with a way to view gene annotations without extra, unnecessary
+information obscuring their view while also embodying a wikipedia-like emphasis
+on collaboration and openness.
 
-Our major driving force of this experiment was to see how Couchbase compared to MySQL for running a web application.
-To test which is better we created a workload harness.
-The main tasks of the workload harness was to define how many times a workload would run and which workloads to run.
-In our case workloads define a certain task.
-For example, getting of the data necessary to display a users profile.
-In our system a workload knows how to do its job in both MySQL and the Couchbase API.
-We leave describing each type workload we ran to the next section.
+Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
+Obispo, offers a bioinformatics course that covers several aspects of gene
+annotation. This project is spearheaded by Ryan Verdon to accommodate Dr.
+Goodman's needs and ideas for ideal gene annotation software.
 
-As each workload is run the workload harness records important stats.
-Some of the stats include how long the workload took to run, the min and max times of an individual query or insertion, and the standard deviations of individual runs.
+This paper primarily focuses on considerations for \textit{CGAT}'s database
+architecture. These considerations are made with regards to a MySQL cluster
+versus a Couchbase cluster to determine what database architecture is most
+efficient for running \textit{CGAT}. Section \ref{sec:motive} discusses our
+motivation to compare couchbase and MySQL in this paper, while section
+\ref{sec:design} talks about how we designed experimental workloads to use for
+our comparison of couchbase vs MySQL. In section \ref{sec:workload} we
+describe the workloads and how they are relevant to \textit{CGAT}, some brief
+implementation details are outlined in section \ref{sec:implementation}, we
+evaluate our results in section \ref{sec:eval} and finally conclude in
+section \ref{sec:conclusions}.
+
+\section{Motivation}\label{sec:motive}
+To evaluate the performance we auto-generated around
+400MB of data and inserted the same data into each cluster.  Obviously using a
+different data model for Couchbase to accommodate for the fact it is a NoSQL
+key-value store instead of a relational database. We then created a set of six
+experiments to run. Five came from common use cases found in the web
+application. While the other was designed to test the performance on each
+cluster of repeatedly writing and reading from the same object. Our major
+metrics are write performance, read performance, and time-delay between writes
+and reads of the same data. The rest of the paper is broken down as follows.
+Section 2 describes in depth how we designed the experiment. Section 3
+discusses the workloads we ran on each system. Section 4 describes our
+implementation.  Section 5 goes into the results of each of our experiments and
+some analysis on the data we obtained. Lastly, section 6 presents our
+conclusions.
+
+\section{Experiment Design}\label{sec:design}
+For our experiments we are using a total of five EC2 instances from the Amazon
+Web Services cloud. For both of our systems four servers were used for the
+database while the last was used for running our workload harness. The MySQL
+system is setup with one master server for writes and three slaves servers that
+replicate all of the masters data. Unlike the Couchbase system where each
+server handled a subset of the data.
+
+Our major driving force of this experiment was to see how Couchbase compared to
+MySQL for running a web application. To test which is better we created a
+workload harness. The main tasks of the workload harness was to define how many
+times a workload would run and which workloads to run. In our case workloads
+define a certain task. For example, getting of the data necessary to display a
+users profile. In our system a workload knows how to do its job in both MySQL
+and the Couchbase API. We leave describing each type workload we ran to the
+next section.
+
+As each workload is run the workload harness records important stats. Some of
+the stats include how long the workload took to run, the min and max times of
+an individual query or insertion, and the standard deviations of individual
+runs.
 
 \subsection{Data}
-The data we are storing comes from Ryan Verdon's masters thesis.
-It is the backend for a community genome annotation database.
-It contains data like genome sequences, genes, and species information.
-The representation of the data is vastly different in the two systems.
+The data we are storing comes from Ryan Verdon's masters thesis. It is the
+backend for a community genome annotation database. It contains data like
+genome sequences, genes, and species information. The representation of the
+data is vastly different in the two systems.
 
-But there exists four main pieces of data that are crucial to either representation.
-They are contigs, annotations, users, and groups.
-Contigs are what a biologist annotates to find genes.
-The main portion of a contig is a DNA sequence.
-Annotations are the record or formal representation of a gene that has been found in a specific contig.
-An annotation has several parts including coordinates in the contig which specify individual exons and meta information about the annotation like who uploaded it and when it was done, etc.
+But there exists four main pieces of data that are crucial to either
+representation. They are contigs, annotations, users, and groups. Contigs are
+what a biologist annotates to find genes. The main portion of a contig is a DNA
+sequence. Annotations are the record or formal representation of a gene that
+has been found in a specific contig. An annotation has several parts including
+coordinates in the contig which specify individual exons and meta information
+about the annotation like who uploaded it and when it was done, etc.
 
-Groups are another important part of the usefulness of the system.
-Users of like minds can join the same group to receive important information.
-Users is a major part of the data and is one of the central pillars that connects the data.
-Users upload contigs, create annotations, and join groups.
-Below we show the different ways we store the data in SQL and JSON.
+Groups are another important part of the usefulness of the system. Users of
+like minds can join the same group to receive important information. Users is a
+major part of the data and is one of the central pillars that connects the
+data. Users upload contigs, create annotations, and join groups. Below we show
+the different ways we store the data in SQL and JSON.
 
 \subsubsection{SQL}
-In SQL the data is represented as nine tables. A diagram of the tables with links to represent foreign key constraints is shown in Figure 1.
+In SQL the data is represented as nine tables. A diagram of the tables with
+links to represent foreign key constraints is shown in Figure 1.
 
 \subsubsection{JSON}
-For Couchbase we used four JSON objects which are shown in Figures 2-5.
-We left out the JSON object for the test table.
-It just consisted of two fields (Id and Name).
-All of the information from the SQL tables is represented in these JSON objects.
-We take advantage of nested objects to reduce as much subsequent requerying of the database as possible.
+For Couchbase we used four JSON objects which are shown in Figures 2-5. We left
+out the JSON object for the test table. It just consisted of two fields (Id and
+Name). All of the information from the SQL tables is represented in these JSON
+objects. We take advantage of nested objects to reduce as much subsequent
+requerying of the database as possible.
+
+\begin{figure}
+	\includegraphics[width=5.5in]{SQLDiagram.eps}
+	\caption{The SQL diagram for CGAT.}
+	\label{fig:SQLDiagram}
+\end{figure}
+
+%\clearpage
 
 \begin{figure}%
-	\includegraphics[width=5.5in]{SQLDiagram.png}%
-	\caption{The SQL diagram for CGAT.}
-	\label{fig:SQLDiagram}
-\end{figure}%
-
-\clearpage
-
-\begin{figure}%
-	\includegraphics[height=90mm,width=100mm]{user.png}
+	\includegraphics{user.eps}
 	\caption{The JSON for a user.}
@@
 \begin{figure}%
-	\includegraphics[height=70mm]{contig.png}
+	\includegraphics[height=70mm]{contig.eps}
   \caption{The JSON for a contig.}
@@
 \begin{figure}%
-	\includegraphics[height=100mm]{annotation.png}
+	\includegraphics[height=100mm]{annotation.eps}
 	\caption{The JSON for an annotation.}
@@
 \begin{figure}%
-	\includegraphics[height=20mm]{group.png}
+	\includegraphics[height=20mm]{group.eps}
 	\caption{The JSON for a group.}
@@
 
-\section{Workloads}
-The following workloads were ran on both the MySQL and Couchbase clusters.
+\section{Workloads}\label{sec:workload}
+The core unit of work that we used to compare couchbase performance against
+MySQL performance was called a \textit{workload}. Each workload was designed
+either to be write-heavy, read-heavy, or balanced. The number of workloads used
+during our comparison was determined by the number of members in our group (as
+specified by Alex Dekhtyar). Unfortunately due to time constraints we were
+unable to do multiple runs of any one workload, however, due to our environment
+and machines, there should be minimal fluctuations in any given run time for a
+workload.
 
-\subsection{Assigning a Task}
-This workload comes from a use-case in the web application. The main idea is that a new contig has been uploaded and some users that care about this contig need to be notified.
+In this section we further describe each task's purpose as well as what type of
+workload it was (write-heavy, etc.). This is necessary to be able to properly
+interpret (and agree or disagree with) the results described in section
+\ref{sec:eval}.
 
-\subsection{Profiling}
-The profiling workload simulates what happens when all of the information about a user is gathered.
-This includes: all groups they are a member of, all full and partial annotations they have submitted, and all tasks they have assigned.
-This is a read-heavy workload.
+\subsection{Profile View}
+The first thing that all users will do in \textit{CGAT} (ideally) is log in.
+When a user logs into \textit{CGAT} the user will be directed to his profile
+view which will contain all groups they are a member of, all completed
+annotations they have submitted, partial annotations that have been saved, and
+all tasks they have assigned. This is a use case that is likely to see many
+additions in the future as it serves as the first and foremost view of the
+system for the user. This means that any information that should be accessible
+-- including when a forum system is integrated into \textit{CGAT} -- will find
+a place on the user's profile.
 
-\subsection{Publishing}
-This workload involves updating a user's experience because of a completed annotation.
-The corresponding annotation record must also be set to complete, have its finished date updated, and have its gained experience set.
-This is a write-heavy workload.
+This workload is clearly a read-heavy workload that entails no disk writes at
+all. Important to note though, this use case may constantly be increasing
+(should constantly be increasing for a successful system) and so the load for
+this workload may increase quite a bit, and whenever new features are added
+there is a potential for a spike in workload.
 
-\subsection{Adding and Removing Group Membership}
-Adding or removing users to a group is a task common for classroom situations.
-In MySQL, this operation requires a join between three tables: Groups, Users, and GroupMembership.
-In Couchbase, this requires retrieving the key for the desired group and removing the user from the members field.
-This is a mixed workload involving reading and writing.
+%TODO make sure overview of CGAT defines what a task is, and mentions user
+%groups
+\subsection{Task Assignment}
+In \textit{CGAT}, it must be possible to request a group of users to annotate
+new or updated contigs. The frequency of this use case is dependent on the
+adoption of \textit{CGAT} -- the more species \textit{CGAT} supports, the more
+frequent this use case will be. More specifically, the relevant entities in the
+\textit{CGAT} schema are \textit{users}, \textit{contigs}, and \textit{groups}.
+Since a task embodies a request for a group $G$ to annotate a contig $C$, every
+user $U$ in $G$ receives a notification including a description of $C$ and a
+message requesting that it be annotated. A task can be assigned to any number
+of groups. Since tasks are assigned to groups that have some vested interest in
+the species or specific contig that has been updated or added, there may be
+many groups that a task should be assigned to. Additionally, as more species
+are supported by \textit{CGAT} or more users join \textit{CGAT}, more users
+will receive notifications of tasks and so there are multiple ways in which
+this use case must be scalable.
 
-\subsection{Uploading Contigs}
-This workload represents a contig being inserted into the database. All of the attributes of a contig are set, including the user who uploaded it and its sequence. This is a write-heavy workload.
+This workload is write-heavy -- adding tasks for each user in a group requires
+adding many records to a relational table or modifying many documents/values
+in a key/value store.
 
-\subsection{Read and Writing Performance}
-This is a test workload to examine the performance of writing and reading from the same object in succession. We are interested in the differences in performance of the two systems.
+%TODO make sure overview of CGAT mentions experience, and defines what an
+%annotation is
+\subsection{Annotation Publication}
+Annotations are the core focus of \textit{CGAT}. Every user is going to be
+doing some annotation, and they should be doing many annotations. Especially
+since \textit{CGAT} defines an annotation to be information about a gene
+isoform on a contig, even annotating a single gene will yield to several
+annotations in \textit{CGAT}. Additionally, when an annotation is submitted,
+the annotation must be marked as complete, the user and annotation then are
+assigned a certain amount of experience--the user gains experience towards
+his total, the annotation gains experience and serves as a history of the
+user's experience.
 
-\section{Implementation}
-Our experiment was implemented completely in Java. To do the MySQL portion of the workload we used
-the MySQL JDBC driver. For the Couchbase portion we used the Couchbase Java library from their website.
+This is a mostly write-, somewhat read-heavy workload. It is write-heavy in the
+sense that it must update a user and the annotation that the user completed.
+However, it also has some reads in the sense that the contig, for which the
+annotation is being submitted, is used to determine the amount of experience
+that submitting the annotation yields for the user. However, given the
+difference in data model between couchbase and MySQL, the couchbase
+implementation is much more write-heavy than the SQL equivalent. This is
+because the couchbase data model includes lists of complete and incomplete
+annotations for each user so that the annotation does not have to be fetched
+for every use case involving users and annotations. Since this is the case,
+more of the document undergoes modification for this use case (though I'm not
+sure if couchbase optimizes its writes in a way that many modifications in the
+same document is slower than few modifications in the same document).
 
-\section{Evaluation}
+\subsection{Group Membership Modification}
+Group membership is likely the second most stable of the workloads used in our
+comparisons of couchbase vs MySQL next to contig uploads. This is ``stable''
+because the majority of users will only join groups, and leaving groups will
+likely be relatively rare. This use case will only ever be as bad as how many
+users join at any one time--once a user registers for \textit{CGAT} and has
+joined some groups the user is unlikely to join or leave groups and so this use
+case poses almost no scalability problems. Although, despite this, there will
+be spikes that coincide with academic schedules, as groups will be created and
+populated for each new class that uses \textit{CGAT}.
+
+This workload is the most balanced of the workloads described in this paper
+(except maybe for ``stupid'' workload). Generally speaking, when a user joins a
+group then there is only the addition of a record in MySQL or the modification
+of the user document in couchbase. And when a user leaves a group, then
+subsequently the additional relational tuple or field in the user document is
+``undone'' by removing the group membership tuple or deleting the field from
+the user document.
+
+\subsection{Contig Uploads}
+Contig uploads is the most stable workload when the manual, or user, use case
+is considered. Although, there will be large spikes of activity when/if the
+reference genome for a species is updated--requiring \textit{CGAT} to update
+any and all of the contigs relevant to the genome. However, reference genomes
+are updated very infrequently and contigs will never be removed from
+\textit{CGAT}.
+
+This workload is write-heavy because it requires no reads--the use case is for
+uploading data, not downloading data.
+
+\subsection{Stupid}
+This is a test workload to examine the performance of writing and reading from
+the same object in succession. This is a workload contrived purely for better
+comparison of couchbase vs MySQL instead of workloads representative of what
+\textit{CGAT} will rely on. This is a naive, balanced workload.
+
+\section{Implementation}\label{sec:implementation}
+Our experiment was implemented completely in Java. To do the MySQL portion of
+the workload we used the MySQL JDBC driver. For the Couchbase portion we used
+the Couchbase Java library from their website.
+
+\section{Evaluation}\label{sec:eval}
 In this section we discuss the results from each workload in turn.
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 TODO

### 2012-11-27 Assignment6/paper/paper.tex
fixed some grammar

@@
 This workload is clearly a read-heavy workload that entails no disk writes at
-all. Important to note though, this use case may constantly be increasing
-(should constantly be increasing for a successful system) and so the load for
+all. Important to note though, this use case should constantly be increasing in a successful system and so the load for
 this workload may increase quite a bit, and whenever new features are added
-there is a potential for a spike in workload.
+there is a potential for a spike in this workload.
 

### 2012-11-27 Assignment6/paper/paper.tex
Edit pass on paper

--- 2012-11-27 Assignment6/paper/paper.tex
@@
 Gene annotation is the process of associating metadata about a gene with the
-contig on which the gene resides. This metadata is necessary for conducting
+contig, a dna sequence, on which the gene resides. This metadata is necessary for conducting
 genomic research projects and analysis involving the contig's genomic sequence.
-Currently, there is no specifically gene annotation software. Genome browsers
-all maintain lots of genomic data and are used to be able to annotate genes and
-contigs, but do not provide a user-friendly interface. While genome browsers
+Currently, there is no specialty gene annotation software. Genome browsers
+all maintain lots of genomic data and are used when manually annotating genes,
+but do not provide a user-friendly interface. While genome browsers
 will likely never be replaced (UCSC genome browser, etc. are well
 established), it would be desirable to have a software system that better
-accomodates the visual and informational needs of gene annotation.
+accommodates the visual and informational needs of gene annotation.
 
@@
 provide users with a way to view gene annotations without extra, unnecessary
-information obscuring their view while also embodying a wikipedia-like emphasis
+information obscuring their view while also embodying a Wikipedia-like emphasis
 on collaboration and openness.
@@
 architecture. These considerations are made with regards to a MySQL cluster
-versus a Couchbase cluster to determine what database architecture is most
+versus a Couchbase cluster to determine which database architecture is most
 efficient for running \textit{CGAT}. Section \ref{sec:motive} discusses our
-motivation to compare couchbase and MySQL in this paper, while section
+motivation to compare Couchbase and MySQL in this paper, while section
 \ref{sec:design} talks about how we designed experimental workloads to use for
-our comparison of couchbase vs MySQL. In section \ref{sec:workload} we
+our comparison of Couchbase vs MySQL. In section \ref{sec:workload} we
 describe the workloads and how they are relevant to \textit{CGAT}, some brief
@@
 \section{Motivation}\label{sec:motive}
+Traditional web applications are almost always backed by a relational database.
+However, there has been a recent movement towards using NoSQL databases for
+web application. Our main goal is to determine whether a NoSQL database, Couchbase,
+can outperform a traditional relational database, MySQL, for \textit{CGAT}'s
+common operations.
+
+\section{Experiment Design}\label{sec:design}
 To evaluate the performance we auto-generated around
@@
 experiments to run. Five came from common use cases found in the web
-application. While the other was designed to test the performance on each
-cluster of repeatedly writing and reading from the same object. Our major
+application. The final experiment was a contrived one designed to test the performance on each
+cluster by repeatedly writing and reading from the same object. Our major
 metrics are write performance, read performance, and time-delay between writes
-and reads of the same data. The rest of the paper is broken down as follows.
-Section 2 describes in depth how we designed the experiment. Section 3
-discusses the workloads we ran on each system. Section 4 describes our
-implementation.  Section 5 goes into the results of each of our experiments and
-some analysis on the data we obtained. Lastly, section 6 presents our
-conclusions.
+and reads of the same data.
 
-\section{Experiment Design}\label{sec:design}
 For our experiments we are using a total of five EC2 instances from the Amazon
-Web Services cloud. For both of our systems four servers were used for the
-database while the last was used for running our workload harness. The MySQL
+Web Services cloud offering. For both of our systems four servers were used for the
+database cluster while the last was used for running our workload harness. The MySQL
 system is setup with one master server for writes and three slaves servers that
@@
 define a certain task. For example, getting of the data necessary to display a
-users profile. In our system a workload knows how to do its job in both MySQL
+user's profile. In our system a workload knows how to do its job in both MySQL
 and the Couchbase API. We leave describing each type workload we ran to the
@@
 \subsection{Data}
-The data we are storing comes from Ryan Verdon's masters thesis. It is the
-backend for a community genome annotation database. It contains data like
-genome sequences, genes, and species information. The representation of the
-data is vastly different in the two systems.
-
-But there exists four main pieces of data that are crucial to either
-representation. They are contigs, annotations, users, and groups. Contigs are
+There exists four main pieces of data that are crucial to either
+representation: contigs, annotations, users, and groups. Contigs are
 what a biologist annotates to find genes. The main portion of a contig is a DNA
@@
 has been found in a specific contig. An annotation has several parts including
-coordinates in the contig which specify individual exons and meta information
+coordinates in the contig which specify individual parts of the gene, exons, and meta information
 about the annotation like who uploaded it and when it was done, etc.
@@
 Groups are another important part of the usefulness of the system. Users of
-like minds can join the same group to receive important information. Users is a
+like minds or research targets can join the same group to receive important information. Users is a
 major part of the data and is one of the central pillars that connects the
 data. Users upload contigs, create annotations, and join groups. Below we show
-the different ways we store the data in SQL and JSON.
+the different ways we store the data in SQL and JSON. We chose to represent all of our
+data in Couchbase in JSON because it is a naturally serializable data format with
+large library support. In addition, using JSON would allow use to share the same
+schema with other popular document-based NoSQL databases, such as MongoDB.
 
@@
 In SQL the data is represented as nine tables. A diagram of the tables with
-links to represent foreign key constraints is shown in Figure 1.
+links to represent foreign key constraints is shown in Figure \ref{fig:SQLDiagram}.
 
 \subsubsection{JSON}
-For Couchbase we used four JSON objects which are shown in Figures 2-5. We left
-out the JSON object for the test table. It just consisted of two fields (Id and
-Name). All of the information from the SQL tables is represented in these JSON
+For Couchbase we used four JSON objects which are shown in Figures \ref{fig:JSON-user}-\ref{fig:JSON-group}.
+All of the information from the SQL tables is represented in these JSON
 objects. We take advantage of nested objects to reduce as much subsequent
@@
 \section{Workloads}\label{sec:workload}
-The core unit of work that we used to compare couchbase performance against
+The core unit of work that we used to compare Couchbase performance against
 MySQL performance was called a \textit{workload}. Each workload was designed
@@
 workload it was (write-heavy, etc.). This is necessary to be able to properly
-interpret (and agree or disagree with) the results described in section
+interpret (and agree or disagree with) the results described in Section
 \ref{sec:eval}.
@@
 The first thing that all users will do in \textit{CGAT} (ideally) is log in.
-When a user logs into \textit{CGAT} the user will be directed to his profile
+When a user logs into \textit{CGAT} the user will be directed to their profile
 view which will contain all groups they are a member of, all completed
@@
 
-%TODO make sure overview of CGAT defines what a task is, and mentions user
-%groups
 \subsection{Task Assignment}
@@
 
-%TODO make sure overview of CGAT mentions experience, and defines what an
-%annotation is
 \subsection{Annotation Publication}
@@
 since \textit{CGAT} defines an annotation to be information about a gene
-isoform on a contig, even annotating a single gene will yield to several
+isoform, variation, on a contig, even annotating a single gene will yield to several
 annotations in \textit{CGAT}. Additionally, when an annotation is submitted,
@@
 that submitting the annotation yields for the user. However, given the
-difference in data model between couchbase and MySQL, the couchbase
+difference in data model between Couchbase and MySQL, the Couchbase
 implementation is much more write-heavy than the SQL equivalent. This is
-because the couchbase data model includes lists of complete and incomplete
+because the Couchbase data model includes lists of complete and incomplete
 annotations for each user so that the annotation does not have to be fetched
 for every use case involving users and annotations. Since this is the case,
-more of the document undergoes modification for this use case (though I'm not
-sure if couchbase optimizes its writes in a way that many modifications in the
-same document is slower than few modifications in the same document).
+more of the document undergoes modification for this use case.
 
@@
 Group membership is likely the second most stable of the workloads used in our
-comparisons of couchbase vs MySQL next to contig uploads. This is ``stable''
+comparisons of Couchbase vs MySQL next to contig uploads. This is ``stable''
 because the majority of users will only join groups, and leaving groups will
@@
 This workload is the most balanced of the workloads described in this paper
-(except maybe for ``stupid'' workload). Generally speaking, when a user joins a
+(except maybe for ``Read/Write'' workload). Generally speaking, when a user joins a
 group then there is only the addition of a record in MySQL or the modification
-of the user document in couchbase. And when a user leaves a group, then
+of the user and group document in Couchbase. And when a user leaves a group, then
 subsequently the additional relational tuple or field in the user document is
 ``undone'' by removing the group membership tuple or deleting the field from
-the user document.
+the user and group document.
 
@@
 
-\subsection{Stupid}
+\subsection{Read/Write}
 This is a test workload to examine the performance of writing and reading from
 the same object in succession. This is a workload contrived purely for better
-comparison of couchbase vs MySQL instead of workloads representative of what
+comparison of Couchbase vs MySQL instead of workloads representative of what
 \textit{CGAT} will rely on. This is a naive, balanced workload.
@@
 the workload we used the MySQL JDBC driver. For the Couchbase portion we used
-the Couchbase Java library from their website.
+the official Couchbase Java library from their website.
 

### 2012-11-28 Assignment6/paper/paper.tex
Added basic tables for most of the stats.
Adding evaluation and conclusion section to the paper. Still need the time for the Assign workload and edits of my engrish.

@@
 
-%section on each workload and the results on each system (could be a table tbh and not a graph)
-%probably some weak analysis on why we expected the results
-%need to ask eriq the # of times each workload was  run, might be in an email?
+\subsection{Assigning a Task}
 
-\subsection{Assigning a Task}
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+	{\bf Metric} & {\bf Result} \\
+	\hline
+        \# of runs & 100000 \\
+        \hline
+        MySQL runtime (ms) & 39325189 \\
+	\hline
+        Couchbase runtime (ms) & ?* \\
+        \hline
+\end{tabular}
+\end{center}
+
+From the results it is obvious that MySQL performs much better than Couchbase. The result we have for
+Couchbase is actually a minimum runtime. It was running so long we killed it to get more results. Even with this
+in mind, MySQL vastly outperforms Couchbase. We believe this is because during the process of assigning a task
+a couple values have to be read into memory before a finally write can be done. Compared to MySQL where its a simple 
+ad-hoc query to find the necessary values prior to inserting the new task.
+
+From our analysis, we believe if we had document or ad-hoc unstructured queries in Couchbase we could vastly improve the performance.
+When doing the Couchbase queries we only need a small subset of the JSON. By only querying for the necessary parts it could save vast amounts of time.
 
@@
 
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+	{\bf Metric} & {\bf Result} \\
+	\hline
+        \# of runs & 100000 \\
+        \hline
+        MySQL runtime (ms) & 2908257 \\
+	\hline
+        Couchbase runtime (ms) & 7008818 \\
+        \hline
+\end{tabular}
+\end{center}
+
+Like the last workload, MySQL outperforms Couchbase. We believe that this is because this workload is read-heavy so several
+full documents are retrieved per run.  Again, this workload could be improved in Couchbase with the use of unstructured queries.
+
 \subsection{Publishing}
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+	{\bf Metric} & {\bf Result} \\
+	\hline
+        \# of runs & 100000 \\
+        \hline
+        MySQL runtime (ms) & 510837 \\
+	\hline
+        Couchbase runtime (ms) & 1830064 \\
+        \hline
+\end{tabular}
+\end{center}
+
+Again, a similar pattern in the results as the last two workloads. This was another workload that required some queries prior to inserting
+a new annotation. We believe this is another case where unstructured queries could vastly improve performance.
 
@@
 
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+	{\bf Metric} & {\bf Result} \\
+	\hline
+        \# of runs & 100000 \\
+        \hline
+        MySQL runtime (ms) & 316673 \\
+	\hline
+        Couchbase runtime (ms) & 9086304 \\
+        \hline
+\end{tabular}
+\end{center}
+
+To our surprise this workload performed extremely bad on Couchbase. We were not expecting this because this is a 50\% read
+and 50\% write workload. Having to do lots of reads prior to inserting new group memberships really hurt Couchbase. This is most definitely 
+a case where having unstructured queries or document support would greatly improve performance. The way we store group memberships
+is as a sub-field in the user and group objects. By only queries for those fields we could reduce the amount of data we return to the workload harness by over 70\%.
+
 \subsection{Uploading Contigs}
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+	{\bf Metric} & {\bf Result} \\
+	\hline
+        \# of runs & 10000 \\
+        \hline
+        MySQL runtime (ms) & 119892 \\
+	\hline
+        Couchbase runtime (ms) & 20133 \\
+        \hline
+\end{tabular}
+\end{center}
+
+The results from this test were no surprise. Couchbase demolished MySQL when most of the workload was insertions. Even with
+contig objects containing strings of over 50k characters Couchbase performed amazingly well. By not having to lock key-value pairs
+Couchbase can get some awesome write performance.
 
@@
 
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+	{\bf Metric} & {\bf Result} \\
+	\hline
+        \# of runs & 100000 \\
+        \hline
+        MySQL runtime (ms) & 103808 \\
+	\hline
+        Couchbase runtime (ms) & 58581 \\
+        \hline
+\end{tabular}
+\end{center}
+
+The results from this test were not that surprising based on the results from our other tests.
+It is obvious that Couchbase is much faster in terms of writes plus Couchbase can leverage Memcache in this workload to speed up queries to the database.
+For this workload we decided to gather a bunch more numbers to see what the minimum, maximum, mean, median and standard deviation of the individual reads and writes for each system.
+Below we list the six tables and discuss what the results mean.
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+  	\multicolumn{2}{|c|}{{\bf SQL read after writes of same object}} \\
+ 	\hline
+  	 Min (ms) & 0 \\
+  	\hline
+         Max (ms) & 25 \\
+  	\hline
+  	Mean (ms) & 0.473 \\
+ 	\hline
+  	Median (ms) & 0 \\
+  	\hline
+         Standard Deviation (ms) & 0.685 \\
+  	\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+  	\multicolumn{2}{|c|}{{\bf SQL write after reads of same object}} \\
+ 	\hline
+  	 Min (ms) & 0 \\
+  	\hline
+         Max (ms) & 70 \\
+  	\hline
+  	Mean (ms) & 0.558 \\
+ 	\hline
+  	Median (ms) & 1 \\
+  	\hline
+         Standard Deviation (ms) & 0.859 \\
+  	\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+  	\multicolumn{2}{|c|}{{\bf SQL combined results}} \\
+ 	\hline
+  	 Min (ms) & 0 \\
+  	\hline
+         Max (ms) & 70 \\
+  	\hline
+  	Mean (ms) & 0.516 \\
+ 	\hline
+  	Median (ms) & 0 \\
+  	\hline
+         Standard Deviation (ms) & 0.778 \\
+  	\hline
+\end{tabular}
+\end{center}
+
+The three prior tables are all of our SQL results. The first table lists the minimum, maximum, mean, median, and standard deviation of reads 
+that occur after writes. The second table lists the same metrics but for writes after reads. Lastly, the third table lists the metrics for
+all of the queries. Below we list the same three tables but this time show the Couchbase results.
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+  	\multicolumn{2}{|c|}{{\bf Couchbase read after writes of same object}} \\
+ 	\hline
+  	 Min (ms) & 0 \\
+  	\hline
+         Max (ms) & 71 \\
+  	\hline
+  	Mean (ms) & 0.562 \\
+ 	\hline
+  	Median (ms) & 1 \\
+  	\hline
+         Standard Deviation (ms) & 0.859 \\
+  	\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+  	\multicolumn{2}{|c|}{{\bf Couchbase write after reads of same object}} \\
+ 	\hline
+  	 Min (ms) & 0 \\
+  	\hline
+         Max (ms) & 48 \\
+  	\hline
+  	Mean (ms) & 0.017 \\
+ 	\hline
+  	Median (ms) & 0 \\
+  	\hline
+         Standard Deviation (ms) & 0.213 \\
+  	\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{|l|l|}
+	\hline
+  	\multicolumn{2}{|c|}{{\bf Couchbase combined results}} \\
+ 	\hline
+  	 Min (ms) & 0 \\
+  	\hline
+         Max (ms) & 71 \\
+  	\hline
+  	Mean (ms) & 0.29 \\
+ 	\hline
+  	Median (ms) & 0 \\
+  	\hline
+         Standard Deviation (ms) & 0.682 \\
+  	\hline
+\end{tabular}
+\end{center}
+
+By examining the read after writes results from the two systems we get the results we expect based on our other tests. MySQL outperforms
+Couchbase when it comes to reading after a write. The really interesting results come from the the write after read tables where it is obvious that
+Couchbase demolished MySQL by a lot. There is a huge difference in the means alone without even looking at the other metrics. This again
+backs up our other results from the contig workload where Couchbase outperformed MySQL in a mainly write task. Overall Couchbase outperformed MySQL because of its blazing fast writes after reads.
+
 \section{Conclusions}\label{sec:conclusions}
-TODO
+From our results MySQL appears to be the winner right now for the CGAT backend.
+The MySQL workloads outperformed Couchbase the majority of the time.
+We believe in a live scenario Couchbase would perform better than our tests mainly because our tests could not take advantage of Memcache.
+When a key had to be looked up there was almost no chance it was sitting in Memcache.
+We also believe that the Couchbase results would greatly improve with the addition of an unstructured query language.
+With the addition of this feature our workloads would be making much smaller requests from the database.
 

### 2012-11-28 Assignment6/paper/paper.tex
Consolidated data into 3 different tables, moved JSON and ER-diagram to be an appendix at the end of the paper, changed things in the tables so that they would fit into a single column.
If someone thinks more description is needed for a table try to put it into the
caption so that the tables themselves fit into a single column (it's a bit
cleaner this way since the numbers we have are not very big, it's just the
names of the columns, etc.)

Also, added the runtime for the assign task workload in couch.log

@@
 % Paper for the ChouchBase assignment.
-
-\documentclass[]{IEEEtran}
+\documentclass[10pt, conference, compsocconf]{IEEEtran}
 
 \usepackage{float}
-\usepackage{url}
+%\usepackage{url}
 \usepackage{graphicx}
+\usepackage{subfig}
 \usepackage{color}
@@
 
-This paper primarily focuses on considerations for \textit{CGAT}'s database
+This paper primarily focuses on considerations for \textit{CGAT}s database
 architecture. These considerations are made with regards to a MySQL cluster
@@
 
-\begin{figure}
-	\includegraphics[width=5.5in]{SQLDiagram.eps}
-	\caption{The SQL diagram for CGAT.}
-	\label{fig:SQLDiagram}
-\end{figure}
-
-%\clearpage
-
-\begin{figure}%
-	\includegraphics{user.eps}
-	\caption{The JSON for a user.}
-	\label{fig:JSON-user}
-\end{figure}%
-
-\begin{figure}%
-	\includegraphics[height=70mm]{contig.eps}
-         \caption{The JSON for a contig.}
-         \label{fig:JSON-contig}
-\end{figure}%
-
-\begin{figure}%
-	\includegraphics[height=100mm]{annotation.eps}
-	\caption{The JSON for an annotation.}
-	\label{fig:JSON-annotation}
-\end{figure}%
-
-\begin{figure}%
-	\includegraphics[height=20mm]{group.eps}
-	\caption{The JSON for a group.}
-	\label{fig:JSON-group}
-\end{figure}%
-
-\clearpage
+\begin{table}[t]
+   \centering
+      \begin{tabular}{| l | c | c | c |}
+         \hline
+            {\textbf{Workload}} &
+            {\textbf{MySQL}} &
+            {\textbf{Couchbase}} \\
+         \hline
+            Task Assignment         & 39325189ms & 157431049ms \\
+         \hline
+            Profile View            & 2908257ms & 7008818ms \\
+         \hline
+            Annotation Publication  & 510837ms & 1830064ms \\
+         \hline
+            Group Modification      & 316673ms & 9086304ms \\
+         \hline
+            Contig Upload           & 119892ms & 20133ms \\
+         \hline
+            Read/Write              & 103808ms & 58581ms \\
+         \hline
+      \end{tabular}
+      \caption{Comparison of runtimes in ms for workloads using MySQL and Couchbase.
+               Each workload was run for 100000 iterations.}
+      \label{tab:workload_perf}
+\end{table}
 
@@
 
-\subsection{Read/Write}
+\subsection{Read/Write Performance}
 This is a test workload to examine the performance of writing and reading from
@@
 
+\begin{table}[t]
+   \centering
+      \begin{tabular}{| l | c | c | c |}
+         \hline
+            \multicolumn{4}{|c|}{{\textbf{MySQL read/write performance}}} \\
+         \hline
+            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
+         \hline
+            Min (ms) & 0 & 0 & 0 \\
+         \hline
+            Max (ms) & 25 & 70 & 70 \\
+         \hline
+            Mean (ms) & 0.473 & 0.558 & 0.516 \\
+         \hline
+            Median (ms) & 0 & 1 & 0 \\
+         \hline
+            Standard Deviation (ms) & 0.685 & 0.859 & 0.778 \\
+         \hline
+      \end{tabular}
+      \caption{Comparison of reads and writes to the same object in MySQL.}
+      \label{tab:sql_perf}
+\end{table}
+
+\begin{table}[t]
+   \centering
+      \begin{tabular}{| l | c | c | c |}
+         \hline
+            \multicolumn{4}{|c|}{{\textbf{Couchbase read/write performance}}} \\
+         \hline
+            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
+         \hline
+            Min (ms) & 0 & 0 & 0 \\
+         \hline
+            Max (ms) & 71 & 48 & 71 \\
+         \hline
+            Mean (ms) & 0.562 & 0.017 & 0.290 \\
+         \hline
+            Median (ms) & 1 & 0 & 0 \\
+         \hline
+            Standard Deviation (ms) & 0.859 & 0.213 & 0.682 \\
+         \hline
+      \end{tabular}
+      \caption{Comparison of reads and writes to the same object in couchbase}
+      \label{tab:couch_perf}
+\end{table}
+
 \section{Evaluation}\label{sec:eval}
@@
 
-\subsection{Assigning a Task}
+\subsection{Task Assignment}
+From the results it is obvious that MySQL performs much better than Couchbase.
+The result we have for Couchbase is actually a minimum runtime. It was running
+so long we killed it to get more results. Even with this in mind, MySQL vastly
+outperforms Couchbase. We believe this is because during the process of
+assigning a task a couple values have to be read into memory before a finally
+write can be done. Compared to MySQL where its a simple ad-hoc query to find
+the necessary values prior to inserting the new task.
 
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-	{\bf Metric} & {\bf Result} \\
-	\hline
-        \# of runs & 100000 \\
-        \hline
-        MySQL runtime (ms) & 39325189 \\
-	\hline
-        Couchbase runtime (ms) & ?* \\
-        \hline
-\end{tabular}
-\end{center}
+From our analysis, we believe if we had document or ad-hoc unstructured queries
+in Couchbase we could vastly improve the performance.  When doing the Couchbase
+queries we only need a small subset of the JSON. By only querying for the
+necessary parts it could save vast amounts of time.
 
-From the results it is obvious that MySQL performs much better than Couchbase. The result we have for
-Couchbase is actually a minimum runtime. It was running so long we killed it to get more results. Even with this
-in mind, MySQL vastly outperforms Couchbase. We believe this is because during the process of assigning a task
-a couple values have to be read into memory before a finally write can be done. Compared to MySQL where its a simple 
-ad-hoc query to find the necessary values prior to inserting the new task.
+\subsection{Profile View}
+Like the last workload, MySQL outperforms Couchbase. We believe that this is
+because this workload is read-heavy so several full documents are retrieved per
+run.  Again, this workload could be improved in Couchbase with the use of
+unstructured queries.
 
-From our analysis, we believe if we had document or ad-hoc unstructured queries in Couchbase we could vastly improve the performance.
-When doing the Couchbase queries we only need a small subset of the JSON. By only querying for the necessary parts it could save vast amounts of time.
+\subsection{Annotation Publication}
+Again, a similar pattern in the results as the last two workloads. This was
+another workload that required some queries prior to inserting a new
+annotation. We believe this is another case where unstructured queries could
+vastly improve performance.
 
-\subsection{Profiling}
+\subsection{Group Modification}
+To our surprise this workload performed extremely bad on Couchbase. We were not
+expecting this because this is a 50\% read and 50\% write workload. Having to
+do lots of reads prior to inserting new group memberships really hurt
+Couchbase. This is most definitely a case where having unstructured queries or
+document support would greatly improve performance. The way we store group
+memberships is as a sub-field in the user and group objects. By only queries
+for those fields we could reduce the amount of data we return to the workload
+harness by over 70\%.
 
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-	{\bf Metric} & {\bf Result} \\
-	\hline
-        \# of runs & 100000 \\
-        \hline
-        MySQL runtime (ms) & 2908257 \\
-	\hline
-        Couchbase runtime (ms) & 7008818 \\
-        \hline
-\end{tabular}
-\end{center}
+\subsection{Contig Uploads}
+The results from this test were no surprise. Couchbase demolished MySQL when
+most of the workload was insertions. Even with contig objects containing
+strings of over 50k characters Couchbase performed amazingly well. By not
+having to lock key-value pairs Couchbase can get some awesome write
+performance.
 
-Like the last workload, MySQL outperforms Couchbase. We believe that this is because this workload is read-heavy so several
-full documents are retrieved per run.  Again, this workload could be improved in Couchbase with the use of unstructured queries.
+\subsection{Read/Write Performance}
+The results from this test were not that surprising based on the results from
+our other tests. It is obvious that Couchbase is much faster in terms of writes
+plus Couchbase can leverage Memcache in this workload to speed up queries to
+the database. For this workload we decided to gather a bunch more numbers to
+see what the minimum, maximum, mean, median and standard deviation of the
+individual reads and writes for each system. Below we list the six tables and
+discuss what the results mean.
 
-\subsection{Publishing}
+The three prior tables are all of our SQL results. The first table lists the
+minimum, maximum, mean, median, and standard deviation of reads that occur
+after writes. The second table lists the same metrics but for writes after
+reads. Lastly, the third table lists the metrics for all of the queries. Below
+we list the same three tables but this time show the Couchbase results.
 
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-	{\bf Metric} & {\bf Result} \\
-	\hline
-        \# of runs & 100000 \\
-        \hline
-        MySQL runtime (ms) & 510837 \\
-	\hline
-        Couchbase runtime (ms) & 1830064 \\
-        \hline
-\end{tabular}
-\end{center}
-
-Again, a similar pattern in the results as the last two workloads. This was another workload that required some queries prior to inserting
-a new annotation. We believe this is another case where unstructured queries could vastly improve performance.
-
-\subsection{Adding and Removing Group Membership}
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-	{\bf Metric} & {\bf Result} \\
-	\hline
-        \# of runs & 100000 \\
-        \hline
-        MySQL runtime (ms) & 316673 \\
-	\hline
-        Couchbase runtime (ms) & 9086304 \\
-        \hline
-\end{tabular}
-\end{center}
-
-To our surprise this workload performed extremely bad on Couchbase. We were not expecting this because this is a 50\% read
-and 50\% write workload. Having to do lots of reads prior to inserting new group memberships really hurt Couchbase. This is most definitely 
-a case where having unstructured queries or document support would greatly improve performance. The way we store group memberships
-is as a sub-field in the user and group objects. By only queries for those fields we could reduce the amount of data we return to the workload harness by over 70\%.
-
-\subsection{Uploading Contigs}
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-	{\bf Metric} & {\bf Result} \\
-	\hline
-        \# of runs & 10000 \\
-        \hline
-        MySQL runtime (ms) & 119892 \\
-	\hline
-        Couchbase runtime (ms) & 20133 \\
-        \hline
-\end{tabular}
-\end{center}
-
-The results from this test were no surprise. Couchbase demolished MySQL when most of the workload was insertions. Even with
-contig objects containing strings of over 50k characters Couchbase performed amazingly well. By not having to lock key-value pairs
-Couchbase can get some awesome write performance.
-
-\subsection{Read and Writing Performance}
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-	{\bf Metric} & {\bf Result} \\
-	\hline
-        \# of runs & 100000 \\
-        \hline
-        MySQL runtime (ms) & 103808 \\
-	\hline
-        Couchbase runtime (ms) & 58581 \\
-        \hline
-\end{tabular}
-\end{center}
-
-The results from this test were not that surprising based on the results from our other tests.
-It is obvious that Couchbase is much faster in terms of writes plus Couchbase can leverage Memcache in this workload to speed up queries to the database.
-For this workload we decided to gather a bunch more numbers to see what the minimum, maximum, mean, median and standard deviation of the individual reads and writes for each system.
-Below we list the six tables and discuss what the results mean.
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-  	\multicolumn{2}{|c|}{{\bf SQL read after writes of same object}} \\
- 	\hline
-  	 Min (ms) & 0 \\
-  	\hline
-         Max (ms) & 25 \\
-  	\hline
-  	Mean (ms) & 0.473 \\
- 	\hline
-  	Median (ms) & 0 \\
-  	\hline
-         Standard Deviation (ms) & 0.685 \\
-  	\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-  	\multicolumn{2}{|c|}{{\bf SQL write after reads of same object}} \\
- 	\hline
-  	 Min (ms) & 0 \\
-  	\hline
-         Max (ms) & 70 \\
-  	\hline
-  	Mean (ms) & 0.558 \\
- 	\hline
-  	Median (ms) & 1 \\
-  	\hline
-         Standard Deviation (ms) & 0.859 \\
-  	\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-  	\multicolumn{2}{|c|}{{\bf SQL combined results}} \\
- 	\hline
-  	 Min (ms) & 0 \\
-  	\hline
-         Max (ms) & 70 \\
-  	\hline
-  	Mean (ms) & 0.516 \\
- 	\hline
-  	Median (ms) & 0 \\
-  	\hline
-         Standard Deviation (ms) & 0.778 \\
-  	\hline
-\end{tabular}
-\end{center}
-
-The three prior tables are all of our SQL results. The first table lists the minimum, maximum, mean, median, and standard deviation of reads 
-that occur after writes. The second table lists the same metrics but for writes after reads. Lastly, the third table lists the metrics for
-all of the queries. Below we list the same three tables but this time show the Couchbase results.
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-  	\multicolumn{2}{|c|}{{\bf Couchbase read after writes of same object}} \\
- 	\hline
-  	 Min (ms) & 0 \\
-  	\hline
-         Max (ms) & 71 \\
-  	\hline
-  	Mean (ms) & 0.562 \\
- 	\hline
-  	Median (ms) & 1 \\
-  	\hline
-         Standard Deviation (ms) & 0.859 \\
-  	\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-  	\multicolumn{2}{|c|}{{\bf Couchbase write after reads of same object}} \\
- 	\hline
-  	 Min (ms) & 0 \\
-  	\hline
-         Max (ms) & 48 \\
-  	\hline
-  	Mean (ms) & 0.017 \\
- 	\hline
-  	Median (ms) & 0 \\
-  	\hline
-         Standard Deviation (ms) & 0.213 \\
-  	\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-\begin{tabular}{|l|l|}
-	\hline
-  	\multicolumn{2}{|c|}{{\bf Couchbase combined results}} \\
- 	\hline
-  	 Min (ms) & 0 \\
-  	\hline
-         Max (ms) & 71 \\
-  	\hline
-  	Mean (ms) & 0.29 \\
- 	\hline
-  	Median (ms) & 0 \\
-  	\hline
-         Standard Deviation (ms) & 0.682 \\
-  	\hline
-\end{tabular}
-\end{center}
-
-By examining the read after writes results from the two systems we get the results we expect based on our other tests. MySQL outperforms
-Couchbase when it comes to reading after a write. The really interesting results come from the the write after read tables where it is obvious that
-Couchbase demolished MySQL by a lot. There is a huge difference in the means alone without even looking at the other metrics. This again
-backs up our other results from the contig workload where Couchbase outperformed MySQL in a mainly write task. Overall Couchbase outperformed MySQL because of its blazing fast writes after reads.
+By examining the read after writes results from the two systems we get the
+results we expect based on our other tests. MySQL outperforms Couchbase when it
+comes to reading after a write. The really interesting results come from the
+the write after read tables where it is obvious that Couchbase demolished MySQL
+by a lot. There is a huge difference in the means alone without even looking at
+the other metrics. This again backs up our other results from the contig
+workload where Couchbase outperformed MySQL in a mainly write task. Overall
+Couchbase outperformed MySQL because of its blazing fast writes after reads.
 
@@
 From our results MySQL appears to be the winner right now for the CGAT backend.
-The MySQL workloads outperformed Couchbase the majority of the time.
-We believe in a live scenario Couchbase would perform better than our tests mainly because our tests could not take advantage of Memcache.
-When a key had to be looked up there was almost no chance it was sitting in Memcache.
-We also believe that the Couchbase results would greatly improve with the addition of an unstructured query language.
-With the addition of this feature our workloads would be making much smaller requests from the database.
+The MySQL workloads outperformed Couchbase the majority of the time. We believe
+in a live scenario Couchbase would perform better than our tests mainly because
+our tests could not take advantage of Memcache. When a key had to be looked up
+there was almost no chance it was sitting in Memcache. We also believe that the
+Couchbase results would greatly improve with the addition of an unstructured
+query language. With the addition of this feature our workloads would be making
+much smaller requests from the database.
 
@@
 
+\newpage
+\appendices
+\begin{figure*}
+   \centering
+   \includegraphics[height=90mm]{SQLDiagram.eps}
+	\caption{The SQL diagram for CGAT.}
+	\label{fig:SQLDiagram}
+\end{figure*}
+
+\begin{figure*}{t}
+   \centering
+   \subfloat[Contig JSON]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
+   \subfloat[Annotation JSON]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
+\end{figure*}
+
+\begin{figure*}{t}
+   \centering
+   \subfloat[User JSON]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
+   \subfloat[Group JSON]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
+\end{figure*}
+
+%\begin{figure*}
+%  \includegraphics[height=90mm]{user.eps}
+%	\caption{The JSON for a user.}
+%	\label{fig:JSON-user}
+%\end{figure*}
+%\begin{figure*}
+%  \includegraphics[height=25mm]{group.eps}
+%	\caption{The JSON for a group.}
+%	\label{fig:JSON-group}
+%\end{figure*}
+%\begin{figure*}
+%  \includegraphics[height=80mm]{contig.eps}
+%  \caption{The JSON for a contig.}
+%  \label{fig:JSON-contig}
+%\end{figure*}
+%\begin{figure*}
+%  \includegraphics[height=90mm]{annotation.eps}
+%	\caption{The JSON for an annotation.}
+%	\label{fig:JSON-annotation}
+%\end{figure*}
+
 \end{document}

### 2012-11-28 ui/js/annotation.js
Changed Start to Begin

@@ -181,7 +181,7 @@ function startEndValidate(start, end, errorDump) {
    if (start < 0 || end < 0 ||
        start >= window.cgat.dna.length ||
        end >= window.cgat.dna.length) {
-      validationError('Start and End must be >= 0 and < ' + window.cgat.dna.length,
+      validationError('Gene Start and End must be >= 0 and < ' + window.cgat.dna.length,
                       errorDump);
       return false;
    } else if (start >= end) {
@@ -291,7 +291,7 @@ function createExonElement(start, end, key) {
    exonElement.classList.add('exon');
 
    var exonElementString =
-         "<span>Start: </span>" +
+         "<span>Begin: </span>" +
          "<input type='number' id='exon-start-" + key + "'" +
             " value=" + start +
             " onchange='updateExon(" + key + ");'/>" +

### 2012-11-28 ui/js/view-annotation.js

@@ -49,7 +49,7 @@ document.addEventListener('DOMContentLoaded', function() {
          if (data.annotation.exons) {
             var exons = '';
             data.annotation.exons.forEach(function(exon) {
-               exons += "<p>Start: " + exon.start + ", End: " + exon.end + "</p>";
+               exons += "<p>Begin: " + exon.start + ", End: " + exon.end + "</p>";
             });
             $('#exons-inset').html(exons);
          }

### 2012-11-28 ui/js/annotation.js
Tick marks on top dna

@@ -203,15 +203,18 @@ function updateBoundingMarkers() {
    var startSize = Math.floor(window.cgat.geneStart * sizeRatio);
    var endSize = Math.floor((window.cgat.dna.length - window.cgat.geneEnd) * sizeRatio);
 
+   console.log("len: " + window.cgat.dna.length + ", geneEnd: " + window.cgat.geneEnd);
+   console.log("Start: " + startSize + ", End: " + endSize);
+
    var startBound = document.createElement('div');
-   startBound.className = 'top-dna-boundary-marker';
+   startBound.className = 'top-dna-boundary-marker top-dna-boundary-marker-begin';
    startBound.style.top = '0px';
    startBound.style.width = '' + startSize + 'px';
    startBound.style.left = '0px';
    document.getElementById('top-dna').appendChild(startBound);
 
    var endBound = document.createElement('div');
-   endBound.className = 'top-dna-boundary-marker';
+   endBound.className = 'top-dna-boundary-marker top-dna-boundary-marker-end';
    endBound.style.top = '0px';
    endBound.style.width = '' + endSize + 'px';
    endBound.style.right = '0px';

### 2012-11-28 ui/js/annotation.js
Exon ordering

@@ -292,6 +292,7 @@ function createExonElement(start, end, key) {
    var exonElement = document.createElement('div');
    exonElement.id = 'exon-' + key;
    exonElement.classList.add('exon');
+   exonElement.setAttribute('data-start', start);
 
    var exonElementString =
          "<span>Begin: </span>" +
@@ -305,7 +306,19 @@ function createExonElement(start, end, key) {
          "<button onclick='removeExon(" + key + ");'>Remove Exon</button>";
 
    exonElement.innerHTML = exonElementString;
-   document.getElementById('exons').appendChild(exonElement);
+
+   // Find the correct place to insert the node.
+   // This is the exon that belongs after this node.
+   // If it remains null, the new exon will be inserted at the end.
+   var afterNode = null;
+   $('.exon').each(function() {
+      if (parseInt(this.getAttribute('data-start')) > start) {
+         afterNode = this;
+         // This breaks.
+         return false;
+      }
+   });
+   document.getElementById('exons').insertBefore(exonElement, afterNode);
 }
 
 function updateExon(key) {

### 2012-11-28 ui/js/annotation.js
Rounded number on ruler

@@ -125,9 +125,6 @@ document.addEventListener('DOMContentLoaded', function () {
 
    // Change the title of the page with the gene name.
    document.getElementById('annotation-name').addEventListener('change', function() {
-      //TEST
-      console.log("THIS");
-
       var newName = document.getElementById('annotation-name').value;
       window.cgat.geneName = newName;
       document.title = 'Annotate: ' + newName;
@@ -203,9 +200,6 @@ function updateBoundingMarkers() {
    var startSize = Math.floor(window.cgat.geneStart * sizeRatio);
    var endSize = Math.floor((window.cgat.dna.length - window.cgat.geneEnd) * sizeRatio);
 
-   console.log("len: " + window.cgat.dna.length + ", geneEnd: " + window.cgat.geneEnd);
-   console.log("Start: " + startSize + ", End: " + endSize);
-
    var startBound = document.createElement('div');
    startBound.className = 'top-dna-boundary-marker top-dna-boundary-marker-begin';
    startBound.style.top = '0px';
@@ -222,7 +216,7 @@ function updateBoundingMarkers() {
 }
 
 function fillRuler(length) {
-   var tick = Math.floor(length / 10);
+   var tick = highestOrderIntRound(Math.floor(length / 10));
    var ruler = [];
 
    for (var i = 1; i <= 10; i++) {
@@ -508,7 +502,6 @@ function collectAnnotationData() {
 // Validate any required information before a submit.
 function validateBeforeSubmit() {
    // The only thing I can think of here is the gene name.
-   console.log($('#annotation-name').val());
    if ($('#annotation-name').val().length === 0) {
       validationError('Must have a Gene Name', 'gene-name-area');
       return false;

### 2012-11-28 Assignment6/paper/paper.tex
Modified quite  a bit of the paper's evaluation and conclusion sections.
Also updated the SQLDiagram eps to reflect the newest SQLDiagram png file,
which removed the Collab entities

--- 2012-11-27 Assignment6/paper/paper.tex
@@
 
-Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
-Obispo, offers a bioinformatics course that covers several aspects of gene
-annotation. This project is spearheaded by Ryan Verdon to accommodate Dr.
-Goodman's needs and ideas for ideal gene annotation software.
+%Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
+%Obispo, offers a bioinformatics course that covers several aspects of gene
+%annotation. This project is spearheaded by Ryan Verdon to accommodate Dr.
+%Goodman's needs and ideas for ideal gene annotation software.
 
-This paper primarily focuses on considerations for \textit{CGAT}'s database
+This paper primarily focuses on considerations for \textit{CGAT}s database
 architecture. These considerations are made with regards to a MySQL cluster
@@
 Traditional web applications are almost always backed by a relational database.
-However, there has been a recent movement towards using NoSQL databases for
-web application. Our main goal is to determine whether a NoSQL database, Couchbase,
+However, there has been a recent movement towards the use of NoSQL databases for
+web applications. Our main goal is to determine whether a NoSQL database, Couchbase,
 can outperform a traditional relational database, MySQL, for \textit{CGAT}'s
@@
 \section{Experiment Design}\label{sec:design}
-To evaluate the performance we auto-generated around
-400MB of data and inserted the same data into each cluster. Obviously using a
-different data model for Couchbase to accommodate for the fact it is a NoSQL
-key-value store instead of a relational database. We then created a set of six
-experiments to run. Five came from common use cases found in the web
-application. The final experiment was a contrived one designed to test the performance on each
-cluster by repeatedly writing and reading from the same object. Our major
-metrics are write performance, read performance, and time-delay between writes
-and reads of the same data.
+To evaluate the performance of couchbase and MySQL, we generated around 400MB
+of data that was used by each database\footnote{Different data models were used
+for the data which are depicted in the appendices at the end of this paper.}.
+We then created a set of six experiments to run. Five came from common use
+cases found in the web application. The final experiment was a contrived one
+designed to test the performance on each cluster by repeatedly writing and
+reading from the same object. Our major metrics are write performance, read
+performance, and time-delay between writes and reads of the same data.
 
-For our experiments we are using a total of five EC2 instances from the Amazon
-Web Services cloud offering. For both of our systems four servers were used for the
-database cluster while the last was used for running our workload harness. The MySQL
-system is setup with one master server for writes and three slaves servers that
-replicate all of the masters data. Unlike the Couchbase system where each
-server handled a subset of the data.
-
-Our major driving force of this experiment was to see how Couchbase compared to
-MySQL for running a web application. To test which is better we created a
-workload harness. The main tasks of the workload harness was to define how many
-times a workload would run and which workloads to run. In our case workloads
-define a certain task. For example, getting of the data necessary to display a
-user's profile. In our system a workload knows how to do its job in both MySQL
-and the Couchbase API. We leave describing each type workload we ran to the
-next section.
-
-As each workload is run the workload harness records important stats. Some of
-the stats include how long the workload took to run, the min and max times of
-an individual query or insertion, and the standard deviations of individual
-runs.
+For our experiments, we are using a total of five EC2 instances from the Amazon
+Web Services cloud offering. For both of our systems, four servers were used
+for the database cluster while the fifth was used for running our workload
+harness. The MySQL system is setup with one master server for writes and three
+slaves servers that replicate all of the masters data. Unlike the Couchbase
+system where each server handled a subset of the data.  The main tasks of the
+workload harness was to define how many times a workload would run and which
+workloads to run. In our case, workloads represent a certain task or use case
+in the \textit{CGAT} system. For example, retrieving data that is displayed on
+a user's profile. In our system, a workload knows how to do its job in both
+MySQL and the Couchbase API.  As each workload is run the workload harness
+records important statistics.  These include the length of time the workload
+took to complete, the min and max completion times of an individual query or
+insertion, and the standard deviations of individual runs.
 
 \subsection{Data}
-There exists four main pieces of data that are crucial to either
-representation: contigs, annotations, users, and groups. Contigs are
-what a biologist annotates to find genes. The main portion of a contig is a DNA
-sequence. Annotations are the record or formal representation of a gene that
-has been found in a specific contig. An annotation has several parts including
-coordinates in the contig which specify individual parts of the gene, exons, and meta information
-about the annotation like who uploaded it and when it was done, etc.
+Our data models have four crucial pieces of information: contigs, annotations,
+users, and groups. Contigs and annotations are biological in nature, while
+users and groups are particularly defined for our system.
+
+When biologists sequence genomes, they sequence them in overlapping segments
+called \textit{contigs}. A contig, which is a DNA sequence of some length, contains some
+number of genes. Biologists record several pieces of information, called
+\textit{annotations}, about genes including coordinates of components of the gene, such
+as exons, and meta information about the annotation like who uploaded it and
+when it was done, etc.
 
 Groups are another important part of the usefulness of the system. Users of
-like minds or research targets can join the same group to receive important information. Users is a
-major part of the data and is one of the central pillars that connects the
-data. Users upload contigs, create annotations, and join groups. Below we show
-the different ways we store the data in SQL and JSON. We chose to represent all of our
-data in Couchbase in JSON because it is a naturally serializable data format with
-large library support. In addition, using JSON would allow use to share the same
-schema with other popular document-based NoSQL databases, such as MongoDB.
+like minds or research interests can join the same group to receive important
+information. Users are a major part of the data and serve as the central hub that connects the
+data. Users upload contigs, create annotations, and join groups.
 
 \subsubsection{SQL}
-In SQL the data is represented as nine tables. A diagram of the tables with
-links to represent foreign key constraints is shown in Figure \ref{fig:SQLDiagram}.
+Our data model in SQL consists of nine relational tables depicted in Figure
+\ref{fig:SQLDiagram} in the appendix of this paper.
+%GroupMembership and Tasks
+%represent many-to-many relationships. GroupMembership represents the
+%relationship between Users and Groups while Tasks links Users and Contigs.
+%Collab entities--CollabAnnotations and CollabExons--are not implemented but
+%represent the end goal of having a collaboratively constructed aggregation of
+%annotations and exons.
 
 \subsubsection{JSON}
-For Couchbase we used four JSON objects which are shown in Figures \ref{fig:JSON-user}-\ref{fig:JSON-group}.
-All of the information from the SQL tables is represented in these JSON
-objects. We take advantage of nested objects to reduce as much subsequent
-requerying of the database as possible.
+We chose to represent all of our data in Couchbase in JSON because it is a
+naturally serializable data format with large library support. In addition,
+using JSON would allow use to share the same schema with other popular
+document-based NoSQL databases, such as MongoDB.  We used four JSON objects
+which are shown in Figures \ref{fig:JSON-contig}, \ref{fig:JSON-annotation},
+\ref{fig:JSON-user}, and \ref{fig:JSON-group} in Appendix
+\ref{sec:data_models}. All of the information from the SQL tables is
+represented in these JSON objects. We take advantage of nested objects to
+reduce as much subsequent requerying of the database as possible.
 
-\begin{figure}
-	\includegraphics[width=5.5in]{SQLDiagram.eps}
-	\caption{The SQL diagram for CGAT.}
-	\label{fig:SQLDiagram}
-\end{figure}
-
-%\clearpage
-
-\begin{figure}%
-	\includegraphics{user.eps}
-	\caption{The JSON for a user.}
-	\label{fig:JSON-user}
-\end{figure}%
-
-\begin{figure}%
-	\includegraphics[height=70mm]{contig.eps}
-  \caption{The JSON for a contig.}
-  \label{fig:JSON-contig}
-\end{figure}%
-
-\begin{figure}%
-	\includegraphics[height=100mm]{annotation.eps}
-	\caption{The JSON for an annotation.}
-	\label{fig:JSON-annotation}
-\end{figure}%
-
-\begin{figure}%
-	\includegraphics[height=20mm]{group.eps}
-	\caption{The JSON for a group.}
-	\label{fig:JSON-group}
-\end{figure}%
-
-\clearpage
+\begin{table}[t]
+   \centering
+      \begin{tabular}{| l | c | c | c |}
+         \hline
+            {\textbf{Workload}} &
+            {\textbf{MySQL}} &
+            {\textbf{Couchbase}} \\
+         \hline
+            Task Assignment         & 39325189ms & 157431049ms\footnotemark{} \\
+         \hline
+            Profile View            & 2908257ms & 7008818ms \\
+         \hline
+            Annotation Publication  & 510837ms & 1830064ms \\
+         \hline
+            Group Modification      & 316673ms & 9086304ms \\
+         \hline
+            Contig Upload           & 119892ms & 20133ms \\
+         \hline
+            Read/Write              & 103808ms & 58581ms \\
+         \hline
+      \end{tabular}
+      \caption{Comparison of runtimes in ms for workloads using MySQL and Couchbase.
+               Each workload was run for 100000 iterations.}
+      \label{tab:workload_perf}
+\end{table}
+%footnote text has to be separate from the mark because the table is floating
+%and so if you just use the usual '\footnote{}' then the mark shows up in the
+%document but not the text. awkies...
+\footnotetext{Task assignment never actually completed. The number reported is the
+approximate length of time spent running the workload before
+giving up. Explanation provided in section \ref{sec:eval}}
 
@@
 during our comparison was determined by the number of members in our group (as
-specified by Alex Dekhtyar). Unfortunately due to time constraints we were
+specified by Alex Dekhtyar). Unfortunately, due to time constraints we were
 unable to do multiple runs of any one workload, however, due to our environment
@@
 
-\subsection{Group Membership Modification}
+\subsection{Group Modification}
 Group membership is likely the second most stable of the workloads used in our
@@
 \section{Implementation}\label{sec:implementation}
-Our experiment was implemented completely in Java. To do the MySQL portion of
+Our experiment was implemented in Java. To do the MySQL portion of
 the workload we used the MySQL JDBC driver. For the Couchbase portion we used
 the official Couchbase Java library from their website.
+%TODO lol whut y so short
+
+\begin{table}[t]
+   \centering
+      \begin{tabular}{| l | c | c | c |}
+         \hline
+            \multicolumn{4}{|c|}{{\textbf{MySQL read/write performance}}} \\
+         \hline
+            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
+         \hline
+            Min (ms) & 0 & 0 & 0 \\
+         \hline
+            Max (ms) & 25 & 70 & 70 \\
+         \hline
+            Mean (ms) & 0.473 & 0.558 & 0.516 \\
+         \hline
+            Median (ms) & 0 & 1 & 0 \\
+         \hline
+            Standard Deviation (ms) & 0.685 & 0.859 & 0.778 \\
+         \hline
+      \end{tabular}
+      \caption{Comparison of reads and writes to the same object in MySQL.}
+      \label{tab:sql_perf}
+\end{table}
+
+\begin{table}[t]
+   \centering
+      \begin{tabular}{| l | c | c | c |}
+         \hline
+            \multicolumn{4}{|c|}{{\textbf{Couchbase read/write performance}}} \\
+         \hline
+            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
+         \hline
+            Min (ms) & 0 & 0 & 0 \\
+         \hline
+            Max (ms) & 71 & 48 & 71 \\
+         \hline
+            Mean (ms) & 0.562 & 0.017 & 0.290 \\
+         \hline
+            Median (ms) & 1 & 0 & 0 \\
+         \hline
+            Standard Deviation (ms) & 0.859 & 0.213 & 0.682 \\
+         \hline
+      \end{tabular}
+      \caption{Comparison of reads and writes to the same object in couchbase}
+      \label{tab:couch_perf}
+\end{table}
 
 \section{Evaluation}\label{sec:eval}
-In this section we discuss the results from each workload in turn.
+In this section we discuss the results for each workload in turn. As a general
+note, each workload included a specific task that was done $N$ number of times.
+While we initially planned for $N$ to be the same for each workload
+($N=100000$), the ContigUpload workload ran into memory constraints of the
+EC2 machine running our queries. This caused us to reduce the number of times
+the ContigUpload task was repeated ($N=10000$) for the ContigUpload workload.
+Additionally, the TaskAssignment workload never completed on couchbase. The
+runtime reported for this workload was the amount of time the couchbase tests
+were running with the times for ContigUpload and ProfileView workloads
+subtracted. We had waited approximately 45 hours (from Friday night to Sunday
+evening) until finally the workloads had been killed without reporting a time
+for the TaskAssignment workload.
 
-%section on each workload and the results on each system (could be a table tbh and not a graph)
-%probably some weak analysis on why we expected the results
-%need to ask eriq the # of times each workload was  run, might be in an email?
+\subsection{Task Assignment}
+From table \ref{tab:workload_perf}, it seems that MySQL significantly
+outperforms Couchbase for the TaskAssignment workload. We think this is due to
+the read-heavy nature of this workload which is made even more read-intensive
+given Couchbase's feature set and data model. Where MySQL's data model allows
+simple ad-hoc query to find necessary values prior to inserting the new task,
+Couchbase must fetch entire JSON documents from the database and it must do
+this for several documents for each task being assigned. We believe if
+Couchbase offered features for document or ad-hoc unstructured queries the
+performance of this workload would be significantly better. Particularly
+because this workload only requires a small subset of the JSON documents to be
+used, querying for only the necessary parts of each JSON document would make a
+big difference in query completion time.
 
-\subsection{Assigning a Task}
+\subsection{Profile View}
+Similarly to the TaskAssignment workload, MySQL outperforms Couchbase for the
+ProfileView workload. Likely, this is for the same reasons that the
+TaskAssignment workload has poor performance on Couchbase. This workload could
+be improved in Couchbase with the use of unstructured queries, preventing the
+need to retrieve several complete documents for each run.
 
-\subsection{Profiling}
+\subsection{Annotation Publication}
+Again, a similar pattern in the results as the last two workloads. This was
+another workload that required some queries prior to inserting a new
+annotation. We believe this is another case where unstructured queries could
+vastly improve performance.
 
-\subsection{Publishing}
+\subsection{Group Modification}
+To our surprise this workload performed extremely poorly on Couchbase. We were not
+expecting this because this is a 50\% read and 50\% write workload. Having to
+do lots of reads prior to inserting new group memberships really hurt
+performance for Couchbase. This is another scenario in which having unstructured queries or
+document support would improve performance given our data model. We store group
+memberships as a sub-field in the user and group objects. Querying only
+these fields could reduce the amount of data we return to the workload
+harness by over 70\%. Given more time, it would be interesting to see if we
+could better adapt our data model for Couchbase's querying capabilities.
+Intuitively, it would seem that perhaps as a key-value store it performs better if we store group
+memberships as a separate document from the user and group documents, however,
+if we do this we only increase the amount of hits to the database we would have
+to make. It seems that Couchbase simply does not lend itself well to highly
+related or inter-connected data.
 
-\subsection{Adding and Removing Group Membership}
+\subsection{Contig Uploads}
+The results from this test were no surprise, the ContigUpload workload is
+entirely write-heavy and so Couchbase performance demolishes MySQL performance.
+Even with contig objects containing strings of over 50k characters, by not
+having to lock key-value pairs Couchbase can achieve great write performance.
 
-\subsection{Uploading Contigs}
+\subsection{Read/Write Performance}
+The results from this test were unsurprising based on the results from our
+other tests, especially the GroupModification workload. Our workloads seem to
+show that Couchbase is much faster in terms of writes when compared to MySQL.
+Additionally, Couchbase can leverage Memcache in this workload to speed up
+queries. For this workload, we decided to gather more statistics to see how the
+database performs. We expanded our statistics gathering to include the minimum,
+maximum, mean, median and standard deviation of the individual reads and writes
+for both Couchbase and MySQL. Table \ref{tab:sql_perf} contains our SQL
+results, while table \ref{tab:couch_perf} contains equivalent results for
+Couchbase. Both tables list the minimum, maximum, mean, median, and standard
+deviation for various types of database access patterns. The types of database
+access patterns we were interested in are reads after writes of the same object
+and writes after reads of the same object.
 
-\subsection{Read and Writing Performance}
+Examining the read after writes results from the two systems, we observe
+expected patterns in times given the results of our other workloads,
+AssignTask, ProfileView, AnnotationPublication, GroupModification, and
+ContigUpload. MySQL consistently outperforms Couchbase when reading an object
+after it has written to that object. In contrast, Couchbase greatly outperforms
+MySQL in write after read access patterns of the data as shown by the
+significant difference between the mean query completion time for MySQL and
+Couchbase. This coincides and seems to agree with our results from write-heavy
+workloads where Couchbase's performance showed improvement over MySQL.
 
 \section{Conclusions}\label{sec:conclusions}
-TODO
+From our results, MySQL appears to provide better performance for the types of
+workloads we expect \textit{CGAT} to primarly encounter. These workloads will
+likely be more read-heavy than write-heavy due to the idea that users will
+often consume information at a higher rate than they will be submitting
+information. Even if this may not be the case for any particular user, as
+\textit{CGAT} expands and gains more users, the amount of users consuming
+information is likely to outpace the amount of users submitting information.
+
+One thing to note, the workloads we prepared for our comparison of
+MySQL and Couchbase do not allow much opportunity for Couchbase to leverage
+Memcache. Our workloads were setup to select random data objects to work on and
+with the way we executed our workloads, the probability that the data being
+retrieved was present in Memcache was relatively low compared to live
+workloads. If Memcache could cache data such that live workloads would
+benefit from cache hits, it would be possible to offset the difference in read
+performance between Couchbase and MySQL.
+
+In the end, it's possible that Couchbase may not perform as badly if we could
+optimize for it, but the amount of effort required to achieve our current
+performance statistics was much less for MySQL than for Couchbase.
+Additionally, the amount of effort required to learn how to optimize our data
+model for Couchbase's querying capabilities and for Memcache to maximize cache
+hits may be more than desired. Especially if MySQL is able to perform better
+for a more dominant data access pattern. While MySQL seems to be a clear choice
+for us (at least compared to Couchbase) the clear takeaway is that our data is
+simply more relational than Couchbase was developed to support.
+
+%We believe that Couchbase performance would greatly improve with the addition of an unstructured
+%query language, though it is possible that Couchbase targets applications that
+%often do not have structured document objects as we do for \textit{CGAT}. If it
+%is possible to add this kind of query language to Couchbase with a low cost, it
+%would be a good way to minimize the amount of data retrieved from the database
+%for workloads similar to ours.
+
+\onecolumn
+\appendices
+
+\section{SQL and JSON data models}\label{sec:data_models}
+\centering
+Here are diagrams depicting the SQL and JSON variants of our data model.
+
+\begin{figure*}[t]
+   \centering
+   \includegraphics[height=90mm]{SQLDiagram.eps}
+	\caption{Entity Relationship diagram for CGAT.}
+	\label{fig:SQLDiagram}
+\end{figure*}
+
+\begin{figure*}[t]
+   \centering
+   \subfloat[Contig format]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
+   \subfloat[Annotation format]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
+   \caption{JSON format for Contigs and Annotations.}
+\end{figure*}
+
+\begin{figure*}[t]
+   \centering
+   \subfloat[User format]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
+   \subfloat[Group format]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
+   \caption{JSON format for Users and Groups.}
+\end{figure*}
+
+%\begin{figure*}
+%  \includegraphics[height=90mm]{user.eps}
+%	\caption{The JSON for a user.}
+%	\label{fig:JSON-user}
+%\end{figure*}
+%\begin{figure*}
+%  \includegraphics[height=25mm]{group.eps}
+%	\caption{The JSON for a group.}
+%	\label{fig:JSON-group}
+%\end{figure*}
+%\begin{figure*}
+%  \includegraphics[height=80mm]{contig.eps}
+%  \caption{The JSON for a contig.}
+%  \label{fig:JSON-contig}
+%\end{figure*}
+%\begin{figure*}
+%  \includegraphics[height=90mm]{annotation.eps}
+%	\caption{The JSON for an annotation.}
+%	\label{fig:JSON-annotation}
+%\end{figure*}
 

### 2012-11-30 ui/js/script.js
Help page template changes

@@ -147,6 +147,13 @@ function makeContig(contig) {
    return contigInfo;
 }
 
+// Generate some html to represent the help info (given the standard information from the api).
+function makeHelp(help) {
+   var helpInfo = '';
+   helpInfo += "<p>" + "hi" + "</p>";
+   return helpInfo;
+}
+
 // Generate some html meant to diplay an inline annotation given stantard api annotation info.
 function makeInlineAnnotation(annotation) {
    var numExons = annotation.exons ? annotation.exons.length : 0;

### 2012-11-30 ui/js/annotation.js
Gene Exon View

@@ -182,7 +182,7 @@ function startEndValidate(start, end, errorDump) {
                       errorDump);
       return false;
    } else if (start >= end) {
-      validationError('Start must be < End', errorDump);
+      validationError('Begin must be < End', errorDump);
       return false;
    }
 
@@ -395,6 +395,42 @@ function placeExons() {
       // TODO(eriq): Append all the kids at the same time.
       document.getElementById('top-dna').appendChild(marker);
    }
+
+   placeExonDiagram();
+}
+
+function placeExonDiagram() {
+   var html = '';
+
+   var count = 0;
+   sortExons(window.cgat.exons).forEach(function(exon) {
+      html += "<div class='gene-diagram-exon gene-diagram-element'>" +
+              "<p class='gene-diagram-exon-begin'>" + exon.start + "</p>" +
+              "<p class='gene-diagram-exon-end'>" + exon.end + "</p>" +
+              "</div>";
+
+      count++;
+
+      if (count != window.cgat.exons.length) {
+         html += "<div class='gene-diagram-element gene-diagram-intron'>" +
+                 "<div class='gene-diagram-intron-line'></div>" +
+                 "</div>";
+      }
+   });
+
+   $('#gene-diagram').html(html);
+}
+
+function sortExons(exons) {
+   var sorted = [];
+
+   for (var exonKey in exons) {
+      sorted.push(exons[exonKey]);
+   }
+
+   return sorted.sort(function(a, b) {
+      return a.start - b.start;
+   });
 }
 
 // markPosition is meant to mark the first nucleotide in each line.

### 2012-11-30 ui/js/annotation.js
Setting the subtitle now will change the actual document title. This takes care of our current needs, but may need to be changed later to only optionally change the true title.

@@ -75,7 +75,6 @@ document.addEventListener('DOMContentLoaded', function () {
 
          // Fill in the gene name
          window.cgat.geneName = data.annotation.isoform_name || '';
-         document.title = 'Annotate: ' + window.cgat.geneName;
          setSubtitle(window.cgat.geneName);
          simpleBoxSelectValue(0, window.cgat.geneName, window.cgat.geneName);
 
@@ -127,7 +126,6 @@ document.addEventListener('DOMContentLoaded', function () {
    document.getElementById('annotation-name').addEventListener('change', function() {
       var newName = document.getElementById('annotation-name').value;
       window.cgat.geneName = newName;
-      document.title = 'Annotate: ' + newName;
       setSubtitle(newName);
    });
 

### 2012-12-06 paper/paper.tex
Paper base
Removed unused sections

--- 2012-11-28 Assignment6/paper/paper.tex
@@
-% Paper for the ChouchBase assignment.
+% Paper for CGAT
 \documentclass[10pt, conference, compsocconf]{IEEEtran}
@@
 
-\title{No Dogs On The Couch$_{Base}$}
+\title{CGAT In The Hat}
 
@@
 
-Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
-Obispo, offers a bioinformatics course that covers several aspects of gene
-annotation. This project is spearheaded by Ryan Verdon to accommodate Dr.
-Goodman's needs and ideas for ideal gene annotation software.
+%Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
+%Obispo, offers a bioinformatics course that covers several aspects of gene
+%annotation. This project is spearheaded by Ryan Verdon to accommodate Dr.
+%Goodman's needs and ideas for ideal gene annotation software.
 
@@
 
-\section{Experiment Design}\label{sec:design}
-To evaluate the performance we auto-generated around
-400MB of data and inserted the same data into each cluster. Obviously using a
-different data model for Couchbase to accommodate for the fact it is a NoSQL
-key-value store instead of a relational database. We then created a set of six
-experiments to run. Five came from common use cases found in the web
-application. The final experiment was a contrived one designed to test the performance on each
-cluster by repeatedly writing and reading from the same object. Our major
-metrics are write performance, read performance, and time-delay between writes
-and reads of the same data.
+\onecolumn
+\appendices
 
-For our experiments we are using a total of five EC2 instances from the Amazon
-Web Services cloud offering. For both of our systems four servers were used for the
-database cluster while the last was used for running our workload harness. The MySQL
-system is setup with one master server for writes and three slaves servers that
-replicate all of the masters data. Unlike the Couchbase system where each
-server handled a subset of the data.
+\section{JSON Data Models}\label{sec:data_models}
+\centering
+Here are diagrams depicting our JSON data model.
 
-Our major driving force of this experiment was to see how Couchbase compared to
-MySQL for running a web application. To test which is better we created a
-workload harness. The main tasks of the workload harness was to define how many
-times a workload would run and which workloads to run. In our case workloads
-define a certain task. For example, getting of the data necessary to display a
-user's profile. In our system a workload knows how to do its job in both MySQL
-and the Couchbase API. We leave describing each type workload we ran to the
-next section.
-
-As each workload is run the workload harness records important stats. Some of
-the stats include how long the workload took to run, the min and max times of
-an individual query or insertion, and the standard deviations of individual
-runs.
-
-\subsection{Data}
-There exists four main pieces of data that are crucial to either
-representation: contigs, annotations, users, and groups. Contigs are
-what a biologist annotates to find genes. The main portion of a contig is a DNA
-sequence. Annotations are the record or formal representation of a gene that
-has been found in a specific contig. An annotation has several parts including
-coordinates in the contig which specify individual parts of the gene, exons, and meta information
-about the annotation like who uploaded it and when it was done, etc.
-
-Groups are another important part of the usefulness of the system. Users of
-like minds or research targets can join the same group to receive important information. Users is a
-major part of the data and is one of the central pillars that connects the
-data. Users upload contigs, create annotations, and join groups. Below we show
-the different ways we store the data in SQL and JSON. We chose to represent all of our
-data in Couchbase in JSON because it is a naturally serializable data format with
-large library support. In addition, using JSON would allow use to share the same
-schema with other popular document-based NoSQL databases, such as MongoDB.
-
-\subsubsection{SQL}
-In SQL the data is represented as nine tables. A diagram of the tables with
-links to represent foreign key constraints is shown in Figure \ref{fig:SQLDiagram}.
-
-\subsubsection{JSON}
-For Couchbase we used four JSON objects which are shown in Figures \ref{fig:JSON-user}-\ref{fig:JSON-group}.
-All of the information from the SQL tables is represented in these JSON
-objects. We take advantage of nested objects to reduce as much subsequent
-requerying of the database as possible.
-
-\begin{table}[t]
+\begin{figure*}[t]
    \centering
-      \begin{tabular}{| l | c | c | c |}
-         \hline
-            {\textbf{Workload}} &
-            {\textbf{MySQL}} &
-            {\textbf{Couchbase}} \\
-         \hline
-            Task Assignment         & 39325189ms & 157431049ms \\
-         \hline
-            Profile View            & 2908257ms & 7008818ms \\
-         \hline
-            Annotation Publication  & 510837ms & 1830064ms \\
-         \hline
-            Group Modification      & 316673ms & 9086304ms \\
-         \hline
-            Contig Upload           & 119892ms & 20133ms \\
-         \hline
-            Read/Write              & 103808ms & 58581ms \\
-         \hline
-      \end{tabular}
-      \caption{Comparison of runtimes in ms for workloads using MySQL and Couchbase.
-               Each workload was run for 100000 iterations.}
-      \label{tab:workload_perf}
-\end{table}
-
-\section{Workloads}\label{sec:workload}
-The core unit of work that we used to compare Couchbase performance against
-MySQL performance was called a \textit{workload}. Each workload was designed
-either to be write-heavy, read-heavy, or balanced. The number of workloads used
-during our comparison was determined by the number of members in our group (as
-specified by Alex Dekhtyar). Unfortunately due to time constraints we were
-unable to do multiple runs of any one workload, however, due to our environment
-and machines, there should be minimal fluctuations in any given run time for a
-workload.
-
-In this section we further describe each task's purpose as well as what type of
-workload it was (write-heavy, etc.). This is necessary to be able to properly
-interpret (and agree or disagree with) the results described in Section
-\ref{sec:eval}.
-
-\subsection{Profile View}
-The first thing that all users will do in \textit{CGAT} (ideally) is log in.
-When a user logs into \textit{CGAT} the user will be directed to their profile
-view which will contain all groups they are a member of, all completed
-annotations they have submitted, partial annotations that have been saved, and
-all tasks they have assigned. This is a use case that is likely to see many
-additions in the future as it serves as the first and foremost view of the
-system for the user. This means that any information that should be accessible
--- including when a forum system is integrated into \textit{CGAT} -- will find
-a place on the user's profile.
-
-This workload is clearly a read-heavy workload that entails no disk writes at
-all. Important to note though, this use case may constantly be increasing
-(should constantly be increasing for a successful system) and so the load for
-this workload may increase quite a bit, and whenever new features are added
-there is a potential for a spike in workload.
-
-\subsection{Task Assignment}
-In \textit{CGAT}, it must be possible to request a group of users to annotate
-new or updated contigs. The frequency of this use case is dependent on the
-adoption of \textit{CGAT} -- the more species \textit{CGAT} supports, the more
-frequent this use case will be. More specifically, the relevant entities in the
-\textit{CGAT} schema are \textit{users}, \textit{contigs}, and \textit{groups}.
-Since a task embodies a request for a group $G$ to annotate a contig $C$, every
-user $U$ in $G$ receives a notification including a description of $C$ and a
-message requesting that it be annotated. A task can be assigned to any number
-of groups. Since tasks are assigned to groups that have some vested interest in
-the species or specific contig that has been updated or added, there may be
-many groups that a task should be assigned to. Additionally, as more species
-are supported by \textit{CGAT} or more users join \textit{CGAT}, more users
-will receive notifications of tasks and so there are multiple ways in which
-this use case must be scalable.
-
-This workload is write-heavy -- adding tasks for each user in a group requires
-adding many records to a relational table or modifying many documents/values
-in a key/value store.
-
-\subsection{Annotation Publication}
-Annotations are the core focus of \textit{CGAT}. Every user is going to be
-doing some annotation, and they should be doing many annotations. Especially
-since \textit{CGAT} defines an annotation to be information about a gene
-isoform, variation, on a contig, even annotating a single gene will yield to several
-annotations in \textit{CGAT}. Additionally, when an annotation is submitted,
-the annotation must be marked as complete, the user and annotation then are
-assigned a certain amount of experience--the user gains experience towards
-his total, the annotation gains experience and serves as a history of the
-user's experience.
-
-This is a mostly write-, somewhat read-heavy workload. It is write-heavy in the
-sense that it must update a user and the annotation that the user completed.
-However, it also has some reads in the sense that the contig, for which the
-annotation is being submitted, is used to determine the amount of experience
-that submitting the annotation yields for the user. However, given the
-difference in data model between Couchbase and MySQL, the Couchbase
-implementation is much more write-heavy than the SQL equivalent. This is
-because the Couchbase data model includes lists of complete and incomplete
-annotations for each user so that the annotation does not have to be fetched
-for every use case involving users and annotations. Since this is the case,
-more of the document undergoes modification for this use case.
-
-\subsection{Group Membership Modification}
-Group membership is likely the second most stable of the workloads used in our
-comparisons of Couchbase vs MySQL next to contig uploads. This is ``stable''
-because the majority of users will only join groups, and leaving groups will
-likely be relatively rare. This use case will only ever be as bad as how many
-users join at any one time--once a user registers for \textit{CGAT} and has
-joined some groups the user is unlikely to join or leave groups and so this use
-case poses almost no scalability problems. Although, despite this, there will
-be spikes that coincide with academic schedules, as groups will be created and
-populated for each new class that uses \textit{CGAT}.
-
-This workload is the most balanced of the workloads described in this paper
-(except maybe for ``Read/Write'' workload). Generally speaking, when a user joins a
-group then there is only the addition of a record in MySQL or the modification
-of the user and group document in Couchbase. And when a user leaves a group, then
-subsequently the additional relational tuple or field in the user document is
-``undone'' by removing the group membership tuple or deleting the field from
-the user and group document.
-
-\subsection{Contig Uploads}
-Contig uploads is the most stable workload when the manual, or user, use case
-is considered. Although, there will be large spikes of activity when/if the
-reference genome for a species is updated--requiring \textit{CGAT} to update
-any and all of the contigs relevant to the genome. However, reference genomes
-are updated very infrequently and contigs will never be removed from
-\textit{CGAT}.
-
-This workload is write-heavy because it requires no reads--the use case is for
-uploading data, not downloading data.
-
-\subsection{Read/Write Performance}
-This is a test workload to examine the performance of writing and reading from
-the same object in succession. This is a workload contrived purely for better
-comparison of Couchbase vs MySQL instead of workloads representative of what
-\textit{CGAT} will rely on. This is a naive, balanced workload.
-
-\section{Implementation}\label{sec:implementation}
-Our experiment was implemented completely in Java. To do the MySQL portion of
-the workload we used the MySQL JDBC driver. For the Couchbase portion we used
-the official Couchbase Java library from their website.
-
-\begin{table}[t]
-   \centering
-      \begin{tabular}{| l | c | c | c |}
-         \hline
-            \multicolumn{4}{|c|}{{\textbf{MySQL read/write performance}}} \\
-         \hline
-            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
-         \hline
-            Min (ms) & 0 & 0 & 0 \\
-         \hline
-            Max (ms) & 25 & 70 & 70 \\
-         \hline
-            Mean (ms) & 0.473 & 0.558 & 0.516 \\
-         \hline
-            Median (ms) & 0 & 1 & 0 \\
-         \hline
-            Standard Deviation (ms) & 0.685 & 0.859 & 0.778 \\
-         \hline
-      \end{tabular}
-      \caption{Comparison of reads and writes to the same object in MySQL.}
-      \label{tab:sql_perf}
-\end{table}
-
-\begin{table}[t]
-   \centering
-      \begin{tabular}{| l | c | c | c |}
-         \hline
-            \multicolumn{4}{|c|}{{\textbf{Couchbase read/write performance}}} \\
-         \hline
-            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
-         \hline
-            Min (ms) & 0 & 0 & 0 \\
-         \hline
-            Max (ms) & 71 & 48 & 71 \\
-         \hline
-            Mean (ms) & 0.562 & 0.017 & 0.290 \\
-         \hline
-            Median (ms) & 1 & 0 & 0 \\
-         \hline
-            Standard Deviation (ms) & 0.859 & 0.213 & 0.682 \\
-         \hline
-      \end{tabular}
-      \caption{Comparison of reads and writes to the same object in couchbase}
-      \label{tab:couch_perf}
-\end{table}
-
-\section{Evaluation}\label{sec:eval}
-In this section we discuss the results from each workload in turn.
-
-\subsection{Task Assignment}
-From the results it is obvious that MySQL performs much better than Couchbase.
-The result we have for Couchbase is actually a minimum runtime. It was running
-so long we killed it to get more results. Even with this in mind, MySQL vastly
-outperforms Couchbase. We believe this is because during the process of
-assigning a task a couple values have to be read into memory before a finally
-write can be done. Compared to MySQL where its a simple ad-hoc query to find
-the necessary values prior to inserting the new task.
-
-From our analysis, we believe if we had document or ad-hoc unstructured queries
-in Couchbase we could vastly improve the performance.  When doing the Couchbase
-queries we only need a small subset of the JSON. By only querying for the
-necessary parts it could save vast amounts of time.
-
-\subsection{Profile View}
-Like the last workload, MySQL outperforms Couchbase. We believe that this is
-because this workload is read-heavy so several full documents are retrieved per
-run.  Again, this workload could be improved in Couchbase with the use of
-unstructured queries.
-
-\subsection{Annotation Publication}
-Again, a similar pattern in the results as the last two workloads. This was
-another workload that required some queries prior to inserting a new
-annotation. We believe this is another case where unstructured queries could
-vastly improve performance.
-
-\subsection{Group Modification}
-To our surprise this workload performed extremely bad on Couchbase. We were not
-expecting this because this is a 50\% read and 50\% write workload. Having to
-do lots of reads prior to inserting new group memberships really hurt
-Couchbase. This is most definitely a case where having unstructured queries or
-document support would greatly improve performance. The way we store group
-memberships is as a sub-field in the user and group objects. By only queries
-for those fields we could reduce the amount of data we return to the workload
-harness by over 70\%.
-
-\subsection{Contig Uploads}
-The results from this test were no surprise. Couchbase demolished MySQL when
-most of the workload was insertions. Even with contig objects containing
-strings of over 50k characters Couchbase performed amazingly well. By not
-having to lock key-value pairs Couchbase can get some awesome write
-performance.
-
-\subsection{Read/Write Performance}
-The results from this test were not that surprising based on the results from
-our other tests. It is obvious that Couchbase is much faster in terms of writes
-plus Couchbase can leverage Memcache in this workload to speed up queries to
-the database. For this workload we decided to gather a bunch more numbers to
-see what the minimum, maximum, mean, median and standard deviation of the
-individual reads and writes for each system. Below we list the six tables and
-discuss what the results mean.
-
-The three prior tables are all of our SQL results. The first table lists the
-minimum, maximum, mean, median, and standard deviation of reads that occur
-after writes. The second table lists the same metrics but for writes after
-reads. Lastly, the third table lists the metrics for all of the queries. Below
-we list the same three tables but this time show the Couchbase results.
-
-By examining the read after writes results from the two systems we get the
-results we expect based on our other tests. MySQL outperforms Couchbase when it
-comes to reading after a write. The really interesting results come from the
-the write after read tables where it is obvious that Couchbase demolished MySQL
-by a lot. There is a huge difference in the means alone without even looking at
-the other metrics. This again backs up our other results from the contig
-workload where Couchbase outperformed MySQL in a mainly write task. Overall
-Couchbase outperformed MySQL because of its blazing fast writes after reads.
-
-\section{Conclusions}\label{sec:conclusions}
-From our results MySQL appears to be the winner right now for the CGAT backend.
-The MySQL workloads outperformed Couchbase the majority of the time. We believe
-in a live scenario Couchbase would perform better than our tests mainly because
-our tests could not take advantage of Memcache. When a key had to be looked up
-there was almost no chance it was sitting in Memcache. We also believe that the
-Couchbase results would greatly improve with the addition of an unstructured
-query language. With the addition of this feature our workloads would be making
-much smaller requests from the database.
-
-% TODO: Uncomment if we use any refs
-% \bibliographystyle{acm}
-% \bibliography{refs}
-
-\newpage
-\appendices
-\begin{figure*}
-   \centering
-   \includegraphics[height=90mm]{SQLDiagram.eps}
-	\caption{The SQL diagram for CGAT.}
-	\label{fig:SQLDiagram}
+   \subfloat[Contig format]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
+   \subfloat[Annotation format]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
+   \caption{JSON format for Contigs and Annotations.}
 \end{figure*}
 
-\begin{figure*}{t}
+\begin{figure*}[t]
    \centering
-   \subfloat[Contig JSON]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
-   \subfloat[Annotation JSON]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
-\end{figure*}
-
-\begin{figure*}{t}
-   \centering
-   \subfloat[User JSON]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
-   \subfloat[Group JSON]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
+   \subfloat[User format]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
+   \subfloat[Group format]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
+   \caption{JSON format for Users and Groups.}
 \end{figure*}
@@
 
+% TODO: Uncomment if we use any refs
+% \bibliographystyle{acm}
+% \bibliography{refs}
+
 \end{document}

### 2012-12-06 paper/paper.tex
Adding some section outlines
Fixing up intro/motivation. Added the last section we need to do (future work). Added the references in refs.bib.

@@
 Gene annotation is the process of associating metadata about a gene with the
-contig, a dna sequence, on which the gene resides. This metadata is necessary for conducting
+contig, a dna sequence, on which the gene resides.\cite{annotation} This metadata is necessary for conducting
 genomic research projects and analysis involving the contig's genomic sequence.
@@
 will likely never be replaced (UCSC genome browser, etc. are well
-established), it would be desirable to have a software system that better
+established\cite{ucscbrowser, ncbi}), it would be desirable to have a software system that better
 accommodates the visual and informational needs of gene annotation.
@@
 \textit{CGAT} is a web-based gene annotation application designed with
-usability, simplicity, and efficiency in mind. It is not a feature and data
-rich genome browser like the \textit{UCSC genome brower}. Nor is \textit{CGAT}
+usability, simplicity, and efficiency in mind. It is not a feature-rich and data-rich
+ genome browser like the \textit{UCSC genome brower}. Nor is \textit{CGAT}
 designed to be a replacement for other genome browsers in any aspect other than
@@
 
-%Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
-%Obispo, offers a bioinformatics course that covers several aspects of gene
-%annotation. This project is spearheaded by Ryan Verdon to accommodate Dr.
-%Goodman's needs and ideas for ideal gene annotation software.
-
-This paper primarily focuses on considerations for \textit{CGAT}s database
-architecture. These considerations are made with regards to a MySQL cluster
-versus a Couchbase cluster to determine which database architecture is most
-efficient for running \textit{CGAT}. Section \ref{sec:motive} discusses our
-motivation to compare Couchbase and MySQL in this paper, while section
-\ref{sec:design} talks about how we designed experimental workloads to use for
-our comparison of Couchbase vs MySQL. In section \ref{sec:workload} we
-describe the workloads and how they are relevant to \textit{CGAT}, some brief
-implementation details are outlined in section \ref{sec:implementation}, we
-evaluate our results in section \ref{sec:eval} and finally conclude in
-section \ref{sec:conclusions}.
+This paper discusses the progress that has been made this quarter and the future work that 
+still needs to be done. The rest of the paper is as follows. Section \ref{sec:motive} discusses the
+motivation for creating CGAT. Section \ref{sec:choice} explains why we chose MongoDB
+as our backend. Section \ref{sec:implementation} goes into the details of how
+our backend and frontend are implemented currently. Section \ref{sec:ui} explains the
+various features of our UI. Lastly, section \ref{sec:future} describes the future work still left to
+do.
 
 \section{Motivation}\label{sec:motive}
-Traditional web applications are almost always backed by a relational database.
-However, there has been a recent movement towards the use of NoSQL databases for
-web applications. Our main goal is to determine whether a NoSQL database, Couchbase,
-can outperform a traditional relational database, MySQL, for \textit{CGAT}'s
-common operations.
+Dr. Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
+Obispo, offers a bioinformatics course that covers several aspects of gene
+annotation. Her work is a part of the Genomics Education Partnership (GEP) at 
+Washington University in St. Louis.\cite{gep} The goal of GEP is to teach undergrad students
+in Biology how to annotate genes and other features in raw sequence data.
+The CGAT project is spearheaded by Ryan Verdon to accommodate Dr.
+Goodman's needs and ideas for ideal gene annotation software.
+
+\section{Database Choice}\label{sec:choice}
+% TODO(ryan): Put decision and considereations here.
+%  Schema will go in implementation section.
+%  Incoude some text about how having the documents in JSON (Mongo is actually BSON) makes it really easy to use in JSON.
+
+\section{Implementation}\label{sec:implementation}
+
+\subsection{Backend}
+
+\subsubsection{Database Schema}
+
+\subsection{Frontend}
+
+\subsubsection{Technologies}
+
+\subsubsection{API}
+
+\section{UI Design}\label{sec:ui}
+
+\section{Future Work}\label{sec:future}
+
+\bibliographystyle{acm}
+\bibliography{refs}
 
@@
 
-\begin{figure*}[t]
+\begin{figure*}[h]
    \centering
-   \subfloat[Contig format]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
-   \subfloat[Annotation format]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
+   \subfloat[Contig format]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.png}}
+   \subfloat[Annotation format]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.png}}
    \caption{JSON format for Contigs and Annotations.}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[h]
    \centering
-   \subfloat[User format]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
-   \subfloat[Group format]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
+   \subfloat[User format]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.png}}
+   \subfloat[Group format]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.png}}
    \caption{JSON format for Users and Groups.}
@@
 
-%\begin{figure*}
-%  \includegraphics[height=90mm]{user.eps}
-%	\caption{The JSON for a user.}
-%	\label{fig:JSON-user}
-%\end{figure*}
-%\begin{figure*}
-%  \includegraphics[height=25mm]{group.eps}
-%	\caption{The JSON for a group.}
-%	\label{fig:JSON-group}
-%\end{figure*}
-%\begin{figure*}
-%  \includegraphics[height=80mm]{contig.eps}
-%  \caption{The JSON for a contig.}
-%  \label{fig:JSON-contig}
-%\end{figure*}
-%\begin{figure*}
-%  \includegraphics[height=90mm]{annotation.eps}
-%	\caption{The JSON for an annotation.}
-%	\label{fig:JSON-annotation}
-%\end{figure*}
-
-% TODO: Uncomment if we use any refs
-% \bibliographystyle{acm}
-% \bibliography{refs}
-
 \end{document}

### 2012-12-06 paper/paper.tex
Changed .png to .eps heheheheheheheheh.

@@
    \centering
-   \subfloat[Contig format]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.png}}
-   \subfloat[Annotation format]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.png}}
+   \subfloat[Contig format]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
+   \subfloat[Annotation format]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
    \caption{JSON format for Contigs and Annotations.}
@@
    \centering
-   \subfloat[User format]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.png}}
-   \subfloat[Group format]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.png}}
+   \subfloat[User format]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
+   \subfloat[Group format]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
    \caption{JSON format for Users and Groups.}

### 2012-12-06 paper/paper.tex
Working on implementation

--- 2012-12-06 paper/paper.tex
@@
 
+\section{Database Choice}
+% TODO(ryan): Put decision and considereations here.
+%  Schema will go in implementation section.
+%  Incoude some text about how having the documents in JSON (Mongo is actually BSON) makes it really easy to use in JSON.
+
+\section{Implementation}\label{sec:implementation}
+CGAT uses a standard LAMP stack with MySQL traded out for MongoDB.
+
+\subsection{Backend}
+The \textit{backend} of the system consists of a PHP server connecting to a MongoDB database.
+We do not use any templating systems.
+We chose PHP for the backend because we did not require most of the functionality of a templating system, and we wanted a simple
+system that can be easily understood when passed off to the next generation.
+
+Mongo's PHP connecter is fully featured.
+TODO(eriq): Cite mongo's php connector.
+
+\subsubsection{Database Schema}
+
+\subsubsection{API}
+CGAT is an entireley API driven product. The API is central to all the behavior.
+The API supplies \textbf{all} the information that is displayed in the UI.
+This is a very important design choice. Having the API be the sole data provider and manipulator for the system
+means that all functionality can be programatically provided and different applications can present the same
+fuinctionality with a different UI.
+
+The API accepts GET and PUT requests (depending on the action) and will always return a JSON response.
+
+The current API supports 21 different calls, 7 GET requests and 14 POST actions.
+
+\paragraph{GET Requests}
+\begin{itemize}
+\item administration_info - Provides the user with meta information information that the user would need to perform administrative tasks such as greating a group or assigning a task.
+\item annotation - Provide the user with information about an specified annotation.
+\item contig - Get information about a specific contig.
+\item gene - Get information about a specific gene.
+\item group - Get information about a specific group.
+\item help - Get help information on a specific topic.
+\item user_profile - Get the profile information about a specific user. If the user is logged in and requests information about themselves, more information is provided.
+\end{itemize}
+
+\paragraph{POST Actions}
+Most POST actions require that the user is logged in.
+
+\begin{itemize}
+\item assign_task - Ask a group of people to perform an annotation.
+\item cancel_notification - Clear a notification from a user.
+\item create_annotation - Create a new annotation of a contig.
+\item create_group - Create a group.
+\item join_group - Join a group.
+\item leave_group - Leave a group.
+\item login - Login.
+\item logout - Logout.
+\item parse_fasta - Upload a FASTA file to the server, parse the contents, and returned the parsed structure.
+\item register - Register a new user.
+\item save_annotation - Save a work-in-progress annotation.
+\item set_help - Upload a help page.
+\item submit_annotation - Submit a finalized annotaion.
+\item upload_contig - Uplaod a contig into CGAT.
+\end{itemize}
+
+\subsection{Frontend}
+
+\section{UI Design}
+
 \onecolumn

### 2012-12-06 paper/paper.tex

--- 2012-12-06 paper/paper.tex
@@
 \section{Implementation}\label{sec:implementation}
+CGAT uses a standard LAMP stack with MySQL traded out for MongoDB.
 
 \subsection{Backend}
+The \textit{backend} of the system consists of a PHP server connecting to a MongoDB database.
+We do not use any templating systems.
+We chose PHP for the backend because we did not require most of the functionality of a templating system, and we wanted a simple
+system that can be easily understood when passed off to the next generation.
+
+Mongo's PHP connecter is fully featured.
+TODO(eriq): Cite mongo's php connector.
 
@@
 
+\subsubsection{API}
+CGAT is an entireley API driven product. The API is central to all the behavior.
+The API supplies \textbf{all} the information that is displayed in the UI.
+This is a very important design choice. Having the API be the sole data provider and manipulator for the system
+means that all functionality can be programatically provided and different applications can present the same
+fuinctionality with a different UI.
+
+The API accepts GET and PUT requests (depending on the action) and will always return a JSON response.
+
+The current API supports 21 different calls, 7 GET requests and 14 POST actions.
+
+\paragraph{GET Requests}
+\begin{itemize}
+\item administration_info - Provides the user with meta information information that the user would need to perform administrative tasks such as greating a group or assigning a task.
+\item annotation - Provide the user with information about an specified annotation.
+\item contig - Get information about a specific contig.
+\item gene - Get information about a specific gene.
+\item group - Get information about a specific group.
+\item help - Get help information on a specific topic.
+\item user_profile - Get the profile information about a specific user. If the user is logged in and requests information about themselves, more information is provided.
+\end{itemize}
+
+\paragraph{POST Actions}
+Most POST actions require that the user is logged in.
+
+\begin{itemize}
+\item assign_task - Ask a group of people to perform an annotation.
+\item cancel_notification - Clear a notification from a user.
+\item create_annotation - Create a new annotation of a contig.
+\item create_group - Create a group.
+\item join_group - Join a group.
+\item leave_group - Leave a group.
+\item login - Login.
+\item logout - Logout.
+\item parse_fasta - Upload a FASTA file to the server, parse the contents, and returned the parsed structure.
+\item register - Register a new user.
+\item save_annotation - Save a work-in-progress annotation.
+\item set_help - Upload a help page.
+\item submit_annotation - Submit a finalized annotaion.
+\item upload_contig - Uplaod a contig into CGAT.
+\end{itemize}
+
 \subsection{Frontend}
-
-\subsubsection{Technologies}
-
-\subsubsection{API}
 

### 2012-12-06 paper/paper.tex
Working on implementation section in paper.

@@
 \section{Implementation}\label{sec:implementation}
-CGAT uses a standard LAMP stack with MySQL traded out for MongoDB.
+CGAT uses a standard LAMP (Linux Apache MySQL PHP) stack with MySQL traded out for MongoDB.
 
@@
 
-Mongo's PHP connecter is fully featured.
-TODO(eriq): Cite mongo's php connector.
+PHP and MongoDB work together very nicely and MongoDB's PHP connector is fully featured \cite{phpMongo}.
 
 \subsubsection{Database Schema}
+Although we use MongoDB which does not enforce a specific schema, by convention we enforce a schema on our documents.
+MongoDB uses BSON (Binary JSON). BSON is an extended JSON format that supports additional types like a datetimes \cite{bson}.
+
+The schemas for our four JSON objects are shown in Figures \ref{fig:JSON-contig}, \ref{fig:JSON-annotation},
+\ref{fig:JSON-user}, and \ref{fig:JSON-group} in Appendix \ref{sec:data_models}.
 
 \subsubsection{API}
-CGAT is an entireley API driven product. The API is central to all the behavior.
+CGAT is an entirely API driven product. The API is central to all the behavior.
 The API supplies \textbf{all} the information that is displayed in the UI.
 This is a very important design choice. Having the API be the sole data provider and manipulator for the system
-means that all functionality can be programatically provided and different applications can present the same
-fuinctionality with a different UI.
+means that all functionality can be programmatically provided and different applications can present the same
+functionality with a different UI.
 
@@
 \begin{itemize}
-\item administration_info - Provides the user with meta information information that the user would need to perform administrative tasks such as greating a group or assigning a task.
+\item administration_info - Provides the user with meta information information that the user would need to perform administrative tasks such as creating a group or assigning a task.
 \item annotation - Provide the user with information about an specified annotation.
@@
 \item set_help - Upload a help page.
-\item submit_annotation - Submit a finalized annotaion.
-\item upload_contig - Uplaod a contig into CGAT.
+\item submit_annotation - Submit a finalized annotation.
+\item upload_contig - Upload a contig into CGAT.
 \end{itemize}
@@
 \subsection{Frontend}
+CGAT's frontend uses standard web technologies (HTML, Javascript, and CSS).
+CGAT uses many HTML5/CSS3 features such as transforms and pseudo-elements.
 
-\section{UI Design}\label{sec:ui}
+Because the entire system is API based, the frontend heavily relies on asynchronous fetches (AJAX).
 

### 2012-12-06 paper/paper.tex
Fixed latex error

@@
 \begin{itemize}
-\item administration_info - Provides the user with meta information information that the user would need to perform administrative tasks such as creating a group or assigning a task.
+\item administration\_info - Provides the user with meta information information that the user would need to perform administrative tasks such as creating a group or assigning a task.
 \item annotation - Provide the user with information about an specified annotation.
@@
 \item help - Get help information on a specific topic.
-\item user_profile - Get the profile information about a specific user. If the user is logged in and requests information about themselves, more information is provided.
+\item user\_profile - Get the profile information about a specific user. If the user is logged in and requests information about themselves, more information is provided.
 \end{itemize}
@@
 \begin{itemize}
-\item assign_task - Ask a group of people to perform an annotation.
-\item cancel_notification - Clear a notification from a user.
-\item create_annotation - Create a new annotation of a contig.
-\item create_group - Create a group.
-\item join_group - Join a group.
-\item leave_group - Leave a group.
+\item assign\_task - Ask a group of people to perform an annotation.
+\item cancel\_notification - Clear a notification from a user.
+\item create\_annotation - Create a new annotation of a contig.
+\item create\_group - Create a group.
+\item join\_group - Join a group.
+\item leave\_group - Leave a group.
 \item login - Login.
 \item logout - Logout.
-\item parse_fasta - Upload a FASTA file to the server, parse the contents, and returned the parsed structure.
+\item parse\_fasta - Upload a FASTA file to the server, parse the contents, and returned the parsed structure.
 \item register - Register a new user.
-\item save_annotation - Save a work-in-progress annotation.
-\item set_help - Upload a help page.
-\item submit_annotation - Submit a finalized annotation.
-\item upload_contig - Upload a contig into CGAT.
+\item save\_annotation - Save a work-in-progress annotation.
+\item set\_help - Upload a help page.
+\item submit\_annotation - Submit a finalized annotation.
+\item upload\_contig - Upload a contig into CGAT.
 \end{itemize}

### 2012-12-06 paper/paper.tex
Added first version of Database choice section.

@@
 \section{Motivation}\label{sec:motive}
-Dr. Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
+Dr. Anya Goodman, a professor in the Chemistry department at Cal Poly, San Luis
 Obispo, offers a bioinformatics course that covers several aspects of gene
@@
 \section{Database Choice}\label{sec:choice}
-% TODO(ryan): Put decision and considereations here.
-%  Schema will go in implementation section.
-%  Incoude some text about how having the documents in JSON (Mongo is actually BSON) makes it really easy to use in JSON.
+The choice for backend database was between MySQL, Couchbase, and MongoDB.
+We included MySQL in the options because its the standard that small websites tend to use. Then we included both
+Couchbase and MongoDB in our considerations because they are popular NoSQL database that are different types of NoSQL
+databases. Couchbase is a key-value store unlike MongoDB which is a document-store.
+
+We ended up using MongoDB for several reasons. Firstly, the document structure was extremely helpful when storing
+relation. We ended up attaching a lot of data on our relations that would of been a pain to store in SQL but being
+able to store the data as complex objects in JSON was super helpful. Secondly, being able to store the data in JSON
+(BSON in MongoDB) made it extremely easy to retrieve and use the data in a web application. Because of these two reasons
+we narrowed our choices down to MongoDB and Couchbase.
+
+When we were deciding between MongoDB and Couchbase we looked at the features they 
+provided and noticed several things. First, Couchbase currently is a pure key-value store it has
+no idea what the structure of the value. Unlike MongoDB, which can query for fields inside an 
+object specified by a key. This feature of MongoDB makes updating relations (which we 
+have a lot of) super easy. Second, Couchbase provides a huge
+cluster/map reduce overhead that is not necessary for our data.
 

### 2012-12-06 paper/paper.tex
Added background section for Aldrin. Finished the first drafts of results/future work. Did some light editing on my stuff.

@@
 This paper discusses the progress that has been made this quarter and the future work that 
-still needs to be done. The rest of the paper is as follows. Section \ref{sec:motive} discusses the
+still needs to be done. The rest of the paper is as follows. Section \ref{sec:background} discusses
+the common terms and biological meanings used in the project. Section \ref{sec:motive} discusses the
 motivation for creating CGAT. Section \ref{sec:choice} explains why we chose MongoDB
 as our backend. Section \ref{sec:implementation} goes into the details of how
-our backend and frontend are implemented currently. Section \ref{sec:ui} explains the
-various features of our UI. Lastly, section \ref{sec:future} describes the future work still left to
-do.
+our backend and frontend are implemented currently. Section \ref{sec:results} discusses
+the results of our demo to our customer. Lastly, section \ref{sec:future} 
+describes the future work still left to do.
+
+\section{Background}\label{sec:background}
 
@@
 The choice for backend database was between MySQL, Couchbase, and MongoDB.
-We included MySQL in the options because its the standard that small websites tend to use. Then we included both
-Couchbase and MongoDB in our considerations because they are popular NoSQL database that are different types of NoSQL
-databases. Couchbase is a key-value store unlike MongoDB which is a document-store.
+We included MySQL in the options because its the standard that websites tend to use. Then we included both
+Couchbase and MongoDB in our considerations because they are popular NoSQL database of different types of NoSQL
+databases. Couchbase is a key-value store whereas MongoDB which is a document-store.
 
-We ended up using MongoDB for several reasons. Firstly, the document structure was extremely helpful when storing
-relation. We ended up attaching a lot of data on our relations that would of been a pain to store in SQL but being
-able to store the data as complex objects in JSON was super helpful. Secondly, being able to store the data in JSON
+We ended up using MongoDB for several reasons. First, the document structure was extremely helpful when storing
+relations. We ended up attaching a lot of data on our relations that would of been a pain to store in SQL but being
+able to store the data as complex objects in JSON was super helpful. Second, being able to store the data in JSON
 (BSON in MongoDB) made it extremely easy to retrieve and use the data in a web application. Because of these two reasons
@@
 have a lot of) super easy. Second, Couchbase provides a huge
-cluster/map reduce overhead that is not necessary for our data.
+cluster/map reduce overhead that is not necessary for our data. Therefore we choose to use
+MongoDB for the CGAT database.
 
@@
 
+\section{Results}\label{sec:results}
+We demoed CGAT to Dr. Anya Goodman on the 19th of November. She was pleased with
+the work accomplished so far. But she had numerous recommendations for things to fix and update.
+We have taken her ideas and attempted to get as much fixed in the last weeks of the quarter.
+Some of the notable fixes were removing some of the "prettiness" from the site, making
+buttons look more like buttons, and changing some of the words and terminology used on the site.
+
 \section{Future Work}\label{sec:future}
+The future work that is left to be done on CGAT falls into several categories.
+
+\subsection{Missing use cases}
+We still need to add some of the main use cases that Dr. Goodman wanted. This includes things like
+user leveling and immediate feedback on annotations.
+
+\subsection{Bug fixes}
+We know of several bugs that still need to be fixed based on feedback from the demo with
+Dr. Goodman.
+
+\subsection{New features}
+New features for CGAT will most likely be desired after user feedback.
+
+\subsection{Accessibility}
+Need to test the accessibility of the site. This includes things like the color set used
+and text size.
 

### 2012-12-06 paper/paper.tex
In the process of expanding the introduction of the paper.

--- 2012-12-06 paper/paper.tex
@@
 \section{Introduction}\label{sec:introduction}
+Genomics research attempts to understand information maintained in the genomes
+of various organisms. A \textit{genome} is a sequence of DNA containing all
+information for an organism. Genomes contain several types of regions,
+\textit{exons}, \textit{introns}, \textit{CDS}s,   In order to link biological information of the organism
+to genomic information in the DNA sequence... to  To understand this information, it
+\cite{annotation}
+
 Gene annotation is the process of associating metadata about a gene with the
@@
 \begin{itemize}
-\item administration\_info - Provides the user with meta information information that the user would need to perform administrative tasks such as creating a group or assigning a task.
+\item administration_info - Provides the user with meta information information that the user would need to perform administrative tasks such as creating a group or assigning a task.
 \item annotation - Provide the user with information about an specified annotation.
@@
 \item help - Get help information on a specific topic.
-\item user\_profile - Get the profile information about a specific user. If the user is logged in and requests information about themselves, more information is provided.
+\item user_profile - Get the profile information about a specific user. If the user is logged in and requests information about themselves, more information is provided.
 \end{itemize}
@@
 \begin{itemize}
-\item assign\_task - Ask a group of people to perform an annotation.
-\item cancel\_notification - Clear a notification from a user.
-\item create\_annotation - Create a new annotation of a contig.
-\item create\_group - Create a group.
-\item join\_group - Join a group.
-\item leave\_group - Leave a group.
+\item assign_task - Ask a group of people to perform an annotation.
+\item cancel_notification - Clear a notification from a user.
+\item create_annotation - Create a new annotation of a contig.
+\item create_group - Create a group.
+\item join_group - Join a group.
+\item leave_group - Leave a group.
 \item login - Login.
 \item logout - Logout.
-\item parse\_fasta - Upload a FASTA file to the server, parse the contents, and returned the parsed structure.
+\item parse_fasta - Upload a FASTA file to the server, parse the contents, and returned the parsed structure.
 \item register - Register a new user.
-\item save\_annotation - Save a work-in-progress annotation.
-\item set\_help - Upload a help page.
-\item submit\_annotation - Submit a finalized annotation.
-\item upload\_contig - Upload a contig into CGAT.
+\item save_annotation - Save a work-in-progress annotation.
+\item set_help - Upload a help page.
+\item submit_annotation - Submit a finalized annotation.
+\item upload_contig - Upload a contig into CGAT.
 \end{itemize}

### 2012-12-07 paper/paper.tex
Made the introduction, motivation, and background awesome. I think. Also added references so that we seem more legit

--- 2012-12-06 paper/paper.tex
@@
 
+% TODO include more background:
+% http://www.weizmann.ac.il/complex/tlusty/courses/landmark/JacobMonod1960.pdf
+% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2443188/
+% http://www.sciencemag.org/content/suppl/2012/04/11/336.6078.179.DC1/Zerbino.SM.pdf
+% http://gep.wustl.edu/community/publications
+% http://gep.wustl.edu/projects/dot_chromosome
 \section{Introduction}\label{sec:introduction}
-Gene annotation is the process of associating metadata about a gene with the
-contig, a dna sequence, on which the gene resides.\cite{annotation} This metadata is necessary for conducting
-genomic research projects and analysis involving the contig's genomic sequence.
-Currently, there is no specialty gene annotation software. Genome browsers
-all maintain lots of genomic data and are used when manually annotating genes,
-but do not provide a user-friendly interface. While genome browsers
-will likely never be replaced (UCSC genome browser, etc. are well
-established\cite{ucscbrowser, ncbi}), it would be desirable to have a software system that better
-accommodates the visual and informational needs of gene annotation.
+Dr. Anya Goodman, a professor in the department of chemistry and biochemistry
+at Cal Poly, San Luis Obispo, offers a bioinformatics course that covers
+several aspects of gene annotation and genomic research. Her work is a part of
+the \textit{Genomics Education Partnership} (GEP) which is based at Washington
+University in St. Louis~\cite{gep}. GEP is an organization that aims to bring
+genomics research to academic institutions, especially undergraduate classes.
+This allows undergraduate students to get the opportunity to learn about, and
+practice, basic genomic research and, in turn, contribute meaningful work to
+the biological research community. However, GEP does not have an existing
+system in which users can input and save gene annotations and get immediate
+feedback regarding their performance. Similarly, available gene annotation
+software--RepeatMasker, Artemis, and others~\cite{repeatmasker, artemis}--is
+not designed to provide immediate feedback for users regarding their
+performance. In this paper, we present Community Genome Annotation Trainer,
+\textit{CGAT}, a web-based gene annotation application designed to provide
+feedback of gene annotation quality to students faster.
 
-\textit{CGAT} is a web-based gene annotation application designed with
-usability, simplicity, and efficiency in mind. It is not a feature-rich and data-rich
- genome browser like the \textit{UCSC genome brower}. Nor is \textit{CGAT}
-designed to be a replacement for other genome browsers in any aspect other than
-gene annotation. Other gene browsers attempt to accommodate many needs of the
-biology community and so offer many features, and a cluttered, chaotic
-interface. Gene annotation, while not a trivial task, can be well-serviced by a
-simple data model. Ideally, by focusing on a simple data model, \textit{CGAT}
-is able to provide a clean, streamlined experience. \textit{CGAT} aims to
-provide users with a way to view gene annotations without extra, unnecessary
-information obscuring their view while also embodying a Wikipedia-like emphasis
-on collaboration and openness.
+\textit{CGAT} is spearheaded by Ryan Verdon and has been built with usability,
+simplicity, and efficiency in mind. It is not a feature-rich and data-rich
+genome browser like \textit{UCSC genome brower}~\cite{ucscbrowser},
+\textit{Ensembl}~\cite{ensembl}, or \textit{BioViz}~\cite{bioviz}, nor is it
+designed to be a replacement for gene annotatioin software such as
+RepeatMasker, Artemis, or any others, in any aspect other than for teaching, or
+practicing, gene annotation. Other gene browsers or annotation software offer
+many features to accommodate the needs of the biology community, and so tend to
+be complex. Gene annotation, while not a trivial task, can be well-serviced by
+a simple data model. Ideally, by focusing on a simple data model, \textit{CGAT}
+is able to provide a clean, streamlined experience capable of providing users
+with a way to view gene annotations made by others or have their gene
+annotations reviewed by experts. In particular, \textit{CGAT} strives towards a
+Wikipedia-like emphasis on collaboration and openness.
 
-This paper discusses the progress that has been made this quarter and the future work that 
-still needs to be done. The rest of the paper is as follows. Section \ref{sec:background} discusses
-the common terms and biological meanings used in the project. Section \ref{sec:motive} discusses the
-motivation for creating CGAT. Section \ref{sec:choice} explains why we chose MongoDB
-as our backend. Section \ref{sec:implementation} goes into the details of how
-our backend and frontend are implemented currently. Section \ref{sec:results} discusses
-the results of our demo to our customer. Lastly, section \ref{sec:future} 
-describes the future work still left to do.
+This paper discusses the progress that has been made during a graduate database
+class at Cal Poly, San Luis Obispo on \textit{CGAT} as well as relevant future
+work. The rest of the paper is as follows. Section \ref{sec:background}
+discusses the common terms and biological meanings used in the project. Section
+\ref{sec:motive} discusses the motivation for creating CGAT. Section
+\ref{sec:choice} explains why we chose MongoDB as our backend. Section
+\ref{sec:implementation} goes into the details of how our backend and frontend
+are implemented currently. Section \ref{sec:results} discusses the results of
+our demo to our customer. Lastly, section \ref{sec:future} describes the future
+work still left to do.
 
 \section{Background}\label{sec:background}
-
-\section{Motivation}\label{sec:motive}
-Dr. Anya Goodman, a professor in the Chemistry department at Cal Poly, San Luis
-Obispo, offers a bioinformatics course that covers several aspects of gene
-annotation. Her work is a part of the Genomics Education Partnership (GEP) at 
-Washington University in St. Louis.\cite{gep} The goal of GEP is to teach undergrad students
-in Biology how to annotate genes and other features in raw sequence data.
-The CGAT project is spearheaded by Ryan Verdon to accommodate Dr.
-Goodman's needs and ideas for ideal gene annotation software.
+Genomics research attempts to understand information contained in the
+\textit{genomes} of various organisms. The genome is the entire DNA sequence
+that encodes information and drives all cellular processes in an organism.
+However, due to the size of genomes, biologists are only able to sequence
+sections, or fragments, of a genome at a time. One or more contiguous fragments
+of the genome together form a \textit{contig}, or simply a contiguous sequence
+of DNA nucleotides. The genome contains several components, or sub-regions,
+that have different functions or roles. These components describe in detail how
+a genome is decoded to drive cellular processes. \textit{Annotations}--the
+process of linking biological information of the organism to the genome--are an
+important step to being able to relate information learned from the sequence back
+to broader understandings of the associated organism and, ideally, related
+organisms~\cite{annotation}. Again, since the genome is so large, biologists
+typically annotate sections of the genome, specifically \textit{genes}. A gene
+is a section of the genome that codes for a functional \textit{protein} used
+for biological processes inside and outside of the cell. A gene contains
+several components, or sub-regions, that have different functions or
+roles--\textit{exons}, \textit{introns}, \textit{CDS}s, \textit{UTR}s, and
+others. These components that are located in a gene some of the most basic but
+most important pieces of information that must be annotated in order to be able
+to make meaningful biological interpretations from information learned from the
+genomic sequence, and where our project, \textit{CGAT}, is useful. Considering
+the scope of an undergraduate genomics research class, it is preferable to have
+a system that emphasizes simplicity and usability as well as houses information
+in a collaborative and educational environment.
 

### 2012-12-07 paper/paper.tex
forgot to change a few things, specifically the fact that I removed the motivation section and I fixed something that latex couldn't handle (a really long line)
Fixed some words in the paper, otherwise its good.

@@
 
-% TODO include more background:
-% http://www.weizmann.ac.il/complex/tlusty/courses/landmark/JacobMonod1960.pdf
-% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2443188/
-% http://www.sciencemag.org/content/suppl/2012/04/11/336.6078.179.DC1/Zerbino.SM.pdf
-% http://gep.wustl.edu/community/publications
-% http://gep.wustl.edu/projects/dot_chromosome
 \section{Introduction}\label{sec:introduction}
@@
 several aspects of gene annotation and genomic research. Her work is a part of
-the \textit{Genomics Education Partnership} (GEP) which is based at Washington
+the \textit{Genomics Education Partnership} (GEP), based at Washington
 University in St. Louis~\cite{gep}. GEP is an organization that aims to bring
@@
 not designed to provide immediate feedback for users regarding their
-performance. In this paper, we present Community Genome Annotation Trainer,
+performance. In this paper, we present Community Genome Annotation Training database,
 \textit{CGAT}, a web-based gene annotation application designed to provide
@@
 discusses the common terms and biological meanings used in the project. Section
-\ref{sec:motive} discusses the motivation for creating CGAT. Section
 \ref{sec:choice} explains why we chose MongoDB as our backend. Section
@@
 When we were deciding between MongoDB and Couchbase we looked at the features they 
-provided and noticed several things. First, Couchbase currently is a pure key-value store it has
-no idea what the structure of the value. Unlike MongoDB, which can query for fields inside an 
+provided and noticed several things. First, Couchbase currently is a pure key-value store. Couchbase has
+no idea what the structure of the value is. Unlike MongoDB, which can query for fields inside an 
 object specified by a key. This feature of MongoDB makes updating relations (which we 

### 2012-12-07 paper/paper.tex
Adding IEEEtran.bst and modified paper.tex so that some of the ridiculous citations (Ensembl and Artemis I think) have et al instead of a list of like 50 authors.

@@
 \pagestyle{empty}
+\bstctlcite{test}
 
@@
 genomics research to academic institutions, especially undergraduate classes.
-This allows undergraduate students to get the opportunity to learn about, and
+This allows undergraduate students the opportunity to learn about, and
 practice, basic genomic research and, in turn, contribute meaningful work to
@@
 not designed to provide immediate feedback for users regarding their
-performance. In this paper, we present Community Genome Annotation Training database,
+performance. In this paper, we present Community Genome Annotation Trainer,
 \textit{CGAT}, a web-based gene annotation application designed to provide
-feedback of gene annotation quality to students faster.
+students with feedback on the quality of their gene annotations faster.
 
@@
 are implemented currently. Section \ref{sec:results} discusses the results of
-our demo to our customer. Lastly, section \ref{sec:future} describes the future
-work still left to do.
+our demo to our customer. Lastly, section \ref{sec:future} describes our future
+work.
 
@@
 When we were deciding between MongoDB and Couchbase we looked at the features they 
-provided and noticed several things. First, Couchbase currently is a pure key-value store. Couchbase has
-no idea what the structure of the value is. Unlike MongoDB, which can query for fields inside an 
+provided and noticed several things. First, Couchbase currently is a pure key-value store it has
+no idea what the structure of the value. Unlike MongoDB, which can query for fields inside an 
 object specified by a key. This feature of MongoDB makes updating relations (which we 
@@
 
-\bibliographystyle{acm}
+%\bibliographystyle{acm}
+\bibliographystyle{IEEEtran}
 \bibliography{refs}
